---
title: "DiffuChannel: comparison of real and diffusion-generated single-channel records"
author: "Barrett-Jolley lab"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width  = 7,
  fig.height = 4.2,
  dpi = 150
)
```

# Purpose

This notebook takes real (seed) patch-clamp records and DiffuChannel-generated records,
computes every similarity metric reported in the manuscript, draws every figure, and writes
a `results_values.csv` whose rows are keyed to the `[[placeholder]]` names used in the
manuscript draft. Fill the manuscript from that file rather than from the console.

Nothing here assumes a particular result. If the model has not learned the noise structure,
the plots will say so.

# Packages

```{r packages}
required <- c("dplyr", "tidyr", "ggplot2", "readr", "purrr", "arrow",
              "scales", "patchwork", "uwot", "Rtsne")

# Optional: only used for the file dialogs. rstudioapi ships with RStudio and
# tcltk with base R on most builds, so this normally needs nothing installed.
if (!any(vapply(c("rstudioapi", "tcltk"), requireNamespace,
                logical(1), quietly = TRUE))) {
  message("Neither rstudioapi nor tcltk is available, so file dialogs will ",
          "fall back to file.choose(). Set the paths in CFG to avoid prompts.")
}

missing <- required[!vapply(required, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing)) {
  message("Install these before running: ", paste(missing, collapse = ", "))
  message('install.packages(c("', paste(missing, collapse = '", "'), '"))')
}

library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(purrr)
library(arrow)
library(patchwork)

set.seed(42)
```

# Configuration

Everything you need to change lives in this block.

```{r config}
CFG <- list(

  # ---- where the data are -------------------------------------------------
  real_dir       = "./pages",   # folder of seed CSV / Parquet. Always a folder.

  # The two generated records are picked with file dialogs on first run and the
  # choice is remembered afterwards. Options for each:
  #   NA      prompt with a dialog (cancel to skip that arm)
  #   "path"  use that file, no dialog
  #   FALSE   skip that arm entirely, no dialog
  gen_file        = NA,      # DDPM / DiffuChannel output
  gan_file        = NA,      # DeepGANnel output, optional
  gen_single_file = FALSE,   # single-stage ablation, optional

  reset_file_choices = FALSE, # TRUE forgets remembered paths and asks again
  filter_dialogs = TRUE,      # FALSE shows every file in the chooser

  # Meta-analysis. Each run of this notebook writes one tidy file into runs_dir;
  # DiffuChannel_meta.Rmd reads every file in that folder. run_id must be unique
  # per seed record or runs will overwrite one another. NA derives it from the
  # seed folder name.
  run_id         = NA,
  runs_dir       = "./runs",

  out_dir        = "./figures",

  # ---- recording parameters ----------------------------------------------
  sample_hz      = 1e4,     # acquisition rate, Hz. Must match SAMPLE_DT in diffusion_app.py
  filter_hz      = 1e3,     # anti-alias / Bessel corner frequency, Hz

  # ---- must match how the model was trained -------------------------------
  seq_len        = 128,     # training window, samples
  slow_factor    = 64,      # decimation factor of the coarse stage
  context_len    = 512,     # carried context at block joins
  block_len      = 32768,   # generation block length
  burst_tc       = 1000,    # BURST_TC in diffusion_app.py: flicker threshold, samples
  emb_dim        = 4,       # per-record embedding width used by the coarse stage
  emb_drop       = 0.1,     # embedding dropout during coarse training

  # ---- analysis knobs ------------------------------------------------------
  n_windows_mmd  = 400,     # windows per group for MMD. O(n^2): keep modest
  n_windows_proj = 600,     # windows per group for UMAP / t-SNE
  proj_win       = 512,     # window length for manifold projections
  psd_seg        = 1024,    # Welch segment length
  acf_lag_max    = 200,     # autocorrelation lag, samples
  ks_subsample   = 2e5,     # cap for KS tests (ks.test is slow on millions)

  # ---- plotting ------------------------------------------------------------
  trace_secs     = 0.5,     # duration of example traces to plot
  base_size      = 10
)

dir.create(CFG$out_dir, showWarnings = FALSE, recursive = TRUE)

# Results are collected here and written out at the end.
RESULTS <- new.env(parent = emptyenv())
RESULTS$rows <- list()

record_value <- function(key, value, note = "") {
  RESULTS$rows[[key]] <- tibble::tibble(
    placeholder = key,
    value       = if (is.numeric(value)) format(value, digits = 4) else as.character(value),
    note        = note
  )
  invisible(value)
}
```

# Theme

A plain publication theme. No decoration that a reviewer will ask you to remove.

```{r theme}
theme_channel <- function(base_size = CFG$base_size) {
  theme_classic(base_size = base_size) +
    theme(
      strip.background = element_blank(),
      strip.text       = element_text(face = "bold", hjust = 0),
      legend.position  = "top",
      legend.title     = element_blank(),
      legend.key.size  = unit(0.8, "lines"),
      plot.title       = element_text(face = "bold", size = base_size + 1),
      axis.text        = element_text(colour = "black")
    )
}

pal_source <- c(
  "Real"          = "#1b1b1b",
  "Generated"     = "#c0392b",
  "Single-stage"  = "#2980b9",
  "GAN"           = "#8e44ad",
  "White noise"   = "#7f8c8d"
)

save_fig <- function(plot, name, width = 7, height = 4.2) {
  ggsave(file.path(CFG$out_dir, paste0(name, ".pdf")), plot,
         width = width, height = height)
  ggsave(file.path(CFG$out_dir, paste0(name, ".png")), plot,
         width = width, height = height, dpi = 300)
  invisible(plot)
}
```

# Helper functions

## Reading records

Handles both the tidy column names and whatever the acquisition software produced.

```{r fn-read}
normalise_names <- function(nms) {
  gsub("[_-]", " ", trimws(tolower(nms)))
}

read_record <- function(path) {
  df <- if (grepl("\\.parquet$", path, ignore.case = TRUE)) {
    as.data.frame(arrow::read_parquet(path))
  } else {
    as.data.frame(readr::read_csv(path, show_col_types = FALSE))
  }

  nn  <- normalise_names(names(df))
  idx <- function(target) {
    hit <- which(nn %in% target)
    if (length(hit)) hit[1] else NA_integer_
  }

  i_chan <- idx(c("channels", "channel", "ideal", "idealisation", "idealization"))
  i_cur  <- idx(c("noisy current", "current", "raw", "raw current", "i"))
  i_time <- idx(c("time", "t"))

  if (is.na(i_chan) || is.na(i_cur)) {
    stop("Could not find Channels / Noisy Current columns in ", basename(path),
         ". Found: ", paste(names(df), collapse = ", "))
  }

  out <- tibble::tibble(
    channels = as.numeric(df[[i_chan]]),
    current  = as.numeric(df[[i_cur]])
  )
  out$time <- if (!is.na(i_time)) as.numeric(df[[i_time]]) else
    (seq_len(nrow(out)) - 1) / CFG$sample_hz

  out |>
    dplyr::filter(is.finite(channels), is.finite(current)) |>
    dplyr::select(time, channels, current)
}

# Sample rate taken from the file's own Time column rather than from CFG. The
# app's export "Decimation factor" multiplies the Time step, so a generated
# file can sit at a different rate from the seed data, and PSD, ACF and dwell
# times are all silently wrong if that goes unnoticed.
fs_of <- function(d, fallback = CFG$sample_hz) {
  if (nrow(d) < 3) return(fallback)
  dt <- stats::median(diff(d$time), na.rm = TRUE)
  if (!is.finite(dt) || dt <= 0) return(fallback)
  1 / dt
}

# ---- picking the generated files ------------------------------------------
# Dialogs are convenient interactively and useless when knitting, because Knit
# runs in a separate non-interactive session where no dialog can appear. So the
# choice is made once, remembered in a small file next to the figures, and
# reused on every subsequent knit. Delete that file, or set
# CFG$reset_file_choices = TRUE, to be asked again.

CHOICE_CACHE <- file.path(CFG$out_dir, "chosen_files.rds")
if (isTRUE(CFG$reset_file_choices) && file.exists(CHOICE_CACHE)) {
  unlink(CHOICE_CACHE)
  message("Forgot the remembered file choices.")
}

read_choices <- function() {
  if (!file.exists(CHOICE_CACHE)) return(list())
  tryCatch(readRDS(CHOICE_CACHE), error = function(e) list())
}

# File-type filters are the usual reason a chooser opens on an empty folder, so
# they are built carefully here and can be switched off with CFG$filter_dialogs.
#
# tk_choose.files groups its filter matrix by DESCRIPTION, collecting every
# extension that shares a label. So several extensions need one row each with
# the label repeated -- "{.csv .parquet}" in a single row is passed through as
# one literal extension and matches nothing at all. Uppercase variants are
# listed separately because Tcl's matching is case sensitive on some platforms.
tk_filters <- function() {
  matrix(c("Records",   ".csv",
           "Records",   ".parquet",
           "Records",   ".CSV",
           "Records",   ".PARQUET",
           "All files", "*"),
         ncol = 2, byrow = TRUE)
}

choose_one <- function(prompt) {
  if (!interactive()) return(NA_character_)
  use_filter <- !isFALSE(CFG$filter_dialogs)
  pick <- NULL

  if (requireNamespace("rstudioapi", quietly = TRUE) &&
      isTRUE(tryCatch(rstudioapi::isAvailable(), error = function(e) FALSE))) {
    # RStudio expects "Label (pattern)" with patterns separated by spaces, not
    # semicolons. A wrong pattern here hides every file rather than erroring.
    filt <- if (use_filter) "Records (*.csv *.parquet)" else "All Files (*)"
    pick <- tryCatch(
      rstudioapi::selectFile(caption = prompt, existing = TRUE, filter = filt),
      error = function(e) NULL)
    # If the filter hid everything, offer the unfiltered dialog rather than
    # leaving the user staring at an empty folder.
    if (use_filter && (is.null(pick) || !length(pick) || is.na(pick[1]))) {
      pick <- tryCatch(
        rstudioapi::selectFile(caption = paste0(prompt, " (all files)"),
                               existing = TRUE, filter = "All Files (*)"),
        error = function(e) NULL)
    }

  } else if (requireNamespace("tcltk", quietly = TRUE)) {
    pick <- tryCatch(
      if (use_filter) {
        tcltk::tk_choose.files(caption = prompt, multi = FALSE,
                               filters = tk_filters())
      } else {
        tcltk::tk_choose.files(caption = prompt, multi = FALSE)
      },
      error = function(e) NULL)
    if (use_filter && (is.null(pick) || !length(pick) || is.na(pick[1]) ||
                       !nzchar(pick[1]))) {
      pick <- tryCatch(
        tcltk::tk_choose.files(caption = paste0(prompt, " (all files)"),
                               multi = FALSE),
        error = function(e) NULL)
    }

  } else {
    message(prompt, "  (cancel to skip)")
    pick <- tryCatch(file.choose(), error = function(e) NULL)
  }

  if (is.null(pick) || !length(pick) || is.na(pick[1]) || !nzchar(pick[1])) {
    NA_character_
  } else {
    pick[1]
  }
}

pick_dataset <- function(key, prompt, cfg_value) {
  if (isFALSE(cfg_value)) return(NA_character_)

  if (is.character(cfg_value) && length(cfg_value) == 1 && nzchar(cfg_value)) {
    if (file.exists(cfg_value)) return(cfg_value)
    warning("CFG path for ", key, " does not exist: ", cfg_value,
            ". Falling back to a dialog.")
  }

  ch <- read_choices()
  if (!is.null(ch[[key]]) && file.exists(ch[[key]])) {
    message("Using remembered ", key, ": ", basename(ch[[key]]))
    return(ch[[key]])
  }

  path <- choose_one(prompt)
  if (is.na(path)) {
    message(if (interactive()) "Cancelled: " else "No remembered choice and no dialog available: ",
            key, ". That arm will be skipped.")
    return(NA_character_)
  }
  ch[[key]] <- path
  try(saveRDS(ch, CHOICE_CACHE), silent = TRUE)
  message("Remembered ", key, ": ", basename(path))
  path
}

read_folder <- function(dir) {
  files <- list.files(dir, pattern = "\\.(csv|parquet)$", full.names = TRUE,
                      ignore.case = TRUE) |> sort()

  # A GAN output living alongside the seed data would otherwise be loaded as
  # though it were seed data, quietly contaminating every "Real" statistic.
  gan_like <- grepl("_GAN\\.(csv|parquet)$", files, ignore.case = TRUE)
  if (any(gan_like)) {
    message("Ignoring ", sum(gan_like), " *_GAN file(s) in ", dir,
            " when loading seed records.")
    files <- files[!gan_like]
  }

  if (!length(files)) stop("No CSV or Parquet files found in ", dir)

  purrr::map(files, ~ dplyr::mutate(read_record(.x), record = basename(.x)))
}
```

## Windowing and MMD

The MMD estimator here deliberately mirrors the Python implementation in
`diffusion_103.py` (biased V-statistic, Gaussian kernel, bandwidth `gamma = 1 / L`)
so that the numbers in the notebook and the numbers printed during training are on
the same scale.

```{r fn-mmd}
make_windows <- function(x, win, n_max = Inf, stride = NULL) {
  n <- length(x)
  if (n < win) return(matrix(numeric(0), nrow = 0, ncol = win))
  if (is.null(stride)) {
    stride <- max(1L, floor((n - win) / max(1L, min(n_max, n))))
  }
  starts <- seq(1L, n - win + 1L, by = stride)
  if (length(starts) > n_max) starts <- sample(starts, n_max)
  t(vapply(starts, function(s) x[s:(s + win - 1L)], numeric(win)))
}

sq_dists <- function(A, B) {
  aa <- rowSums(A^2)
  bb <- rowSums(B^2)
  d  <- outer(aa, bb, "+") - 2 * tcrossprod(A, B)
  pmax(d, 0)
}

# Bandwidth by the median heuristic: gamma = 1 / median pairwise squared
# distance, so that a typical pair sits where the kernel actually varies.
# This matters more than it sounds. The fixed gamma = 1/L used in
# diffusion_103.py is right for standardised windows but far too narrow for
# current in pA: on a record with a 7 pA standard deviation, gamma * median(d2)
# comes out around 100, every kernel evaluation underflows to zero, and the
# statistic collapses onto its finite-sample floor of 1/m + 1/n regardless of
# what the two samples actually contain.
median_gamma <- function(X, Y, n_sub = 200) {
  A <- X[sample(nrow(X), min(n_sub, nrow(X))), , drop = FALSE]
  B <- Y[sample(nrow(Y), min(n_sub, nrow(Y))), , drop = FALSE]
  m <- stats::median(sq_dists(A, B))
  if (!is.finite(m) || m <= 0) 1 / ncol(X) else 1 / m
}

# unbiased = TRUE drops the i == j diagonal (the U-statistic). The diagonal
# terms are all exactly 1 and contribute (1/m + 1/n) of pure bias to the
# V-statistic, which is what puts a floor under the biased version.
mmd_rbf <- function(X, Y, gamma = NULL, unbiased = TRUE) {
  m <- nrow(X); n <- nrow(Y)
  if (m < 2 || n < 2) return(NA_real_)
  if (is.null(gamma)) gamma <- median_gamma(X, Y)

  Kxx <- exp(-gamma * sq_dists(X, X))
  Kyy <- exp(-gamma * sq_dists(Y, Y))
  Kxy <- exp(-gamma * sq_dists(X, Y))

  if (unbiased) {
    (sum(Kxx) - sum(diag(Kxx))) / (m * (m - 1)) +
      (sum(Kyy) - sum(diag(Kyy))) / (n * (n - 1)) -
      2 * mean(Kxy)
  } else {
    mean(Kxx) + mean(Kyy) - 2 * mean(Kxy)
  }
}

# Permutation test: pool the two sets, reshuffle the labels, and see where the
# observed statistic falls in the null it generates. This answers the question
# the raw number cannot -- given that two draws from one stochastic process are
# never identical, is this difference larger than relabelling would produce?
mmd_permutation_p <- function(X, Y, n_perm = 200, gamma = NULL) {
  if (is.null(gamma)) gamma <- median_gamma(X, Y)
  obs <- mmd_rbf(X, Y, gamma)
  Z <- rbind(X, Y); m <- nrow(X)
  null <- replicate(n_perm, {
    i <- sample(nrow(Z))
    mmd_rbf(Z[i[seq_len(m)], , drop = FALSE],
            Z[i[-seq_len(m)], , drop = FALSE], gamma)
  })
  list(mmd = obs, p = (1 + sum(null >= obs)) / (n_perm + 1),
       null_mean = mean(null), null_sd = sd(null), gamma = gamma)
}
```

## Welch power spectral density

Written out rather than pulled from a package, so that the estimate is transparent
and the notebook has one fewer dependency.

```{r fn-psd}
welch_psd <- function(x, fs, nperseg = 1024, noverlap = NULL, detrend = TRUE) {
  x <- as.numeric(x[is.finite(x)])
  if (length(x) < nperseg) return(NULL)
  if (is.null(noverlap)) noverlap <- nperseg %/% 2

  step   <- nperseg - noverlap
  starts <- seq(1L, length(x) - nperseg + 1L, by = step)
  w      <- 0.5 - 0.5 * cos(2 * pi * (seq_len(nperseg) - 1) / (nperseg - 1))  # Hann
  U      <- sum(w^2)

  acc <- numeric(nperseg %/% 2 + 1)
  for (s in starts) {
    seg <- x[s:(s + nperseg - 1L)]
    if (detrend) seg <- seg - mean(seg)
    X <- stats::fft(seg * w)[seq_len(nperseg %/% 2 + 1)]
    acc <- acc + (Mod(X)^2) / (fs * U)
  }
  psd <- acc / length(starts)
  psd[-c(1, length(psd))] <- 2 * psd[-c(1, length(psd))]  # one-sided

  tibble::tibble(
    freq = seq(0, fs / 2, length.out = length(psd)),
    psd  = psd
  ) |> dplyr::filter(freq > 0)
}
```

## Dwell times and a simple idealiser

The half-amplitude threshold criterion is the classical approach and is what we use
to re-idealise the *generated* current in Section 6, so that the missed-event losses
reported there are the ones a conventional analysis would actually incur.

```{r fn-dwell}
dwell_times <- function(ideal, fs) {
  r <- rle(as.integer(round(ideal)))
  tibble::tibble(
    level    = r$values,
    n_sample = r$lengths,
    dwell_ms = 1000 * r$lengths / fs
  ) |>
    dplyr::filter(n_sample > 0) |>
    dplyr::mutate(state = ifelse(level == 0, "Closed", "Open"))
}

# Half-amplitude threshold idealiser. Levels are estimated from the amplitude
# histogram of the closed baseline and the mode of the open population.
# The original mad()-based amplitude estimate is fragile on a strongly bimodal
# record: if it fails it returns all-NA, and rle() then reports one "dwell" per
# sample, which is what silently destroyed the ground-truth check on the first
# run. Levels are now taken from the two largest peaks of the amplitude density,
# which is what an experimenter reads off the all-points histogram anyway, and
# the baseline is taken from the closed population rather than the global median.
idealise_half_amplitude <- function(current, unit_pA = NULL, baseline = NULL,
                                    n_levels = 1) {
  d <- density(current, n = 4096)

  # Local maxima of the density, strongest first.
  is_peak <- c(FALSE, diff(sign(diff(d$y))) == -2, FALSE)
  pk <- data.frame(x = d$x[is_peak], y = d$y[is_peak])
  pk <- pk[order(-pk$y), , drop = FALSE]

  if (is.null(baseline)) baseline <- if (nrow(pk)) pk$x[1] else median(current)

  if (is.null(unit_pA)) {
    # The open level is the strongest peak that is not the baseline peak.
    far <- pk[abs(pk$x - baseline) > 0.5 * sd(current), , drop = FALSE]
    unit_pA <- if (nrow(far)) far$x[1] - baseline else NA_real_
  }

  if (!is.finite(unit_pA) || abs(unit_pA) < .Machine$double.eps^0.5) {
    warning("Could not estimate the unitary amplitude from the density. ",
            "Pass unit_pA (and baseline) explicitly: read them off Figure 2E.")
    return(rep(NA_integer_, length(current)))
  }

  lvl <- (current - baseline) / unit_pA
  pmin(pmax(as.integer(round(lvl)), 0L), as.integer(n_levels))
}

block_mean <- function(x, factor) {
  factor <- max(1L, as.integer(factor))
  nb <- length(x) %/% factor
  if (nb < 1) return(mean(x))
  colMeans(matrix(x[seq_len(nb * factor)], nrow = factor))
}
```

## Subsampled KS test

```{r fn-ks}
ks_safe <- function(a, b, n_max = CFG$ks_subsample) {
  a <- a[is.finite(a)]; b <- b[is.finite(b)]
  if (length(a) > n_max) a <- sample(a, n_max)
  if (length(b) > n_max) b <- sample(b, n_max)
  suppressWarnings(stats::ks.test(a, b))
}
```

---

# 1. Load the data

```{r load}
real_list  <- read_folder(CFG$real_dir)
real       <- dplyr::bind_rows(real_list)
real_paths <- normalizePath(
  list.files(CFG$real_dir, pattern = "\\.(csv|parquet)$", full.names = TRUE,
             ignore.case = TRUE),
  winslash = "/", mustWork = FALSE)

gen_path    <- pick_dataset("ddpm",   "Choose the DDPM (DiffuChannel) record",
                            CFG$gen_file)
gan_path    <- pick_dataset("gan",    "Choose the GAN (DeepGANnel) record, or cancel",
                            CFG$gan_file)
single_path <- pick_dataset("single", "Choose the single-stage ablation record, or cancel",
                            CFG$gen_single_file)

# A generated file that also sits in the seed folder would be counted as real
# data as well as synthetic, which corrupts every "Real" statistic including
# the within-real MMD baseline that everything else is judged against.
warn_if_seed <- function(path, label) {
  if (is.na(path)) return(invisible(NULL))
  if (normalizePath(path, winslash = "/", mustWork = FALSE) %in% real_paths) {
    warning(label, " (", basename(path), ") is inside CFG$real_dir and is ",
            "therefore also being loaded as a seed record. Move it out.")
  }
}
warn_if_seed(gen_path, "The DDPM record")
warn_if_seed(gan_path, "The GAN record")
warn_if_seed(single_path, "The single-stage record")

if (is.na(gen_path)) {
  stop("No DDPM record chosen. The whole notebook is built around that file, ",
       "so there is nothing to analyse without it. Re-run and pick one, or set ",
       "CFG$gen_file to a path. (A GAN-only run is not supported: every figure ",
       "and table would be labelled 'Generated' while showing GAN output, which ",
       "is exactly the sort of mislabelling that ends up in a paper.)")
}

gen <- read_record(gen_path)
gen$record <- "generated"

has_single <- !is.na(single_path)
gen_single <- if (has_single) {
  x <- read_record(single_path); x$record <- "single-stage"; x
} else {
  message("No single-stage record. Ablation sections will be skipped.")
  NULL
}

has_gan <- !is.na(gan_path)
gan <- if (has_gan) {
  x <- read_record(gan_path); x$record <- "GAN"; x
} else {
  message("No GAN record. Comparison sections will be skipped.")
  NULL
}

FS_REAL <- fs_of(real_list[[1]])
FS_GEN  <- fs_of(gen)
FS_SING <- if (has_single) fs_of(gen_single) else NA_real_
FS_GAN  <- if (has_gan)    fs_of(gan)         else NA_real_

cat(sprintf("Sample rate  real %.1f Hz | generated %.1f Hz\n", FS_REAL, FS_GEN))
cat(sprintf("Real:      %d record(s), %s samples (%.1f s)\n",
            length(real_list), format(nrow(real), big.mark = ","),
            nrow(real) / FS_REAL))
cat(sprintf("DDPM:      %s samples (%.1f s) from %s\n",
            format(nrow(gen), big.mark = ","), nrow(gen) / FS_GEN,
            basename(gen_path)))
if (has_gan) {
  cat(sprintf("GAN:       %s samples (%.1f s) from %s\n",
              format(nrow(gan), big.mark = ","), nrow(gan) / FS_GAN,
              basename(gan_path)))
}
record_value("gan_included", if (has_gan) "yes" else "no",
             "DeepGANnel comparison arm present")

if (abs(FS_REAL - FS_GEN) / FS_REAL > 0.01) {
  warning(sprintf(paste("Real and generated sample rates differ (%.1f vs %.1f Hz).",
                        "Almost certainly the export Decimation factor was > 1.",
                        "Per-file rates are used throughout, but a generated record",
                        "coarser than the seed data cannot reproduce its high-frequency",
                        "noise and the PSD comparison will be truncated."),
                  FS_REAL, FS_GEN))
}

record_value("n_records",  length(real_list), "seed records")
record_value("n_samples",  format(nrow(real), big.mark = ","), "total seed samples")
record_value("n_seconds",  round(nrow(real) / FS_REAL, 1), "seconds of seed data")
record_value("sample_khz", FS_REAL / 1000)
record_value("filter_khz", CFG$filter_hz / 1000)
record_value("seq_len",    CFG$seq_len)
record_value("slow_factor", CFG$slow_factor)
record_value("context_len", CFG$context_len)
record_value("block_len",   CFG$block_len)
record_value("burst_tc",    CFG$burst_tc, "flicker threshold for the burst clock")
record_value("emb_dim",     CFG$emb_dim)
record_value("emb_drop",    CFG$emb_drop)
record_value("seq_len_ms",  round(1000 * CFG$seq_len / FS_REAL, 2),
             "training window duration")
record_value("long_len_s",  round(nrow(gen) / FS_GEN, 1))
record_value("n_blocks",    max(1, ceiling(nrow(gen) /
                              (CFG$block_len - CFG$context_len))))
```

## Table 1: seed data summary

```{r table1}
table1 <- purrr::map_dfr(real_list, function(d) {
  tibble::tibble(
    Record       = d$record[1],
    Samples      = nrow(d),
    `Duration (s)` = round(nrow(d) / FS_REAL, 2),
    `Max levels` = max(d$channels, na.rm = TRUE),
    `Open prob.` = round(mean(d$channels > 0), 4),
    `Mean (pA)`  = round(mean(d$current), 3),
    `SD (pA)`    = round(sd(d$current), 3)
  )
})

knitr::kable(table1, caption = "Table 1. Seed records used for training and evaluation.")
readr::write_csv(table1, file.path(CFG$out_dir, "table1_seed_data.csv"))
```

# 2. Figure 2: traces and amplitude distribution

## 2B–D: example traces

```{r fig2-traces, fig.height = 6}
n_show_real <- round(CFG$trace_secs * FS_REAL)
n_show_gen  <- round(CFG$trace_secs * FS_GEN)

trace_df <- dplyr::bind_rows(
  real |> dplyr::slice_head(n = n_show_real) |>
    dplyr::select(time, channels, current) |> dplyr::mutate(source = "Real"),
  gen  |> dplyr::slice_head(n = n_show_gen) |>
    dplyr::select(time, channels, current) |> dplyr::mutate(source = "Generated"),
  if (has_gan) {
    gan |> dplyr::slice_head(n = round(CFG$trace_secs * FS_GAN)) |>
      dplyr::select(time, channels, current) |> dplyr::mutate(source = "GAN")
  }
) |>
  dplyr::mutate(
    time_ms = 1000 * (time - min(time)),
    source  = factor(source, levels = c("Real", "Generated", "GAN"))
  ) |>
  tidyr::pivot_longer(c(channels, current), names_to = "signal", values_to = "value") |>
  dplyr::mutate(signal = factor(
    signal, levels = c("channels", "current"),
    labels = c("Idealisation", "Current (pA)")
  ))

p_traces <- ggplot(trace_df, aes(time_ms, value, colour = source)) +
  geom_line(linewidth = 0.25) +
  facet_grid(signal ~ source, scales = "free_y", switch = "y") +
  scale_colour_manual(values = pal_source, guide = "none") +
  labs(x = "Time (ms)", y = NULL) +
  theme_channel() +
  theme(strip.placement = "outside")

p_traces
save_fig(p_traces, "fig2_traces", height = 5)
```

## 2E: all-points amplitude histograms

```{r fig2-hist}
amp_df <- dplyr::bind_rows(
  tibble::tibble(current = real$current, source = "Real"),
  tibble::tibble(current = gen$current,  source = "Generated"),
  if (has_gan) tibble::tibble(current = gan$current, source = "GAN")
) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Generated", "GAN")))

p_amp <- ggplot(amp_df, aes(current, colour = source, fill = source)) +
  geom_density(alpha = 0.18, linewidth = 0.6, adjust = 0.6) +
  scale_colour_manual(values = pal_source) +
  scale_fill_manual(values = pal_source) +
  labs(x = "Current (pA)", y = "Density") +
  theme_channel()

p_amp
save_fig(p_amp, "fig2E_amplitude_histogram")

# Modes and dispersion
mode_of <- function(x) { d <- density(x, n = 4096); d$x[which.max(d$y)] }

closed_real <- real$current[real$channels == 0]
closed_gen  <- gen$current[gen$channels == 0]

record_value("modes_real", round(mode_of(real$current), 3), "modal current, real, pA")
record_value("modes_gen",  round(mode_of(gen$current), 3),  "modal current, generated, pA")
record_value("sd_real",    round(sd(closed_real), 4), "closed-state SD, real, pA")
record_value("sd_gen",     round(sd(closed_gen), 4),  "closed-state SD, generated, pA")

ks_amp <- ks_safe(real$current, gen$current)
record_value("ks_D", round(unname(ks_amp$statistic), 4),
             "KS statistic, all-points amplitude, real vs generated")
ks_amp

if (has_gan) {
  closed_gan <- gan$current[gan$channels == 0]
  ks_gan <- ks_safe(real$current, gan$current)
  record_value("ks_D_gan", round(unname(ks_gan$statistic), 4),
               "KS statistic, all-points amplitude, real vs GAN")
  record_value("modes_gan", round(mode_of(gan$current), 3), "modal current, GAN, pA")
  record_value("sd_gan", round(sd(closed_gan), 4), "closed-state SD, GAN, pA")
  print(ks_gan)
}
```

# 3. Maximum mean discrepancy

Three numbers, and the first is meaningless without the other two. The within-real value
is the floor from finite sampling; the white-noise value is what a naive noise model scores.

```{r mmd}
W_real <- make_windows(real$current, CFG$seq_len, CFG$n_windows_mmd)
W_gen  <- make_windows(gen$current,  CFG$seq_len, CFG$n_windows_mmd)

half   <- floor(nrow(real) / 2)
W_r1   <- make_windows(real$current[1:half],              CFG$seq_len, CFG$n_windows_mmd)
W_r2   <- make_windows(real$current[(half + 1):nrow(real)], CFG$seq_len, CFG$n_windows_mmd)

white  <- rnorm(nrow(real), mean(real$current), sd(real$current))
W_white <- make_windows(white, CFG$seq_len, CFG$n_windows_mmd)

# All three share one bandwidth, taken from the real-vs-generated comparison,
# so the numbers can be read against each other.
g0 <- median_gamma(W_real, W_gen)
mmd_main        <- mmd_rbf(W_real, W_gen,   gamma = g0)
mmd_within_real <- mmd_rbf(W_r1,   W_r2,    gamma = g0)
mmd_white       <- mmd_rbf(W_real, W_white, gamma = g0)

mmd_tbl <- tibble::tibble(
  Comparison = c("Real vs generated", "Real vs real (split half)",
                 "Real vs Gaussian white noise"),
  MMD        = c(mmd_main, mmd_within_real, mmd_white)
)

if (has_single) {
  W_gs <- make_windows(gen_single$current, CFG$seq_len, CFG$n_windows_mmd)
  mmd_tbl <- dplyr::bind_rows(
    mmd_tbl,
    tibble::tibble(Comparison = "Real vs single-stage", MMD = mmd_rbf(W_real, W_gs, gamma = g0))
  )
}

if (has_gan) {
  W_gan <- make_windows(gan$current, CFG$seq_len, CFG$n_windows_mmd)
  mmd_gan <- mmd_rbf(W_real, W_gan, gamma = g0)
  mmd_tbl <- dplyr::bind_rows(
    mmd_tbl,
    tibble::tibble(Comparison = "Real vs GAN (DeepGANnel)", MMD = mmd_gan)
  )
  record_value("mmd_gan", signif(mmd_gan, 4), "MMD, real vs DeepGANnel")
}

# One bandwidth for every comparison, or the numbers are not on one scale.
GAMMA <- median_gamma(W_real, W_gen)
cat(sprintf("Median-heuristic gamma = %.3e (fixed 1/L would be %.3e)\n",
            GAMMA, 1 / CFG$seq_len))

mmd_tbl$`MMD (gamma = 1/L, biased)` <- c(
  mmd_rbf(W_real, W_gen,  gamma = 1 / CFG$seq_len, unbiased = FALSE),
  mmd_rbf(W_r1,   W_r2,   gamma = 1 / CFG$seq_len, unbiased = FALSE),
  mmd_rbf(W_real, W_white, gamma = 1 / CFG$seq_len, unbiased = FALSE)
)[seq_len(nrow(mmd_tbl))]

knitr::kable(mmd_tbl, digits = 6,
             caption = paste("Maximum mean discrepancy. The first column uses",
                             "the median-heuristic bandwidth and the unbiased",
                             "estimator; the second reproduces the fixed 1/L",
                             "biased version printed during training."))

perm <- mmd_permutation_p(W_real, W_gen, n_perm = 200, gamma = GAMMA)
cat(sprintf(paste("\nPermutation test, real vs generated: MMD = %.5f,",
                  "null mean %.5f (SD %.5f), p = %.3f\n"),
            perm$mmd, perm$null_mean, perm$null_sd, perm$p))
record_value("mmd_perm_p", signif(perm$p, 3),
             "permutation p, real vs generated windows")
record_value("mmd_gamma", signif(GAMMA, 4), "median-heuristic kernel bandwidth")

if (has_gan) {
  perm_gan <- mmd_permutation_p(W_real, W_gan, n_perm = 200, gamma = GAMMA)
  cat(sprintf("Permutation test, real vs GAN:       MMD = %.5f, p = %.3f\n",
              perm_gan$mmd, perm_gan$p))
  record_value("mmd_perm_p_gan", signif(perm_gan$p, 3),
               "permutation p, real vs GAN windows")
}

record_value("mmd_main",        signif(mmd_main, 4))
record_value("mmd_within_real", signif(mmd_within_real, 4))
record_value("mmd_white",       signif(mmd_white, 4))
```

# 4. Figure 3: noise structure

## 3A: power spectral density of the closed state

Computed on closed-state segments only, so gating does not contaminate the estimate.
This is the plot that tells you whether the model learned the recording noise or just
its variance.

```{r fig3-psd}
longest_closed_run <- function(d, min_len) {
  r <- rle(d$channels == 0)
  ends <- cumsum(r$lengths)
  starts <- ends - r$lengths + 1
  keep <- which(r$values & r$lengths >= min_len)
  if (!length(keep)) return(numeric(0))
  unlist(lapply(keep, function(i) d$current[starts[i]:ends[i]]))
}

closed_real_seg <- longest_closed_run(real, CFG$psd_seg)
closed_gen_seg  <- longest_closed_run(gen,  CFG$psd_seg)

psd_labelled <- function(x, label, fs) {
  p <- welch_psd(x, fs, CFG$psd_seg)
  if (is.null(p)) {
    warning("Not enough closed-state samples for a PSD estimate: ", label,
            ". Lower CFG$psd_seg or use a record with longer closed periods.")
    return(NULL)
  }
  dplyr::mutate(p, source = label)
}

psd_df <- dplyr::bind_rows(
  psd_labelled(closed_real_seg, "Real",      FS_REAL),
  psd_labelled(closed_gen_seg,  "Generated", FS_GEN)
)

stopifnot(nrow(psd_df) > 0)

if (has_single) {
  cs <- longest_closed_run(gen_single, CFG$psd_seg)
  if (length(cs) >= CFG$psd_seg) {
    psd_df <- dplyr::bind_rows(psd_df, psd_labelled(cs, "Single-stage", FS_SING))
  }
}

closed_gan_seg <- if (has_gan) longest_closed_run(gan, CFG$psd_seg) else numeric(0)
if (length(closed_gan_seg) >= CFG$psd_seg) {
  psd_df <- dplyr::bind_rows(psd_df, psd_labelled(closed_gan_seg, "GAN", FS_GAN))
}

p_psd <- ggplot(psd_df, aes(freq, psd, colour = source)) +
  geom_line(linewidth = 0.5) +
  geom_vline(xintercept = CFG$filter_hz, linetype = "dashed",
             colour = "grey50", linewidth = 0.3) +
  annotate("text", x = CFG$filter_hz, y = Inf, label = "filter corner",
           hjust = -0.1, vjust = 1.6, size = 2.8, colour = "grey40") +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_colour_manual(values = pal_source) +
  labs(x = "Frequency (Hz)", y = expression(PSD~(pA^2/Hz))) +
  theme_channel()

p_psd
save_fig(p_psd, "fig3A_psd")

# Quantify agreement as the RMS difference in log10 PSD below the filter corner.
psd_cmp <- psd_df |>
  dplyr::filter(freq < CFG$filter_hz, source %in% c("Real", "Generated")) |>
  tidyr::pivot_wider(names_from = source, values_from = psd) |>
  tidyr::drop_na()

psd_rmse <- sqrt(mean((log10(psd_cmp$Real) - log10(psd_cmp$Generated))^2))

# Same statistic for the GAN, so the two methods are scored identically.
psd_rmse_gan <- NA_real_
if (has_gan && "GAN" %in% psd_df$source) {
  pc <- psd_df |>
    dplyr::filter(freq < CFG$filter_hz, source %in% c("Real", "GAN")) |>
    tidyr::pivot_wider(names_from = source, values_from = psd) |>
    tidyr::drop_na()
  if (nrow(pc)) {
    psd_rmse_gan <- sqrt(mean((log10(pc$Real) - log10(pc$GAN))^2))
    record_value("psd_log_rmse_gan", round(psd_rmse_gan, 3),
                 "RMS difference in log10 PSD below filter corner, GAN")
  }
}
record_value("psd_log_rmse", round(psd_rmse, 3),
             "RMS difference in log10 PSD below filter corner")
record_value("psd_description",
  sprintf(paste("Real and generated spectra agree to an RMS difference of %.2f",
                "log10 units below the %g Hz filter corner."),
          psd_rmse, CFG$filter_hz),
  "REPLACE with your own reading of Figure 3A")
```

## 3B: autocorrelation of the baseline residual

```{r fig3-acf}
acf_of <- function(x, lag_max, fs) {
  a <- stats::acf(x - mean(x), lag.max = lag_max, plot = FALSE, demean = TRUE)
  tibble::tibble(lag_ms = 1000 * a$lag[, 1, 1] / fs, acf = a$acf[, 1, 1])
}

acf_df <- dplyr::bind_rows(
  dplyr::mutate(acf_of(closed_real_seg, CFG$acf_lag_max, FS_REAL), source = "Real"),
  dplyr::mutate(acf_of(closed_gen_seg,  CFG$acf_lag_max, FS_GEN),  source = "Generated"),
  if (length(closed_gan_seg) > CFG$acf_lag_max) {
    dplyr::mutate(acf_of(closed_gan_seg, CFG$acf_lag_max, FS_GAN), source = "GAN")
  }
)

p_acf <- ggplot(acf_df, aes(lag_ms, acf, colour = source)) +
  geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.3) +
  geom_line(linewidth = 0.5) +
  scale_colour_manual(values = pal_source) +
  labs(x = "Lag (ms)", y = "Autocorrelation") +
  theme_channel()

p_acf
save_fig(p_acf, "fig3B_autocorrelation")

# Compare on lag index rather than lag time, so a rate mismatch does not
# silently drop every row when the two lag grids fail to line up.
acf_wide <- acf_df |>
  dplyr::group_by(source) |>
  dplyr::mutate(lag_n = dplyr::row_number()) |>
  dplyr::ungroup() |>
  dplyr::select(lag_n, source, acf) |>
  tidyr::pivot_wider(names_from = source, values_from = acf) |>
  tidyr::drop_na()
acf_rmse <- sqrt(mean((acf_wide$Real - acf_wide$Generated)^2))

acf_rmse_gan <- NA_real_
if (has_gan && "GAN" %in% acf_df$source) {
  aw <- acf_df |>
    dplyr::group_by(source) |>
    dplyr::mutate(lag_n = dplyr::row_number()) |>
    dplyr::ungroup() |>
    dplyr::filter(source %in% c("Real", "GAN")) |>
    dplyr::select(lag_n, source, acf) |>
    tidyr::pivot_wider(names_from = source, values_from = acf) |>
    tidyr::drop_na()
  if (nrow(aw)) {
    acf_rmse_gan <- sqrt(mean((aw$Real - aw$GAN)^2))
    record_value("acf_rmse_gan", round(acf_rmse_gan, 4),
                 "RMS difference in ACF, real vs GAN")
  }
}

record_value("acf_rmse", round(acf_rmse, 4), "RMS difference in ACF")
record_value("acf_description",
  sprintf("ACF curves differ by an RMS of %.3f over the first %.1f ms of lag.",
          acf_rmse, 1000 * CFG$acf_lag_max / FS_REAL),
  "REPLACE with your own reading of Figure 3B")

p_noise <- p_psd + p_acf + patchwork::plot_annotation(tag_levels = "A")
save_fig(p_noise, "fig3_noise", width = 10, height = 4)
```

# 5. Figure 4: the slow envelope

```{r fig4-envelope}
env_real <- block_mean(real$current, CFG$slow_factor)
env_gen  <- block_mean(gen$current,  CFG$slow_factor)

env_len <- min(length(env_real), length(env_gen))
env_df  <- dplyr::bind_rows(
  tibble::tibble(idx = seq_len(env_len), value = env_real[seq_len(env_len)],
                 source = "Real"),
  tibble::tibble(idx = seq_len(env_len), value = env_gen[seq_len(env_len)],
                 source = "Generated")
)

if (has_single) {
  env_single <- block_mean(gen_single$current, CFG$slow_factor)
  el <- min(env_len, length(env_single))
  env_df <- dplyr::bind_rows(
    env_df,
    tibble::tibble(idx = seq_len(el), value = env_single[seq_len(el)],
                   source = "Single-stage")
  )
}

if (has_gan) {
  env_gan <- block_mean(gan$current, CFG$slow_factor)
  eg <- min(env_len, length(env_gan))
  env_df <- dplyr::bind_rows(
    env_df,
    tibble::tibble(idx = seq_len(eg), value = env_gan[seq_len(eg)], source = "GAN")
  )
}

env_fs <- c("Real" = FS_REAL, "Generated" = FS_GEN,
            "Single-stage" = FS_SING, "GAN" = FS_GAN)
env_df <- dplyr::mutate(env_df,
  time_s = idx * CFG$slow_factor / unname(env_fs[source]))

p_env <- ggplot(env_df, aes(time_s, value, colour = source)) +
  geom_line(linewidth = 0.4) +
  facet_wrap(~ source, ncol = 1, scales = "free_y") +
  scale_colour_manual(values = pal_source, guide = "none") +
  labs(x = "Time (s)", y = "Envelope current (pA)") +
  theme_channel()

p_env
save_fig(p_env, "fig4A_envelope", height = 5)

p_env_dist <- ggplot(env_df, aes(value, colour = source, fill = source)) +
  geom_density(alpha = 0.18, linewidth = 0.5) +
  scale_colour_manual(values = pal_source) +
  scale_fill_manual(values = pal_source) +
  labs(x = "Envelope current (pA)", y = "Density") +
  theme_channel()

save_fig(p_env_dist, "fig4C_envelope_distribution")

record_value("env_sd_real",     round(sd(env_real), 4))
record_value("env_sd_cascade",  round(sd(env_gen), 4))

We_r <- make_windows(env_real, min(64, length(env_real)), CFG$n_windows_mmd)
We_g <- make_windows(env_gen,  min(64, length(env_gen)),  CFG$n_windows_mmd)
record_value("mmd_env_cascade", signif(mmd_rbf(We_r, We_g), 4))

if (has_gan) {
  record_value("env_sd_gan", round(sd(env_gan), 4))
  We_gan <- make_windows(env_gan, min(64, length(env_gan)), CFG$n_windows_mmd)
  record_value("mmd_env_gan", signif(mmd_rbf(We_r, We_gan), 4))
}

if (has_single) {
  record_value("env_sd_single", round(sd(env_single), 4))
  We_s <- make_windows(env_single, min(64, length(env_single)), CFG$n_windows_mmd)
  record_value("mmd_env_single", signif(mmd_rbf(We_r, We_s), 4))
} else {
  record_value("env_sd_single",  "NOT RUN", "no single-stage ablation available")
  record_value("mmd_env_single", "NOT RUN", "no single-stage ablation available")
}
```

# 6. Figure 5: block joins

If the context-carry inpainting works, the distribution of sample-to-sample differences
in the neighbourhood of a block boundary should be indistinguishable from that within
blocks. If it does not work, you will see it here before a reviewer does.

```{r fig5-seams}
step <- CFG$block_len - CFG$context_len
boundaries <- seq(step, nrow(gen) - 1, by = step)
boundaries <- boundaries[boundaries > CFG$context_len &
                         boundaries < (nrow(gen) - CFG$context_len)]

diffs <- diff(gen$current)

if (length(boundaries)) {
  near_idx <- unique(unlist(lapply(boundaries, function(b)
    seq(b - CFG$context_len %/% 2, b + CFG$context_len %/% 2))))
  near_idx <- near_idx[near_idx >= 1 & near_idx <= length(diffs)]
  far_idx  <- setdiff(seq_along(diffs), near_idx)

  ks_seam <- ks_safe(diffs[near_idx], diffs[far_idx])
  record_value("ks_seam", round(unname(ks_seam$statistic), 4),
               "KS statistic, sample differences near vs away from block joins")
  print(ks_seam)

  seam_df <- dplyr::bind_rows(
    tibble::tibble(d = diffs[near_idx], where = "Near block join"),
    tibble::tibble(d = diffs[sample(far_idx, min(length(far_idx),
                                                 length(near_idx)))],
                   where = "Within block")
  )

  p_seam <- ggplot(seam_df, aes(d, colour = where)) +
    geom_density(linewidth = 0.6, adjust = 0.8) +
    scale_colour_manual(values = c("Near block join" = "#c0392b",
                                   "Within block"    = "#1b1b1b")) +
    labs(x = "Sample-to-sample difference (pA)", y = "Density") +
    theme_channel()
  print(p_seam)
  save_fig(p_seam, "fig5B_seam_test")
} else {
  message("Generated record spans a single block; no joins to test.")
  record_value("ks_seam", "N/A", "single block, no joins")
}

# The long trace itself, with boundaries marked.
decim <- max(1L, nrow(gen) %/% 20000L)
long_df <- gen[seq(1, nrow(gen), by = decim), ]

p_long <- ggplot(long_df, aes(time, current)) +
  geom_line(linewidth = 0.15, colour = pal_source[["Generated"]]) +
  { if (length(boundaries))
      geom_vline(xintercept = gen$time[boundaries], linetype = "dashed",
                 colour = "grey55", linewidth = 0.3) } +
  labs(x = "Time (s)", y = "Current (pA)") +
  theme_channel()

p_long
save_fig(p_long, "fig5A_long_record", height = 3)
```

# 7. Figure 6: dwell times and the ground-truth check

Two comparisons. First, dwell times from the *simulated conditioning idealisation* --
this is the exact ground truth. Second, dwell times recovered by idealising the
*generated current trace* with a conventional half-amplitude criterion, as though it
were an experimental record. The gap between them is the missed-event loss, and being
able to measure it at all is the point of the whole conditional design.

```{r fig6-dwell}
dw_truth <- dwell_times(gen$channels, FS_GEN) |>
  dplyr::mutate(origin = "Ground truth (simulated)")

n_lev <- max(gen$channels, na.rm = TRUE)
ideal_recovered <- idealise_half_amplitude(gen$current, n_levels = max(1, n_lev))

if (all(is.na(ideal_recovered))) {
  stop("The half-amplitude idealiser failed on the generated trace. ",
       "Read the baseline and open level off Figure 2E and pass them in:\n",
       "  idealise_half_amplitude(gen$current, unit_pA = <open - baseline>, ",
       "baseline = <closed level>, n_levels = max(1, n_lev))")
}
if (length(rle(ideal_recovered)$lengths) > 0.5 * length(ideal_recovered)) {
  warning("The recovered idealisation switches on almost every sample, which ",
          "means the level estimate is wrong rather than that the record is ",
          "fast. Check unit_pA before believing anything in Figure 6C-D.")
}

dw_recov <- dwell_times(ideal_recovered, FS_GEN) |>
  dplyr::mutate(origin = "Recovered from generated current")

dw_real <- dwell_times(real$channels, FS_REAL) |>
  dplyr::mutate(origin = "Real (expert idealised)")

# The GAN emitted its Channels column itself, so those labels are a network
# output rather than a known truth. They are plotted for comparison but they
# are not a ground truth and must not be described as one.
dw_gan <- if (has_gan) {
  dwell_times(gan$channels, FS_GAN) |>
    dplyr::mutate(origin = "GAN (network's own labels)")
} else NULL

dw_all <- dplyr::bind_rows(dw_truth, dw_recov, dw_real, dw_gan) |>
  dplyr::filter(dwell_ms > 0)

# Sigworth-Sine display: log abscissa, square-root ordinate.
p_dwell <- ggplot(dw_all, aes(dwell_ms, colour = origin)) +
  geom_freqpoly(aes(y = after_stat(sqrt(count))), bins = 40, linewidth = 0.55) +
  scale_x_log10(labels = scales::label_number()) +
  facet_wrap(~ state, scales = "free_y") +
  scale_colour_manual(values = c(
    "Ground truth (simulated)"         = "#1b1b1b",
    "Recovered from generated current" = "#c0392b",
    "Real (expert idealised)"          = "#2980b9",
    "GAN (network's own labels)"       = "#8e44ad")) +
  labs(x = "Dwell time (ms)", y = expression(sqrt(count))) +
  theme_channel() +
  theme(legend.position = "top", legend.direction = "vertical")

p_dwell
save_fig(p_dwell, "fig6_dwell_times", width = 8, height = 4.6)

dwell_summary <- dw_all |>
  dplyr::group_by(origin, state) |>
  dplyr::summarise(
    n         = dplyr::n(),
    mean_ms   = mean(dwell_ms),
    median_ms = median(dwell_ms),
    .groups   = "drop"
  )

knitr::kable(dwell_summary, digits = 3,
             caption = "Dwell-time summary by origin and state.")
readr::write_csv(dwell_summary, file.path(CFG$out_dir, "dwell_summary.csv"))

get_mean <- function(df, o, s) {
  v <- df |> dplyr::filter(origin == o, state == s) |> dplyr::pull(mean_ms)
  if (length(v)) round(v[1], 3) else NA_real_
}

record_value("tau_open_true",       get_mean(dwell_summary, "Ground truth (simulated)", "Open"))
record_value("tau_closed_true",     get_mean(dwell_summary, "Ground truth (simulated)", "Closed"))
record_value("tau_open_recovered",  get_mean(dwell_summary, "Recovered from generated current", "Open"))
record_value("tau_closed_recovered", get_mean(dwell_summary, "Recovered from generated current", "Closed"))
```

Note that `mean_ms` is a mean dwell time, not a fitted exponential time constant. If you
want the time constants quoted in the manuscript to be fitted rather than empirical,
fit a mixture of exponentials to `dw_all` and substitute those. The empirical mean is a
defensible summary and is what the placeholders currently carry.

# 8. Figure 7: manifold projections

```{r fig7-manifold, fig.height = 4.5}
Wp_real <- make_windows(real$current, CFG$proj_win, CFG$n_windows_proj)
Wp_gen  <- make_windows(gen$current,  CFG$proj_win, CFG$n_windows_proj)

Wp_gan <- if (has_gan) make_windows(gan$current, CFG$proj_win, CFG$n_windows_proj)
          else matrix(numeric(0), nrow = 0, ncol = CFG$proj_win)

X   <- rbind(Wp_real, Wp_gen, Wp_gan)
lab <- rep(c("Real", "Generated", "GAN"),
           c(nrow(Wp_real), nrow(Wp_gen), nrow(Wp_gan)))

# Robust scaling before projection, as in the DeepGANnel pipeline.
X <- scale(X, center = TRUE, scale = TRUE)
X[!is.finite(X)] <- 0

um <- uwot::umap(X, n_neighbors = 15, min_dist = 0.1, n_components = 2)
ts <- Rtsne::Rtsne(X, dims = 2, perplexity = 30, check_duplicates = FALSE,
                   pca = TRUE, verbose = FALSE)$Y

proj_df <- dplyr::bind_rows(
  tibble::tibble(x = um[, 1], y = um[, 2], source = lab, method = "UMAP"),
  tibble::tibble(x = ts[, 1], y = ts[, 2], source = lab, method = "t-SNE")
) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Generated", "GAN")))

p_proj <- ggplot(proj_df, aes(x, y, colour = source)) +
  geom_point(size = 0.6, alpha = 0.55) +
  facet_wrap(~ method, scales = "free") +
  scale_colour_manual(values = pal_source) +
  labs(x = NULL, y = NULL) +
  theme_channel() +
  theme(axis.text = element_blank(), axis.ticks = element_blank())

p_proj
save_fig(p_proj, "fig7_manifold", width = 8, height = 4.2)

readr::write_csv(proj_df, file.path(CFG$out_dir, "manifold_coordinates.csv"))
```

Cluster separation testing with `ClusterSignificance` (as in DeepGANnel) is not run here
because the package is a Bioconductor dependency that many installations lack. If you
have it, the projection coordinates have been written to `manifold_coordinates.csv` and
can be fed to `ClusterSignificance::Pcp()` directly. Record the resulting p value against
the `cluster_p` placeholder.

```{r cluster-p}
record_value("cluster_p", "TO BE FILLED",
             "run ClusterSignificance on manifold_coordinates.csv")
record_value("manifold_description", "REPLACE with your reading of Figure 7",
             "describe the degree of overlap you actually see")
```

# 9. Table 3: collected similarity metrics

```{r table3}
# Head to head. Every metric is computed the same way for both methods, and
# the reference column is what the real record gives against itself where that
# is meaningful.
num <- function(x, d = 4) if (length(x) && is.finite(x)) round(x, d) else NA_real_

table3 <- tibble::tibble(
  Metric = c(
    "MMD vs real (median gamma, unbiased)",
    "MMD permutation p",
    "KS statistic, amplitude distribution",
    "Closed-state SD (pA)",
    "Log10 PSD RMS difference vs real",
    "ACF RMS difference vs real",
    "Envelope SD (pA)",
    "MMD, envelope vs real"
  ),
  `Reference (real)` = c(
    num(mmd_within_real, 6), NA_real_, NA_real_,
    num(sd(closed_real)), NA_real_, NA_real_,
    num(sd(env_real)), NA_real_
  ),
  Diffusion = c(
    num(mmd_main, 6), num(perm$p, 3), num(unname(ks_amp$statistic)),
    num(sd(closed_gen)), num(psd_rmse, 3), num(acf_rmse),
    num(sd(env_gen)), num(mmd_rbf(We_r, We_g), 6)
  )
)

if (has_gan) {
  table3$GAN <- c(
    num(mmd_gan, 6), num(perm_gan$p, 3), num(unname(ks_gan$statistic)),
    num(sd(closed_gan)), num(psd_rmse_gan, 3), num(acf_rmse_gan),
    num(sd(env_gan)), num(mmd_rbf(We_r, We_gan), 6)
  )
}

knitr::kable(
  table3,
  caption = paste("Table 3. Similarity metrics.",
                  if (has_gan)
                    paste("Diffusion and GAN columns are computed identically.",
                          "Lower is better throughout except the permutation p,",
                          "where a large value means the difference from real is",
                          "no bigger than relabelling produces.")
                  else "")
)

readr::write_csv(table3, file.path(CFG$out_dir, "table3_metrics.csv"))
```

## What this comparison does and does not settle

Both methods are scored identically on every row above, so the numbers are
comparable as measures of *how much the synthetic data resemble the real
record*. That is the only question they answer.

They do not settle the question the paper is actually about. The GAN emits its
raw trace and its Channels column together, so its labels are a network output;
the diffusion pipeline is handed its labels before it starts and never touches
them. A GAN record could match the real data better on every metric here and
still be the less useful training set, because a downstream idealiser trained
against it is being scored on another network's opinion rather than on truth.
The dwell-time figure keeps the two apart deliberately: the diffusion arm has a
"Ground truth (simulated)" trace, the GAN arm is labelled "network's own
labels", and there is no equivalent ground-truth line for it to be compared to.

If the GAN wins on similarity, say so plainly and make the ground-truth argument
on its own terms. If it loses, the similarity result and the ground-truth
argument point the same way, which is a stronger position but not a different
one.

# 10. Write the placeholder values

```{r write-results}
results_tbl <- dplyr::bind_rows(RESULTS$rows) |> dplyr::arrange(placeholder)

readr::write_csv(results_tbl, file.path(CFG$out_dir, "results_values.csv"))

knitr::kable(results_tbl,
             caption = "Values keyed to the [[placeholders]] in the manuscript draft.")

cat("\nWritten to", file.path(CFG$out_dir, "results_values.csv"), "\n")
```

## Tidy metrics for the meta-analysis

One row per metric per method, written to `runs_dir` so that
`DiffuChannel_meta.Rmd` can pool several seed records. The `direction` column
is what makes pooling safe: metrics are not all oriented the same way, and
averaging a "lower is better" distance with a "closer to the real value is
better" summary statistic would be meaningless.

```{r tidy-metrics}
RUN_ID <- if (is.na(CFG$run_id)) {
  paste0(basename(normalizePath(CFG$real_dir, mustWork = FALSE)),
         "_", tools::file_path_sans_ext(basename(gen_path)))
} else {
  as.character(CFG$run_id)
}
RUN_ID <- gsub("[^A-Za-z0-9._-]+", "-", RUN_ID)

dir.create(CFG$runs_dir, showWarnings = FALSE, recursive = TRUE)

# key -> what it measures, whose number it is, and which way is better.
#   lower       smaller is closer to the real record
#   higher      larger is better (permutation p: a big value means the
#               difference from real is no bigger than relabelling produces)
#   match_real  the target is the real value, not zero, so the comparable
#               quantity is the absolute difference from the reference
#   reference   the real record's own value, used by match_real rows
metric_map <- tibble::tribble(
  ~key,               ~metric,                ~method,        ~direction,
  "mmd_main",         "MMD vs real",          "Diffusion",    "lower",
  "mmd_gan",          "MMD vs real",          "GAN",          "lower",
  "mmd_within_real",  "MMD vs real",          "Real floor",   "reference",
  "mmd_white",        "MMD vs real",          "White noise",  "reference",
  "mmd_perm_p",       "MMD permutation p",    "Diffusion",    "higher",
  "mmd_perm_p_gan",   "MMD permutation p",    "GAN",          "higher",
  "ks_D",             "KS amplitude",         "Diffusion",    "lower",
  "ks_D_gan",         "KS amplitude",         "GAN",          "lower",
  "psd_log_rmse",     "log10 PSD RMS",        "Diffusion",    "lower",
  "psd_log_rmse_gan", "log10 PSD RMS",        "GAN",          "lower",
  "acf_rmse",         "ACF RMS",              "Diffusion",    "lower",
  "acf_rmse_gan",     "ACF RMS",              "GAN",          "lower",
  "mmd_env_cascade",  "MMD envelope",         "Diffusion",    "lower",
  "mmd_env_gan",      "MMD envelope",         "GAN",          "lower",
  "mmd_env_single",   "MMD envelope",         "Single-stage", "lower",
  "sd_gen",           "Closed-state SD (pA)", "Diffusion",    "match_real",
  "sd_gan",           "Closed-state SD (pA)", "GAN",          "match_real",
  "sd_real",          "Closed-state SD (pA)", "Real",         "reference",
  "env_sd_cascade",   "Envelope SD (pA)",     "Diffusion",    "match_real",
  "env_sd_gan",       "Envelope SD (pA)",     "GAN",          "match_real",
  "env_sd_single",    "Envelope SD (pA)",     "Single-stage", "match_real",
  "env_sd_real",      "Envelope SD (pA)",     "Real",         "reference",
  "ks_seam",          "KS block seam",        "Diffusion",    "lower"
)

vals <- dplyr::bind_rows(RESULTS$rows) |>
  dplyr::mutate(num = suppressWarnings(as.numeric(gsub(",", "", value))))

metrics_long <- metric_map |>
  dplyr::inner_join(dplyr::select(vals, key = placeholder, num), by = "key") |>
  dplyr::filter(is.finite(num)) |>
  dplyr::transmute(kind = "metric", run_id = RUN_ID, metric, method,
                   direction, value = num)

# Provenance, so a pooled result can be traced back to what produced it.
info <- tibble::tibble(
  kind = "info", run_id = RUN_ID,
  metric = c("seed_file", "ddpm_file", "gan_file", "n_seed_samples",
             "fs_real_hz", "fs_gen_hz", "fs_gan_hz", "seq_len",
             "slow_factor", "mmd_gamma", "analysed"),
  method = "run",
  direction = "info",
  value = NA_real_,
  text = c(paste(basename(unlist(lapply(real_list, function(d) d$record[1]))),
                 collapse = "; "),
           basename(gen_path),
           if (has_gan) basename(gan_path) else "none",
           format(nrow(real)), format(FS_REAL), format(FS_GEN),
           if (has_gan) format(FS_GAN) else NA_character_,
           format(CFG$seq_len), format(CFG$slow_factor),
           format(signif(GAMMA, 4)), format(Sys.time()))
)

run_tbl <- dplyr::bind_rows(dplyr::mutate(metrics_long, text = NA_character_), info)

run_path <- file.path(CFG$runs_dir, paste0(RUN_ID, "__metrics.csv"))
readr::write_csv(run_tbl, run_path)

cat("Run written to", run_path, "\n")
cat("Pool several of these with DiffuChannel_meta.Rmd.\n")
if (!has_gan) {
  cat("Note: no GAN arm in this run, so it contributes to the pooled\n",
      "     Diffusion summary but not to the paired comparison.\n", sep = "")
}

knitr::kable(dplyr::select(metrics_long, metric, method, direction, value),
             digits = 6, caption = "Tidy metrics exported for pooling.")
cat("Any row reading TO BE FILLED, NOT RUN or REPLACE needs your attention",
    "before the manuscript goes anywhere.\n")
```

# Session information

```{r session}
sessionInfo()
```
