---
title: 'DiffuChannel benchmark: diffusion vs DeepGANnel across all phenotypes'
author: "Barrett-Jolley lab"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    df_print: paged
  pdf_document:
    toc: true
    number_sections: false
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 8, fig.height = 5, dpi = 150)
```

# Purpose

One pass over every phenotype folder. For each one it reads the real seed record,
the DeepGANnel record and the diffusion (DDPM) record, computes the same metric
set on the two synthetic arms, and then pools the five paired comparisons.

Every arm is scored identically. Whether that flatters the diffusion model OR NOT is not
this paper's business.

```{r packages}
required <- c("dplyr", "tidyr", "ggplot2", "readr", "purrr", "arrow",
              "scales", "patchwork", "tibble","Rtsne", "uwot")
missing  <- required[!vapply(required, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing)) {
  stop("Install first: install.packages(c(\"",
       paste(missing, collapse = "\", \""), "\"))")
}

library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(purrr)
library(arrow)
library(patchwork)
library(uwot)
library(Rtsne)

set.seed(42)
```

# Configuration

```{r config}
CFG <- list(
  # ---- where things live --------------------------------------------------
  pheno_root = "./Phenotypes",   # holds Phenotype_A, Phenotype_B, ...
  batch_dir  = "./batch_out",    # holds Phenotype_A_DDPM.parquet, ...
  out_dir    = "./benchmark_out",

  raw_pattern = "raw",           # substring marking the seed csv (case-insensitive)
  gan_pattern = "gan",           # substring marking the DeepGANnel csv
  only        = NULL,            # e.g. c("Phenotype_A") to run a subset

  # ---- recording parameters ----------------------------------------------
  sample_hz  = 1e4,              # fallback only; a Time column always wins
  filter_hz  = 1e3,

  # ---- how to treat a generated file's own time column --------------------
  #   "seed"  every arm inherits the raw record's rate (treat as equivalent:
  #           the generated file's time column is cosmetic and wrong)
  #   "own"   an arm with a time column is taken at its word (treat as honest:
  #           it really was produced at that rate)
  rate_policy  = "seed",
  rate_compare = TRUE,           # also run the other policy where they differ

  # ---- must match training ------------------------------------------------
  seq_len     = 128,
  slow_factor = 64,

  # ---- analysis knobs -----------------------------------------------------
  n_windows_mmd = 400,           # O(n^2) in the kernel: keep modest
  n_perm        = 200,
  psd_seg       = "auto",        # "auto" = largest power of 2 every record supports
  psd_seg_min   = 256,           # below this, stop rather than report a coarse PSD
  acf_lag_max   = 200,
  # Which conductance level supplies the noise segments. "auto" picks, per
  # phenotype, whichever of closed or open every arm can support; a record that
  # barely closes has no baseline but plenty of open-level material.
  noise_level   = "auto",        # "auto", "closed" or "open"
  ks_subsample  = 2e5,
  trace_secs    = 0.3,

  # ---- optional extras ----------------------------------------------------
  do_manifold    = TRUE,        # UMAP / t-SNE, needs uwot + Rtsne, slow
  n_windows_proj = 400,

  proj_win       = 512,

  cache = FALSE                   # reuse per-phenotype results if present
  #Careful with caching - it can cause real confusion!!
)

dir.create(CFG$out_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(file.path(CFG$out_dir, "cache"), showWarnings = FALSE, recursive = TRUE)

pal <- c("Real" = "#1b1b1b", "Diffusion" = "#c0392b", "GAN" = "#8e44ad",
         "Real floor" = "#7f8c8d", "White noise" = "#bdc3c7")

theme_channel <- function(base_size = 10) {
  theme_classic(base_size = base_size) +
    theme(strip.background = element_blank(),
          strip.text = element_text(face = "bold", hjust = 0),
          legend.position = "top",
          legend.title = element_blank(),
          legend.key.size = unit(0.8, "lines"),
          axis.text = element_text(colour = "black"))
}

save_fig <- function(plot, name, width = 8, height = 5) {
  ggsave(file.path(CFG$out_dir, paste0(name, ".pdf")), plot,
         width = width, height = height)
  ggsave(file.path(CFG$out_dir, paste0(name, ".png")), plot,
         width = width, height = height, dpi = 300)
  invisible(plot)
}
```

# Reading records

The seed and GAN csv files were not written by this pipeline, so their headers
cannot be assumed. Named columns are used when they are recognisable; otherwise
the columns are inferred the same way `batch_run.py` infers them, which is the
only way the R and Python sides can agree on which column was which. The
resolved mapping is printed in the manifest below and is worth reading once.

```{r fn-read}
normalise_names <- function(nms) gsub("[_-]", " ", trimws(tolower(nms)))

# 10%-of-range quantisation, as in batch_run.py stage_seed()
quantise <- function(x) {
  span <- diff(range(x, na.rm = TRUE))
  if (span > 0) round(x / (0.1 * span)) else rep(0, length(x))
}

# A channels column may hold no more than this many discrete levels (MAX_LEVELS
# in batch_run.py). Used both to infer the column and to check a named one.
MAX_LEVELS <- 10

is_time_col <- function(x) {
  d <- diff(x)
  length(d) > 1 && all(d > 0) &&
    isTRUE(all.equal(d, rep(d[1], length(d)), tolerance = 1e-3))
}

infer_cols <- function(df) {
  # The column count decides whether a time column exists at all: three columns
  # means one of them is time, two means there is none. Which one it is still
  # has to be found, because the order varies.
  time_col <- NA_integer_
  if (ncol(df) >= 3) {
    hit <- which(vapply(df, is_time_col, logical(1)))
    if (!length(hit)) {
      stop(ncol(df), " columns but none of them increases in constant steps, ",
           "so the time column cannot be identified")
    }
    time_col <- hit[1]
  }
  rest <- setdiff(seq_along(df), if (is.na(time_col)) integer(0) else time_col)
  lev  <- lapply(rest, function(j) sort(unique(quantise(df[[j]]))))
  n    <- vapply(lev, length, integer(1))
  k    <- which.min(n)
  if (n[k] > MAX_LEVELS) {
    stop("no column with <= ", MAX_LEVELS, " discrete levels; level counts ",
         paste(n, collapse = ", "))
  }
  others <- setdiff(seq_along(rest), k)
  list(time   = time_col,
       chan   = rest[k],
       cur    = rest[others[which.max(n[others])]],
       levels = lev[[k]])
}

read_record <- function(path) {
  df <- if (grepl("\\.parquet$", path, ignore.case = TRUE)) {
    as.data.frame(arrow::read_parquet(path))
  } else {
    d <- suppressWarnings(as.data.frame(
      readr::read_csv(path, show_col_types = FALSE, name_repair = "minimal")))
    # A file with no header gets its first data row promoted to names; spot that
    # by the names parsing as numbers, and re-read.
    if (all(!is.na(suppressWarnings(as.numeric(names(d)))))) {
      d <- as.data.frame(readr::read_csv(path, col_names = FALSE,
                                         show_col_types = FALSE))
    }
    d
  }

  df <- df[, vapply(df, function(x) is.numeric(x) || is.logical(x) ||
                      !all(is.na(suppressWarnings(as.numeric(as.character(x))))),
                    logical(1)), drop = FALSE]
  df[] <- lapply(df, function(x) suppressWarnings(as.numeric(as.character(x))))
  df <- df[stats::complete.cases(df), , drop = FALSE]
  if (ncol(df) < 2 || nrow(df) < 100) stop("too few usable rows/columns")

  nn  <- normalise_names(names(df))
  idx <- function(target) { h <- which(nn %in% target); if (length(h)) h[1] else NA_integer_ }
  i_chan <- idx(c("channels", "channel", "ideal", "idealisation", "idealization"))
  i_cur  <- idx(c("noisy current", "current", "raw", "raw current", "i"))
  i_time <- idx(c("time", "t"))

  # Headers are either absent or correct, so a recognised pair is taken at face
  # value. Anything else is inferred. Column order is never assumed either way.
  if (!is.na(i_chan) && !is.na(i_cur) && i_chan != i_cur) {
    how  <- "named columns"
    chan <- as.integer(round(df[[i_chan]]))
  } else {
    inf <- infer_cols(df)
    i_chan <- inf$chan; i_cur <- inf$cur; i_time <- inf$time
    how  <- "inferred"
    chan <- match(quantise(df[[i_chan]]), inf$levels) - 1L
  }

  out <- tibble::tibble(channels = as.numeric(chan),
                        current  = as.numeric(df[[i_cur]]))
  out$time <- if (!is.na(i_time)) as.numeric(df[[i_time]]) else
    (seq_len(nrow(out)) - 1) / CFG$sample_hz
  out <- dplyr::filter(out, is.finite(channels), is.finite(current))

  attr(out, "map") <- sprintf(
    "%s: chan=col%d cur=col%d time=%s", how, i_chan, i_cur,
    if (is.na(i_time)) sprintf("none, assuming %g Hz", CFG$sample_hz)
    else paste0("col", i_time))
  dplyr::select(out, time, channels, current)
}

fs_of <- function(d, fallback = CFG$sample_hz) {
  if (nrow(d) < 3) return(fallback)
  dt <- stats::median(diff(d$time), na.rm = TRUE)
  if (!is.finite(dt) || dt <= 0) fallback else 1 / dt
}
```

# Metric functions

Carried over from `DiffuChannel_analysis.Rmd` without change.

```{r fn-metrics}
make_windows <- function(x, win, n_max = Inf, stride = NULL) {
  n <- length(x)
  if (n < win) return(matrix(numeric(0), nrow = 0, ncol = win))
  if (is.null(stride)) stride <- max(1L, floor((n - win) / max(1L, min(n_max, n))))
  starts <- seq(1L, n - win + 1L, by = stride)
  if (length(starts) > n_max) starts <- sample(starts, n_max)
  t(vapply(starts, function(s) x[s:(s + win - 1L)], numeric(win)))
}

sq_dists <- function(A, B) {
  pmax(outer(rowSums(A^2), rowSums(B^2), "+") - 2 * tcrossprod(A, B), 0)
}

# Median heuristic: gamma = 1 / median pairwise squared distance. The fixed 1/L
# used during training underflows to zero on current in pA and pins the
# statistic to its finite-sample floor.
median_gamma <- function(X, Y, n_sub = 200) {
  A <- X[sample(nrow(X), min(n_sub, nrow(X))), , drop = FALSE]
  B <- Y[sample(nrow(Y), min(n_sub, nrow(Y))), , drop = FALSE]
  m <- stats::median(sq_dists(A, B))
  if (!is.finite(m) || m <= 0) 1 / ncol(X) else 1 / m
}

mmd_rbf <- function(X, Y, gamma = NULL, unbiased = TRUE) {
  m <- nrow(X); n <- nrow(Y)
  if (m < 2 || n < 2) return(NA_real_)
  if (is.null(gamma)) gamma <- median_gamma(X, Y)
  Kxx <- exp(-gamma * sq_dists(X, X))
  Kyy <- exp(-gamma * sq_dists(Y, Y))
  Kxy <- exp(-gamma * sq_dists(X, Y))
  if (unbiased) {
    (sum(Kxx) - sum(diag(Kxx))) / (m * (m - 1)) +
      (sum(Kyy) - sum(diag(Kyy))) / (n * (n - 1)) - 2 * mean(Kxy)
  } else {
    mean(Kxx) + mean(Kyy) - 2 * mean(Kxy)
  }
}

mmd_permutation_p <- function(X, Y, n_perm = CFG$n_perm, gamma = NULL) {
  if (is.null(gamma)) gamma <- median_gamma(X, Y)
  obs <- mmd_rbf(X, Y, gamma)
  Z <- rbind(X, Y); m <- nrow(X)
  null <- replicate(n_perm, {
    i <- sample(nrow(Z))
    mmd_rbf(Z[i[seq_len(m)], , drop = FALSE], Z[i[-seq_len(m)], , drop = FALSE], gamma)
  })
  s <- stats::sd(null)
  list(mmd = obs,
       p   = (1 + sum(null >= obs)) / (n_perm + 1),
       z   = if (is.finite(s) && s > 0) (obs - mean(null)) / s else NA_real_)
}

# Noise segments are taken at a single conductance level, so that no gating step
# falls inside a segment. Runs are returned separately and never concatenated:
# gluing them together puts a step discontinuity at every join, and a step is
# broadband, lifting the high-frequency end of the spectrum and dragging the ACF
# towards zero in proportion to how often the record gates. That bias falls
# hardest on exactly the busy records the comparison is meant to include.
at_level <- function(d, level) {
  if (identical(level, "closed")) d$channels == 0 else d$channels >= max(d$channels)
}

level_runs <- function(d, level, min_len = 1L) {
  r <- rle(at_level(d, level))
  ends <- cumsum(r$lengths); starts <- ends - r$lengths + 1
  keep <- which(r$values & r$lengths >= min_len)
  lapply(keep, function(i) d$current[starts[i]:ends[i]])
}

longest_at <- function(d, level) {
  r <- rle(at_level(d, level))
  if (!any(r$values)) 0L else max(r$lengths[r$values])
}

# Welch across a list of runs: every segment from every run long enough to hold
# one contributes to a single average periodogram.
welch_psd <- function(runs, fs, nperseg) {
  if (is.numeric(runs)) runs <- list(runs)
  runs <- runs[vapply(runs, length, integer(1)) >= nperseg]
  if (!length(runs)) stop("no closed run holds ", nperseg, " samples")

  w <- 0.5 - 0.5 * cos(2 * pi * (seq_len(nperseg) - 1) / (nperseg - 1))  # Hann
  U <- sum(w^2)
  acc <- numeric(nperseg %/% 2 + 1)
  nseg <- 0L
  for (x in runs) {
    x <- as.numeric(x[is.finite(x)])
    for (s in seq(1L, length(x) - nperseg + 1L, by = nperseg %/% 2)) {
      seg <- x[s:(s + nperseg - 1L)]
      seg <- seg - mean(seg)
      X <- stats::fft(seg * w)[seq_len(nperseg %/% 2 + 1)]
      acc <- acc + (Mod(X)^2) / (fs * U)
      nseg <- nseg + 1L
    }
  }
  psd <- acc / nseg
  psd[-c(1, length(psd))] <- 2 * psd[-c(1, length(psd))]
  tibble::tibble(freq = seq(0, fs / 2, length.out = length(psd)), psd = psd,
                 n_seg = nseg) |>
    dplyr::filter(freq > 0)
}

# Length-weighted mean of the per-run ACFs, for the same reason.
acf_of <- function(runs, lag_max, fs) {
  if (is.numeric(runs)) runs <- list(runs)
  runs <- runs[vapply(runs, length, integer(1)) > lag_max]
  if (!length(runs)) stop("no closed run longer than ", lag_max, " samples")
  w <- vapply(runs, length, integer(1))
  A <- vapply(runs, function(x)
    stats::acf(x - mean(x), lag.max = lag_max, plot = FALSE,
               demean = TRUE)$acf[, 1, 1], numeric(lag_max + 1))
  tibble::tibble(lag_n  = seq_len(lag_max + 1),
                 lag_ms = 1000 * (seq_len(lag_max + 1) - 1) / fs,
                 acf    = as.numeric(A %*% w / sum(w)))
}

dwell_times <- function(ideal, fs) {
  r <- rle(as.integer(round(ideal)))
  tibble::tibble(level = r$values, n_sample = r$lengths,
                 dwell_ms = 1000 * r$lengths / fs) |>
    dplyr::mutate(state = ifelse(level == 0, "Closed", "Open"))
}

block_mean <- function(x, factor) {
  factor <- max(1L, as.integer(factor))
  nb <- length(x) %/% factor
  if (nb < 1) return(mean(x))
  colMeans(matrix(x[seq_len(nb * factor)], nrow = factor))
}

ks_safe <- function(a, b, n_max = CFG$ks_subsample) {
  a <- a[is.finite(a)]; b <- b[is.finite(b)]
  if (length(a) > n_max) a <- sample(a, n_max)
  if (length(b) > n_max) b <- sample(b, n_max)
  suppressWarnings(stats::ks.test(a, b))
}

# Window choice for the example traces, ported from _best_window in
# batch_run.py so the R figures and the Python previews pick the same way:
# prefer a stretch whose Popen sits in [lo, hi], else the one nearest the middle
# of that band. Taking the first n samples instead lands on whatever the record
# happens to open with, which for a sparse phenotype is often no events at all.
best_window <- function(chan, n, lo = 0.4, hi = 0.8) {
  if (length(chan) <= n) return(1L)
  cs <- c(0, cumsum(as.numeric(chan)))
  k  <- (cs[(n + 1):length(cs)] - cs[1:(length(cs) - n)]) / n
  inside <- which(k >= lo & k <= hi)
  if (length(inside)) return(inside[ceiling(length(inside) / 2)])
  which.min(abs(k - 0.5 * (lo + hi)))
}

rms_diff <- function(a, b) sqrt(mean((a - b)^2))
```

# 1. Find the files

```{r discover}
find_one <- function(folder, pattern) {
  f <- list.files(
    folder,
    pattern = "\\.(csv|parquet)$",
    full.names = TRUE,
    ignore.case = TRUE
  )

  hit <- f[
    grepl(pattern, basename(f), ignore.case = TRUE)
  ]

  if (!length(hit)) return(NA_character_)

  if (length(hit) > 1) {
    warning(
      folder, ": ", length(hit),
      " files match '", pattern,
      "', using ", basename(hit[1])
    )
  }

  hit[1]
}

folders <- list.dirs(CFG$pheno_root, recursive = FALSE)
if (!is.null(CFG$only)) folders <- folders[basename(folders) %in% CFG$only]
if (!length(folders)) stop("No phenotype folders in ", CFG$pheno_root)

parquets <- list.files(CFG$batch_dir, pattern = "\\.parquet$", full.names = TRUE)

manifest <- purrr::map_dfr(folders, function(fd) {
  tag <- gsub(" ", "_", basename(fd))
  ddpm <- parquets[grepl(paste0("^", tag), basename(parquets))]
  if (!length(ddpm)) ddpm <- parquets[grepl(tag, basename(parquets), fixed = TRUE)]
  tibble::tibble(
    phenotype = tag,
    raw  = find_one(fd, CFG$raw_pattern),
    gan  = find_one(fd, CFG$gan_pattern),
    ddpm = if (length(ddpm)) ddpm[1] else NA_character_
  )
})

knitr::kable(
  manifest |> dplyr::mutate(dplyr::across(c(raw, gan, ddpm), basename)),
  caption = "Manifest. Check that each row's three files belong to the same phenotype."
)

bad <- manifest |> dplyr::filter(is.na(raw) | is.na(ddpm))
if (nrow(bad)) {
  warning("Dropping phenotypes with no seed or no DDPM file: ",
          paste(bad$phenotype, collapse = ", "))
  manifest <- dplyr::filter(manifest, !is.na(raw), !is.na(ddpm))
}
if (!nrow(manifest)) stop("Nothing left to analyse.")
```

# 2. Pre-flight

Every record is read and checked before a single metric is computed, so that a
record which cannot support the analysis stops the run here rather than becoming
an `NA` that propagates quietly into the pooled tables.

All records are held in memory for the rest of the notebook. That is the price
of reading each file exactly once; if it becomes a problem, run fewer phenotypes
at a time with `CFG$only`.

```{r preflight}
recs <- purrr::map(seq_len(nrow(manifest)), function(i) {
  row <- manifest[i, ]
  out <- list(Real = read_record(row$raw), Diffusion = read_record(row$ddpm))
  if (!is.na(row$gan)) out$GAN <- read_record(row$gan)
  out
})
names(recs) <- manifest$phenotype

inventory <- purrr::imap_dfr(recs, function(arms, ph) {
  purrr::imap_dfr(arms, function(d, a) {
    closed <- d$current[d$channels == 0]
    open   <- d$current[d$channels > 0]
    tibble::tibble(
      phenotype = ph, arm = a, n = nrow(d),
      has_time = !grepl("time=none", attr(d, "map")),
      fs = fs_of(d),
      longest_closed = longest_at(d, "closed"),
      longest_open   = longest_at(d, "open"),
      popen = mean(d$channels > 0), max_level = max(d$channels),
      baseline = median(closed),
      unit_amp = median(open) - median(closed),
      cols = attr(d, "map"))
  })
})

# A record that has been min-max scaled, or left in the wrong units, has a
# unitary amplitude unlike its seed's. Every amplitude metric then compares
# units rather than distributions, and the arm loses on bookkeeping.
scale_check <- inventory |>
  dplyr::group_by(phenotype) |>
  dplyr::mutate(rel = unit_amp / unit_amp[arm == "Real"]) |>
  dplyr::ungroup() |>
  dplyr::filter(arm != "Real", rel < 1 / 3 | rel > 3)
if (nrow(scale_check)) {
  stop("Unitary amplitude differs from the seed by more than 3x in: ",
       paste(sprintf("%s/%s (%.3g vs %.3g)", scale_check$phenotype,
                     scale_check$arm, scale_check$unit_amp,
                     scale_check$unit_amp / scale_check$rel), collapse = "; "),
       ". That is a scaling or units problem, not a modelling result: the ",
       "amplitude metrics would compare units rather than distributions. ",
       "Check whether a min-max scaled copy has been used.")
}

knitr::kable(inventory, digits = 4,
             caption = "Pre-flight inventory. Every record that will be analysed.")
readr::write_csv(inventory, file.path(CFG$out_dir, "preflight_inventory.csv"))
```

## Sample rate

Only the seed record's time column reflects a real acquisition.
`diffusion_app.py` writes its export time as `arange(n) * d * SAMPLE_DT` with
`SAMPLE_DT` fixed at 1e-4, so a DDPM record always claims 10 kHz whatever it was
trained on. A GAN record's time column is whatever DeepGANnel wrote, and when
that disagrees with the seed there are two readings, neither of which can be
settled from the file alone:

- **equivalent** (`rate_policy = "seed"`): the time column is cosmetic. The
  record is a sample-for-sample stand-in for the seed and inherits its rate.
- **honest** (`rate_policy = "own"`): the record really was produced at that
  rate, so it has half the bandwidth and its events last twice as long in
  wall-clock terms.

Relabelling a genuinely decimated record makes its dwell times and spectrum
wrong; taking a cosmetic label at face value does the same in the other
direction. So both are computed and compared below.

```{r rate}
seed_fs <- inventory |>
  dplyr::filter(arm == "Real") |>
  dplyr::mutate(fs = dplyr::if_else(has_time, fs, CFG$sample_hz))
seed_fs <- setNames(seed_fs$fs, seed_fs$phenotype)

rates_under <- function(policy) {
  inventory |>
    dplyr::transmute(
      phenotype, arm,
      fs = if (identical(policy, "own")) {
        dplyr::if_else(has_time, fs, unname(seed_fs[phenotype]))
      } else {
        unname(seed_fs[phenotype])
      })
}

rate_map <- function(policy) {
  r <- rates_under(policy)
  split(setNames(r$fs, r$arm), r$phenotype)
}

disputed <- rates_under("own") |>
  dplyr::rename(fs_own = fs) |>
  dplyr::left_join(dplyr::rename(rates_under("seed"), fs_seed = fs),
                   by = c("phenotype", "arm")) |>
  dplyr::filter(abs(fs_own - fs_seed) / fs_seed > 0.01)

knitr::kable(rates_under("seed") |>
               tidyr::pivot_wider(names_from = arm, values_from = fs),
             digits = 1, caption = "Rate under the equivalent policy (Hz).")

if (nrow(disputed)) {
  knitr::kable(disputed, digits = 1,
               caption = "Arms where the two policies disagree. These are the only ones whose rate-dependent metrics change between policies.")
} else {
  cat("No arm's own time column disagrees with its seed; the two policies coincide.\n")
}

FS_MAIN <- rate_map(CFG$rate_policy)
FS_ALT  <- rate_map(if (identical(CFG$rate_policy, "seed")) "own" else "seed")
```

The lag-1 autocorrelation of the single-level noise, reported after the analysis,
is the physical evidence: decimation decorrelates adjacent samples, so a record
genuinely at half the rate sits well below its own seed while unaffected
phenotypes match.

## Noise level and segment length for the spectral metrics

The noise metrics need contiguous samples at one conductance level. Which level
that can be depends on the record: a quiet channel offers long closed periods, a
record at Popen 0.99 offers long open ones and closures only a sample or two
wide. The level is therefore chosen per phenotype as whichever *every* arm of
that phenotype can support, and reported so it can be stated in the legend.

Reporting an open-level spectrum is not a fudge: the requirement is that no
gating transition falls inside a segment, which the open level satisfies exactly
as well. It does measure something slightly different — open-channel noise sits
on top of the instrumentation noise — but it is the same quantity in both arms
of the comparison, which is all the paired metric needs.

```{r level-and-seg}
cap_at <- inventory |>
  dplyr::group_by(phenotype) |>
  dplyr::summarise(closed = min(longest_closed), open = min(longest_open),
                   .groups = "drop")

levels_used <- cap_at |>
  dplyr::mutate(
    level = if (identical(CFG$noise_level, "auto")) {
      dplyr::if_else(open > closed, "open", "closed")
    } else {
      CFG$noise_level
    },
    cap = dplyr::if_else(level == "open", open, closed))

if (any(levels_used$cap == 0)) {
  z <- levels_used |> dplyr::filter(cap == 0)
  stop("No usable ", paste(unique(z$level), collapse = "/"), " segments at all in: ",
       paste(z$phenotype, collapse = ", "),
       ". Set CFG$noise_level to the other level, or exclude these phenotypes.")
}

knitr::kable(levels_used, digits = 0,
             caption = "Longest run every arm can supply at each level, and the level used. 'cap' is the binding constraint on segment length.")

cap <- min(levels_used$cap)
limiting <- levels_used |> dplyr::slice_min(cap, n = 1, with_ties = FALSE)

NPERSEG <- if (identical(CFG$psd_seg, "auto")) 2^floor(log2(cap)) else as.integer(CFG$psd_seg)

if (NPERSEG > cap) {
  stop("CFG$psd_seg = ", NPERSEG, " but ", limiting$phenotype,
       " supports only ", cap, " samples at its ", limiting$level,
       " level. Set CFG$psd_seg = \"auto\" or reduce it to at most ",
       2^floor(log2(cap)), ".")
}
if (NPERSEG < CFG$psd_seg_min) {
  stop(limiting$phenotype, " supports only ", cap, " contiguous samples at ",
       "either level (closed ", limiting$closed, ", open ", limiting$open,
       "), allowing ", NPERSEG, " samples per segment, below CFG$psd_seg_min = ",
       CFG$psd_seg_min, ". A spectrum that coarse is not worth reporting. ",
       "Exclude that phenotype from the spectral metrics with CFG$only, or ",
       "lower psd_seg_min deliberately and say so in the methods.")
}
if (NPERSEG <= CFG$acf_lag_max) {
  stop("Segment length ", NPERSEG, " is not longer than CFG$acf_lag_max = ",
       CFG$acf_lag_max, ", so the ACF would be estimated over runs too short ",
       "to support it. Lower acf_lag_max.")
}

LEVEL_OF <- setNames(levels_used$level, levels_used$phenotype)

cat(sprintf("Segment length %d samples, set by %s at its %s level (cap %d).\n",
            NPERSEG, limiting$phenotype, limiting$level, cap))
for (r in sort(unique(unlist(FS_MAIN)))) {
  cat(sprintf("  at %g Hz: %.1f ms, %.1f Hz resolution\n", r,
              1000 * NPERSEG / r, r / NPERSEG))
}
if (dplyr::n_distinct(levels_used$level) > 1) {
  cat("\nLevels differ between phenotypes, so the noise metrics are paired\n",
      "within a phenotype but not comparable in absolute terms across them.\n",
      "State the level per panel in the figure legend.\n", sep = "")
}
```
# 3. Per-phenotype analysis

Each phenotype is reduced to a long table of metrics, plus the curves needed for
the pooled figures. `direction` records which way is better, because pooling a
distance with a summary statistic that should *match* the real record would be
meaningless otherwise.

```{r analyse}
analyse_one <- function(row, arms_all, fs, level) {
  ph <- row$phenotype

  real    <- arms_all$Real
  arms    <- arms_all[setdiff(names(arms_all), "Real")]
  has_gan <- "GAN" %in% names(arms)

  # ---- windows and MMD ----------------------------------------------------
  W <- list(Real = make_windows(real$current, CFG$seq_len, CFG$n_windows_mmd))
  for (a in names(arms)) W[[a]] <- make_windows(arms[[a]]$current, CFG$seq_len,
                                                CFG$n_windows_mmd)
  half <- floor(nrow(real) / 2)
  W_r1 <- make_windows(real$current[1:half], CFG$seq_len, CFG$n_windows_mmd)
  W_r2 <- make_windows(real$current[(half + 1):nrow(real)], CFG$seq_len, CFG$n_windows_mmd)
  W_wn <- make_windows(rnorm(nrow(real), mean(real$current), sd(real$current)),
                       CFG$seq_len, CFG$n_windows_mmd)

  # One bandwidth for every comparison in this phenotype, or the numbers are
  # not on a single scale.
  gam <- median_gamma(W$Real, W$Diffusion)

  # ---- single-level noise -------------------------------------------------
  # Runs are kept separate and every record has already been shown to hold at
  # least one of NPERSEG samples, so there is no missing-data branch here.
  runs <- lapply(arms_all, level_runs, level = level, min_len = NPERSEG)

  psd <- purrr::imap_dfr(runs, function(x, a)
    dplyr::mutate(welch_psd(x, fs[[a]], NPERSEG), source = a))
  acf_tbl <- purrr::imap_dfr(runs, function(x, a)
    dplyr::mutate(acf_of(x, CFG$acf_lag_max, fs[[a]]), source = a))

  # Lag-1 autocorrelation of the single-level noise, per arm. A record genuinely
  # at a lower rate than the one assumed for it sits well below its own seed.
  lag1 <- acf_tbl |> dplyr::filter(lag_n == 2)
  lag1 <- setNames(lag1$acf, lag1$source)

  # Two arms at different rates have different frequency and lag grids, so both
  # comparisons interpolate onto a shared axis rather than pivoting on the grid
  # value. Under the equivalent policy the grids already coincide and the
  # interpolation is a no-op.
  curve <- function(tbl, a, xvar, yvar) {
    d <- tbl[tbl$source == a, ]
    list(x = d[[xvar]], y = d[[yvar]])
  }

  psd_rms <- function(a) {
    r <- curve(psd, "Real", "freq", "psd")
    g <- curve(psd, a, "freq", "psd")
    hi <- min(CFG$filter_hz, max(r$x), max(g$x))
    lo <- max(min(r$x), min(g$x))
    grid <- exp(seq(log(lo), log(hi), length.out = 512))
    rms_diff(approx(r$x, log10(r$y), grid)$y, approx(g$x, log10(g$y), grid)$y)
  }

  acf_rms <- function(a) {
    r <- curve(acf_tbl, "Real", "lag_ms", "acf")
    g <- curve(acf_tbl, a, "lag_ms", "acf")
    grid <- seq(0, min(max(r$x), max(g$x)), length.out = CFG$acf_lag_max + 1)
    rms_diff(approx(r$x, r$y, grid)$y, approx(g$x, g$y, grid)$y)
  }

  # ---- slow envelope ------------------------------------------------------
  env <- list(Real = block_mean(real$current, CFG$slow_factor))
  for (a in names(arms)) env[[a]] <- block_mean(arms[[a]]$current, CFG$slow_factor)
  We <- lapply(env, function(e) make_windows(e, min(64, length(e)), CFG$n_windows_mmd))

  # ---- dwell times --------------------------------------------------------
  dw <- purrr::imap_dfr(arms_all, function(d, a)
    dplyr::mutate(dwell_times(d$channels, fs[[a]]), source = a))
  mean_dwell <- function(a, s) {
    v <- dw |> dplyr::filter(source == a, state == s) |> dplyr::pull(dwell_ms)
    if (!length(v)) stop(ph, "/", a, ": no ", s, " dwells to average")
    mean(v)
  }

  level_pts <- function(d) d$current[at_level(d, level)]

  # ---- assemble -----------------------------------------------------------
  # One metric name whatever the level, so the pooled comparison keeps all
  # phenotypes; the level itself is recorded in the provenance table.
  sd_name <- "Noise SD (pA)"
  M <- function(metric, method, direction, value)
    tibble::tibble(phenotype = ph, metric = metric, method = method,
                   direction = direction, value = as.numeric(value))

  rows <- dplyr::bind_rows(
    M("MMD vs real",          "Real floor",  "reference", mmd_rbf(W_r1, W_r2, gamma = gam)),
    M("MMD vs real",          "White noise", "reference", mmd_rbf(W$Real, W_wn, gamma = gam)),
    M(sd_name,                "Real",        "reference", sd(level_pts(real))),
    M("Envelope SD (pA)",     "Real",        "reference", sd(env$Real)),
    M("Open probability",     "Real",        "reference", mean(real$channels > 0)),
    M("Mean open dwell (ms)", "Real",        "reference", mean_dwell("Real", "Open")),
    M("Mean closed dwell (ms)", "Real",      "reference", mean_dwell("Real", "Closed")),
    purrr::map_dfr(names(arms), function(a) {
      pm <- mmd_permutation_p(W$Real, W[[a]], gamma = gam)
      dplyr::bind_rows(
        M("MMD vs real",            a, "lower",  mmd_rbf(W$Real, W[[a]], gamma = gam)),
        M("MMD permutation p",      a, "higher", pm$p),
        M("MMD z-score",            a, "lower",  pm$z),
        M("KS amplitude",           a, "lower",
          unname(ks_safe(real$current, arms[[a]]$current)$statistic)),
        M("log10 PSD RMS",          a, "lower",  psd_rms(a)),
        M("ACF RMS",                a, "lower",  acf_rms(a)),
        M("MMD envelope",           a, "lower",  mmd_rbf(We$Real, We[[a]])),
        M(sd_name,                  a, "match_real", sd(level_pts(arms[[a]]))),
        M("Envelope SD (pA)",       a, "match_real", sd(env[[a]])),
        M("Open probability",       a, "match_real", mean(arms[[a]]$channels > 0)),
        M("Mean open dwell (ms)",   a, "match_real", mean_dwell(a, "Open")),
        M("Mean closed dwell (ms)", a, "match_real", mean_dwell(a, "Closed"))
      )
    })
  )

  # ---- curves and provenance ---------------------------------------------
  # time_ms is built from the sample index and the rate assigned to the arm, not
  # from the file's own time column, which for a DDPM export is a fixed 1e-4
  # step regardless of what it was trained on and makes panels differ in width.
  traces <- purrr::imap_dfr(arms_all, function(d, a) {
    n <- min(nrow(d), round(CFG$trace_secs * fs[[a]]))
    s <- best_window(d$channels, n)
    d[s:(s + n - 1), ] |>
      dplyr::mutate(source = a, phenotype = ph,
                    time_ms = 1000 * (dplyr::row_number() - 1) / fs[[a]],
                    win_start = s, win_popen = mean(channels > 0))
  })

  amp <- purrr::imap_dfr(arms_all, function(d, a)
    tibble::tibble(current = d$current, source = a, phenotype = ph))

  info <- tibble::tibble(
    phenotype = ph,
    seed_file = basename(row$raw),
    gan_file  = if (has_gan) basename(row$gan) else "none",
    ddpm_file = basename(row$ddpm),
    n_real = nrow(real),
    n_ddpm = nrow(arms$Diffusion),
    n_gan  = if (has_gan) nrow(arms$GAN) else NA_integer_,
    popen_real = mean(real$channels > 0),
    max_level_real = max(real$channels),
    fs_real = fs[["Real"]],
    fs_ddpm = fs[["Diffusion"]],
    fs_gan  = if (has_gan) fs[["GAN"]] else NA_real_,
    noise_level = level,
    psd_nperseg = NPERSEG,
    psd_seg_ms = 1000 * NPERSEG / fs[["Real"]],
    psd_n_seg_real = psd$n_seg[1],
    lag1_real = unname(lag1["Real"]),
    lag1_ddpm = unname(lag1["Diffusion"]),
    lag1_gan  = if (has_gan) unname(lag1["GAN"]) else NA_real_,
    gamma = gam
  )

  list(metrics = rows,
       psd = dplyr::mutate(psd, phenotype = ph),
       acf = dplyr::mutate(acf_tbl, phenotype = ph),
       dwell = dplyr::mutate(dw, phenotype = ph),
       env = purrr::imap_dfr(env, function(e, a)
         tibble::tibble(idx = seq_along(e), value = e, source = a, phenotype = ph)),
       traces = traces,
       amp = amp,
       info = info)
}

# The cache is keyed on the segment length, the rates in force and the policy
# label: results computed under different assumptions are not comparable with
# these and must not be silently reused.
run_one <- function(row, fs, policy) {
  key <- sprintf("%s_seg%d_%s_%s_%s", row$phenotype, NPERSEG,
                 LEVEL_OF[[row$phenotype]], policy,
                 paste(sprintf("%s%g", substr(names(fs), 1, 1), fs), collapse = ""))
  cache_file <- file.path(CFG$out_dir, "cache", paste0(key, ".rds"))
  if (CFG$cache && file.exists(cache_file)) {
    message("Reusing cached ", key)
    return(readRDS(cache_file))
  }
  message("Analysing ", row$phenotype, " (", policy, " rates)")
  res <- analyse_one(row, recs[[row$phenotype]], fs, LEVEL_OF[[row$phenotype]])
  if (CFG$cache) saveRDS(res, cache_file)
  res
}

results <- purrr::map(seq_len(nrow(manifest)), function(i) {
  row <- manifest[i, ]
  run_one(row, FS_MAIN[[row$phenotype]], CFG$rate_policy)
})
names(results) <- manifest$phenotype

pull_all <- function(what) purrr::map_dfr(results, what)

metrics <- pull_all("metrics")
info    <- pull_all("info")

# Nothing missing gets past this point. An NA here would become a phenotype
# silently dropped from a paired comparison, which is worse than a failed knit.
bad <- dplyr::filter(metrics, !is.finite(value))
if (nrow(bad)) {
  stop("Non-finite metrics, which must be understood rather than dropped:\n",
       paste(sprintf("  %s / %s / %s", bad$phenotype, bad$method, bad$metric),
             collapse = "\n"))
}

knitr::kable(
  info |> dplyr::select(phenotype, seed_file, ddpm_file, gan_file,
                        n_real, n_ddpm, n_gan, popen_real, max_level_real),
  digits = 4, caption = "Table 1. Records analysed."
)
knitr::kable(info |> dplyr::select(phenotype, noise_level, fs_real, fs_ddpm,
                                   fs_gan, psd_nperseg, psd_seg_ms,
                                   psd_n_seg_real, gamma),
             digits = 4,
             caption = "Rate applied, spectral segment length, segments contributing to the real spectrum, and kernel bandwidth.")

knitr::kable(info |> dplyr::select(phenotype, lag1_real, lag1_ddpm, lag1_gan),
             digits = 4,
             caption = "Lag-1 autocorrelation of the single-level noise. An arm well below its own real value is genuinely at a lower rate than the one assumed for it, not merely mislabelled.")
readr::write_csv(info, file.path(CFG$out_dir, "provenance.csv"))
readr::write_csv(metrics, file.path(CFG$out_dir, "metrics_long.csv"))
```

## Equivalent versus honest

Only rate-dependent metrics can move between the policies. The amplitude
metrics, MMD and open probability are computed per sample and are identical
either way, which is itself useful: it means the choice cannot be dodged by
appealing to the metrics that happen to be convenient.

```{r policy-compare, eval = CFG$rate_compare}
alt_phenos <- unique(disputed$phenotype)

if (!length(alt_phenos)) {
  cat("No arm's rate is disputed, so the two policies give identical results.\n")
} else {
  alt <- purrr::map(alt_phenos, function(ph) {
    row <- manifest[manifest$phenotype == ph, ]
    run_one(row, FS_ALT[[ph]], if (identical(CFG$rate_policy, "seed")) "own" else "seed")
  })
  names(alt) <- alt_phenos

  rate_sensitive <- c("log10 PSD RMS", "ACF RMS",
                      "Mean open dwell (ms)", "Mean closed dwell (ms)")

  cmp <- dplyr::bind_rows(
    metrics |> dplyr::filter(phenotype %in% alt_phenos) |>
      dplyr::mutate(policy = CFG$rate_policy),
    purrr::map_dfr(alt, "metrics") |>
      dplyr::mutate(policy = if (identical(CFG$rate_policy, "seed")) "own" else "seed")
  ) |>
    dplyr::filter(metric %in% rate_sensitive) |>
    dplyr::mutate(policy = dplyr::recode(policy, seed = "equivalent", own = "honest")) |>
    dplyr::select(phenotype, metric, method, policy, value) |>
    tidyr::pivot_wider(names_from = policy, values_from = value) |>
    dplyr::mutate(ratio = honest / equivalent)

  knitr::kable(cmp, digits = 4,
               caption = "Rate-dependent metrics under both readings. Everything else is unchanged by construction.")
  readr::write_csv(cmp, file.path(CFG$out_dir, "rate_policy_comparison.csv"))

  lag_cmp <- dplyr::bind_rows(
    info |> dplyr::filter(phenotype %in% alt_phenos) |>
      dplyr::mutate(policy = CFG$rate_policy),
    purrr::map_dfr(alt, "info") |>
      dplyr::mutate(policy = if (identical(CFG$rate_policy, "seed")) "own" else "seed")
  ) |>
    dplyr::select(phenotype, policy, lag1_real, lag1_ddpm, lag1_gan)
  knitr::kable(lag_cmp, digits = 4,
               caption = "Lag-1 autocorrelation, which does not depend on the rate label at all and is therefore the evidence.")
}
```

### How to read this

The lag-1 table is the part that decides it, because it is computed per sample
and is untouched by whichever label is applied. If the disputed arm's value sits
close to its own seed record, the samples are correlated like the seed's are and
the record is a sample-for-sample stand-in with a cosmetic time column:
**equivalent** is right, and the honest reading inflates its dwell times by the
rate ratio for no physical reason. If it sits well below, adjacent samples really
are less correlated, the record was produced at the lower rate, and **honest** is
right — but then the record has less bandwidth than the seed, which is a genuine
limitation of that arm and belongs in the results rather than being corrected
away.

The dwell ratios are the sanity check on whichever answer you accept: under the
honest reading a 5 kHz record's mean dwells double exactly, so if the honest
dwell times land implausibly far from the seed's while the equivalent ones sit
close, that is evidence for the equivalent reading and vice versa.

# 4. Scoring

Distances should be small, the permutation p should be large, and summary
statistics such as the noise SD should *match* the real record rather
than approach zero. The score column puts all three on a "smaller is better"
footing except the p, which is flagged separately.

```{r score}
refs <- metrics |>
  dplyr::filter(direction == "reference", method == "Real") |>
  dplyr::select(phenotype, metric, ref = value)

scored <- metrics |>
  dplyr::filter(direction %in% c("lower", "higher", "match_real")) |>
  dplyr::left_join(refs, by = c("phenotype", "metric")) |>
  dplyr::mutate(
    score  = dplyr::if_else(direction == "match_real", abs(value - ref), value),
    better = dplyr::if_else(direction == "higher", "higher", "lower")
  ) |>
  dplyr::filter(is.finite(score))

knitr::kable(
  scored |> dplyr::arrange(metric, method, phenotype) |>
    dplyr::select(metric, method, phenotype, value, ref, score),
  digits = 5, caption = "Table 2. Raw values and the comparable score derived from them."
)
readr::write_csv(scored, file.path(CFG$out_dir, "scores_long.csv"))
```

## Table 3: head to head, per phenotype

```{r table3}
table3 <- scored |>
  dplyr::select(metric, phenotype, method, value) |>
  tidyr::pivot_wider(names_from = method, values_from = value) |>
  dplyr::left_join(refs, by = c("phenotype", "metric")) |>
  dplyr::rename(`Reference (real)` = ref) |>
  dplyr::arrange(metric, phenotype)

knitr::kable(table3, digits = 5,
             caption = paste("Table 3. Similarity metrics, computed identically",
                             "for both methods. Lower is better throughout except",
                             "the permutation p, where a large value means the",
                             "difference from real is no bigger than relabelling",
                             "produces."))
readr::write_csv(table3, file.path(CFG$out_dir, "table3_metrics.csv"))
```

# 5. Pooled comparison

The design is paired: both methods see the same seed record, so the quantity of
interest is the within-phenotype difference. Between-phenotype variation in
noise, open probability and duration is far larger than any method effect and
would swamp it if the records were treated as independent samples.

```{r paired}
pairable <- scored |>
  dplyr::filter(method %in% c("Diffusion", "GAN")) |>
  dplyr::group_by(metric, phenotype) |>
  dplyr::filter(dplyr::n_distinct(method) == 2) |>
  dplyr::ungroup()

n_ph <- dplyr::n_distinct(pairable$phenotype)

p_slope <- ggplot(pairable, aes(factor(method, levels = c("Diffusion", "GAN")),
                                score, group = phenotype)) +
  geom_line(colour = "grey60", linewidth = 0.4) +
  geom_point(aes(colour = method), size = 2.2) +
  facet_wrap(~ metric, scales = "free_y") +
  scale_colour_manual(values = pal, guide = "none") +
  labs(x = NULL, y = "Score (lower is better, except the permutation p)") +
  theme_channel()
p_slope
save_fig(p_slope, "fig_paired_metrics", width = 9, height = 6.5)
```

Each line is one phenotype. Lines that all slope the same way are the reportable
result; lines that cross mean the ranking depends on which recording you happen
to analyse.

```{r pooled-summary}
wide <- pairable |>
  dplyr::select(metric, phenotype, better, method, score) |>
  tidyr::pivot_wider(names_from = method, values_from = score) |>
  dplyr::mutate(diff = Diffusion - GAN)

paired_stats <- wide |>
  dplyr::group_by(metric, better) |>
  dplyr::summarise(
    n              = dplyr::n(),
    n_ties         = sum(diff == 0),
    median_diff    = median(diff),
    mean_diff      = mean(diff),
    sd_diff        = sd(diff),
    n_favour_diff  = sum(if (dplyr::first(better) == "lower") diff < 0 else diff > 0),
    p_paired_t     = if (dplyr::n() > 2 && sd(diff) > 0)
                       stats::t.test(diff)$p.value else NA_real_,
    ci_lo          = if (dplyr::n() > 2 && sd(diff) > 0)
                       stats::t.test(diff)$conf.int[1] else NA_real_,
    ci_hi          = if (dplyr::n() > 2 && sd(diff) > 0)
                       stats::t.test(diff)$conf.int[2] else NA_real_,
    # The sign test must be taken in the same tail as `better`, and exact ties
    # are absence of evidence rather than evidence against: counting them as
    # unfavourable reported 0.0625 for a metric that was three-fifths tied.
    p_sign         = {
      fav <- if (dplyr::first(better) == "lower") sum(diff < 0) else sum(diff > 0)
      nz  <- sum(diff != 0)
      if (nz > 0) stats::binom.test(fav, nz)$p.value else NA_real_
    },
    .groups = "drop"
  ) |>
  dplyr::mutate(
    n_eff   = n - n_ties,
    verdict = dplyr::case_when(
      n_eff == 0             ~ "no separation: all differences exactly tied",
      n_favour_diff == n_eff ~ "consistent: Diffusion",
      n_favour_diff == 0     ~ "consistent: GAN",
      TRUE                   ~ "inconsistent across phenotypes"),
    verdict = dplyr::if_else(n_ties > 0 & n_eff > 0,
                             paste0(verdict, " (", n_ties, " tied)"), verdict)) |>
  dplyr::select(metric, better, n, n_ties, median_diff, mean_diff, sd_diff,
                n_favour_diff, p_paired_t, ci_lo, ci_hi, p_sign, verdict)

knitr::kable(paired_stats, digits = 5,
             caption = paste("Table 4. Pooled paired comparison, Diffusion minus",
                             "GAN within each phenotype. The t-test is on the",
                             "within-phenotype differences; the sign test excludes",
                             "exact ties, so with five records it cannot fall below",
                             "0.0625 and the unanimity count is the reportable",
                             "result."))
readr::write_csv(paired_stats, file.path(CFG$out_dir, "table4_paired_summary.csv"))
```

## What this many phenotypes will and will not support

```{r power}
min_p_sign <- min(1, 2 * 0.5^n_ph)
cat(sprintf("Phenotypes pooled: %d\n", n_ph))
cat(sprintf("Smallest two-sided p a sign test can return here: %.4f\n", min_p_sign))

if (min_p_sign > 0.05) {
  cat("\nThat is above 0.05, so no sign test on these data can reach significance,\n",
      "however large or however unanimous the effect. The same floor applies to the\n",
      "Wilcoxon signed-rank test at this n. The p_sign column above is therefore\n",
      "reported for completeness and should not be leaned on.\n",
      "\nWhat these data do support: the per-phenotype values, the direction, whether\n",
      "it is unanimous, and the paired t interval on the differences. Those are the\n",
      "defensible summaries at this sample size.\n", sep = "")
} else {
  cat("\nA sign test is at least capable of reaching p < 0.05 here.\n")
}

cat(sprintf("\nUnanimous metrics: %d of %d.\n",
            sum(paired_stats$verdict != "inconsistent across phenotypes"),
            nrow(paired_stats)))
```

The paired t-test above assumes the five differences are roughly normal, which
five points cannot demonstrate. It is quoted as an effect estimate with an
interval rather than as a hypothesis test, and the interval is wide by
construction. A metric that comes out inconsistent is a finding in itself: it
means the two methods are close enough that recording-specific detail decides
the ranking.

# 6. Publication figures

## Figure: example traces

```{r fig-traces, fig.height = 7}
traces <- pull_all("traces") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")))

# Popen of the window actually shown, so a panel cannot quietly misrepresent the
# record it came from.
win_lab <- traces |>
  dplyr::group_by(phenotype, source) |>
  dplyr::summarise(win_popen = dplyr::first(win_popen),
                   win_start = dplyr::first(win_start),
                   x = min(time_ms), y = max(current), .groups = "drop") |>
  dplyr::mutate(label = sprintf("Popen %.2f (from %d)", win_popen, win_start))

p_traces <- ggplot(traces, aes(time_ms, current, colour = source)) +
  geom_line(linewidth = 0.2) +
  geom_text(data = win_lab, aes(x, y, label = label), hjust = 0, vjust = 1,
            size = 2.4, colour = "grey30", inherit.aes = FALSE) +
  facet_grid(phenotype ~ source, scales = "free_y") +
  scale_colour_manual(values = pal, guide = "none") +
  labs(x = "Time (ms)", y = "Current (pA)") +
  theme_channel()
p_traces
save_fig(p_traces, "fig_traces", width = 8, height = 1.4 * n_ph + 1)
```

The window is chosen by the same rule the Python previews use, so a panel shows
gating rather than whichever stretch the file happens to start with. Where no
window falls inside the target band — Phenotype_B is too quiet, Phenotype_D too
busy — the nearest available is used, and its Popen is printed so the gap is
visible rather than implied.

## Figure: idealisations

```{r fig-ideal, fig.height = 7}
p_ideal <- ggplot(traces, aes(time_ms, channels, colour = source)) +
  geom_step(linewidth = 0.3) +
  facet_grid(phenotype ~ source) +
  scale_colour_manual(values = pal, guide = "none") +
  labs(x = "Time (ms)", y = "Open channels") +
  theme_channel()
p_ideal
save_fig(p_ideal, "fig_idealisations", width = 8, height = 1.2 * n_ph + 1)
```

## Figure: all-points amplitude histograms

```{r fig-amp, fig.height = 6}
p_amp <- pull_all("amp") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  ggplot(aes(current, colour = source, fill = source)) +
  geom_density(alpha = 0.15, linewidth = 0.5, adjust = 0.6) +
  facet_wrap(~ phenotype, scales = "free", ncol = 2) +
  scale_colour_manual(values = pal) + scale_fill_manual(values = pal) +
  labs(x = "Current (pA)", y = "Density") +
  theme_channel()
p_amp
save_fig(p_amp, "fig_amplitude", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
```

## Figure: single-level noise

```{r fig-noise, fig.height = 6}
# Panels are labelled with the level they were computed at, so the legend can
# state it and a phenotype that offers no baseline is not silently different.
panel_label <- setNames(sprintf("%s (%s level)", names(LEVEL_OF), LEVEL_OF),
                        names(LEVEL_OF))

p_psd <- pull_all("psd") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")),
                phenotype = panel_label[phenotype]) |>
  ggplot(aes(freq, psd, colour = source)) +
  geom_line(linewidth = 0.4) +
  geom_vline(xintercept = CFG$filter_hz, linetype = "dashed",
             colour = "grey60", linewidth = 0.3) +
  facet_wrap(~ phenotype, ncol = 2) +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_colour_manual(values = pal) +
  labs(x = "Frequency (Hz)", y = expression(PSD~(pA^2/Hz))) +
  theme_channel()

p_acf <- pull_all("acf") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")),
                phenotype = panel_label[phenotype]) |>
  ggplot(aes(lag_ms, acf, colour = source)) +
  geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.3) +
  geom_line(linewidth = 0.4) +
  facet_wrap(~ phenotype, ncol = 2) +
  scale_colour_manual(values = pal) +
  labs(x = "Lag (ms)", y = "Autocorrelation") +
  theme_channel()

p_psd
p_acf
save_fig(p_psd, "fig_psd", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
save_fig(p_acf, "fig_acf", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
```

## Figure: dwell times

Sigworth-Sine display: log abscissa, square-root ordinate. The diffusion arm's
`Channels` column is the conditioning idealisation it was handed, so it is a
known ground truth. The GAN emitted its own labels, so those are a network
output. They are plotted together for comparison and must not be described as
equivalent in the manuscript.

```{r fig-dwell, fig.height = 7}
p_dwell <- pull_all("dwell") |>
  dplyr::filter(dwell_ms > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  ggplot(aes(dwell_ms, colour = source)) +
  geom_area(aes(y = after_stat(sqrt(count)), fill = source),
            stat = "bin", bins = 40, alpha = 0.15,
            position = "identity", colour = NA) +
  geom_freqpoly(aes(y = after_stat(sqrt(count))), bins = 40, linewidth = 0.45) +
  facet_grid(phenotype ~ state, scales = "free_y") +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(x = "Dwell time (ms)", y = expression(sqrt(count))) +
  theme_channel()
p_dwell
save_fig(p_dwell, "fig_dwell", width = 8, height = 1.5 * n_ph + 1)
```

and bars
```{r dwellbars,  fig.height = 7}
d_dwell <- pull_all("dwell") |>
  dplyr::filter(dwell_ms > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")))

# Log-spaced breaks, so every bar is the same width on the log axis. Equal
# linear bins would be wide at short dwells and thin at long ones.
brk <- 10^seq(log10(min(d_dwell$dwell_ms)), log10(max(d_dwell$dwell_ms)),
              length.out = 41)

p_dwell <- ggplot(d_dwell, aes(dwell_ms, colour = source)) +
  geom_histogram(aes(y = after_stat(sqrt(count)), fill = source),
                 breaks = brk, alpha = 0.15, position = "identity", colour = NA) +
  geom_histogram(aes(y = after_stat(sqrt(count))), fill = NA,
                 breaks = brk, linewidth = 0.4, position = "identity") +
  facet_grid(phenotype ~ state, 
             scales = "free_y",
             labeller = labeller(
             phenotype = function(x) gsub("_", "\n", x, fixed = TRUE)
    )) +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(x = "Dwell time (ms)", y = expression(sqrt(count))) +
  theme_channel()+
  theme(
    strip.text.y = element_text(
      angle = 270,
      hjust = 0.5,
      vjust = 0.5
    )
  ) +
    guides(colour = guide_legend(override.aes = list(
    fill = scales::alpha(pal[c("Real", "Diffusion", "GAN")], 0.15),
    linewidth = 0.4)))
p_dwell

save_fig(p_dwell, "fig_dwell", width = 8, height = 1.5 * n_ph + 1)
```



## Bursts

```{r bursts, fig.height = 6}
# Two-component exponential mixture on the closed times by EM, then the
# Colquhoun-Sakmann equal-areas tcrit: the gap length at which as many brief
# closures are misread as between-burst as long ones are misread as within.
fit_tcrit <- function(closed_ms, n_iter = 200) {
  x <- closed_ms[is.finite(closed_ms) & closed_ms > 0]
  if (length(x) < 50) return(NA_real_)
  tau <- stats::quantile(x, c(0.25, 0.9), names = FALSE)
  if (!(tau[2] > tau[1])) return(NA_real_)
  a <- c(0.5, 0.5)
  for (i in seq_len(n_iter)) {
    d1 <- a[1] * stats::dexp(x, 1 / tau[1])
    d2 <- a[2] * stats::dexp(x, 1 / tau[2])
    w  <- d1 / (d1 + d2); w[!is.finite(w)] <- 0.5
    a   <- c(mean(w), 1 - mean(w))
    tau <- c(sum(w * x) / sum(w), sum((1 - w) * x) / sum(1 - w))
    if (any(!is.finite(tau)) || any(tau <= 0)) return(NA_real_)
  }
  o <- order(tau); tau <- tau[o]; a <- a[o]
  # A record with one closed component has no bursts to find, and forcing a
  # tcrit onto it invents them.
  if (tau[2] / tau[1] < 3 || min(a) < 0.02) return(NA_real_)
  f <- function(t) a[1] * exp(-t / tau[1]) - a[2] * (1 - exp(-t / tau[2]))
  lo <- tau[1] * 1e-3; hi <- tau[2]
  if (!is.finite(f(lo)) || !is.finite(f(hi)) || f(lo) * f(hi) > 0) return(NA_real_)
  stats::uniroot(f, c(lo, hi), tol = 1e-9)$root
}

burst_of <- function(dw, tcrit) {
  if (!is.finite(tcrit)) return(NULL)
  sep <- dw$state == "Closed" & dw$dwell_ms >= tcrit
  b   <- cumsum(sep)
  d <- dw[!sep, , drop = FALSE]; b <- b[!sep]
  if (!nrow(d)) return(NULL)
  # The first and last bursts are cut off by the ends of the record, so both
  # would drag the mean burst length down.
  inner <- b > min(b) & b < max(b)
  d <- d[inner, , drop = FALSE]; b <- b[inner]
  if (!nrow(d)) return(NULL)
  tibble::tibble(burst = b, state = d$state, ms = d$dwell_ms) |>
    dplyr::group_by(burst) |>
    dplyr::summarise(length_ms = sum(ms),
                     n_open    = sum(state == "Open"),
                     open_ms   = sum(ms[state == "Open"]),
                     n_gap     = sum(state == "Closed"),
                     gap_ms    = sum(ms[state == "Closed"]),
                     .groups = "drop") |>
    dplyr::filter(n_open > 0)
}

res_burst <- purrr::map(manifest$phenotype, function(ph) {
  arms <- recs[[ph]]; fs <- FS_MAIN[[ph]]
  dwl <- lapply(names(arms), function(a) dwell_times(arms[[a]]$channels, fs[[a]]))
  names(dwl) <- names(arms)
  # One tcrit per phenotype, taken from the real record and imposed on every
  # arm. Per-arm tcrits would make the burst counts incomparable.
  tcrit <- fit_tcrit(dwl$Real$dwell_ms[dwl$Real$state == "Closed"])
  per_burst <- purrr::map_dfr(names(arms), function(a) {
    b <- burst_of(dwl[[a]], tcrit)
    if (is.null(b) || !nrow(b)) return(NULL)
    dplyr::mutate(b, phenotype = ph, source = a)
  })
  if (!nrow(per_burst)) return(list(summary = NULL, per_burst = NULL))
  list(
    per_burst = per_burst,
    summary = per_burst |>
      dplyr::group_by(phenotype, source) |>
      dplyr::summarise(
        tcrit_ms = tcrit, n_bursts = dplyr::n(),
        `Burst length (ms)`       = mean(length_ms),
        `Openings per burst`      = mean(n_open),
        `Open time in burst (ms)` = sum(open_ms) / sum(n_open),
        `Gap in burst (ms)`       = if (sum(n_gap)) sum(gap_ms) / sum(n_gap) else NA_real_,
        `Burst Popen`             = sum(open_ms) / sum(length_ms),
        .groups = "drop") |>
      dplyr::mutate(`Bursts per s` = purrr::map_dbl(source, function(a)
        sum(per_burst$source == a) / (nrow(arms[[a]]) / fs[[a]]))))
})

burst      <- purrr::map_dfr(res_burst, "summary")
bursts_raw <- purrr::map_dfr(res_burst, "per_burst")

knitr::kable(dplyr::select(burst, phenotype, source, tcrit_ms, n_bursts,
                           dplyr::everything()), digits = 4,
             caption = paste("Table 5. Burst kinetics. tcrit is fitted on the",
                             "real record of each phenotype and applied to all",
                             "three arms; a phenotype whose closed times are",
                             "monoexponential yields no tcrit and drops out."))

burst_cmp <- burst |>
  dplyr::select(-tcrit_ms, -n_bursts) |>
  tidyr::pivot_longer(-c(phenotype, source), names_to = "metric") |>
  tidyr::pivot_wider(names_from = source, values_from = value) |>
  dplyr::filter(is.finite(Real), is.finite(Diffusion), is.finite(GAN)) |>
  dplyr::mutate(err_diffusion = abs(Diffusion - Real),
                err_gan       = abs(GAN - Real),
                favours = dplyr::if_else(err_diffusion < err_gan, "Diffusion", "GAN"))

knitr::kable(burst_cmp, digits = 4,
             caption = "Table 6. Burst metrics, absolute error against the real record.")

knitr::kable(
  burst_cmp |>
    dplyr::group_by(metric) |>
    dplyr::summarise(n = dplyr::n(),
                     n_favour_diffusion = sum(favours == "Diffusion"),
                     .groups = "drop"),
  caption = "Table 7. Unanimity of the burst metrics across phenotypes.")

readr::write_csv(burst,     file.path(CFG$out_dir, "burst_kinetics.csv"))
readr::write_csv(burst_cmp, file.path(CFG$out_dir, "burst_comparison.csv"))
```

## Burst Figures
```{r burst_fig, fig.height = 7 }
p_burst <- bursts_raw |>
  dplyr::transmute(phenotype, source,
                   `Burst length (ms)` = length_ms,
                   `Openings per burst` = as.numeric(n_open)) |>
  tidyr::pivot_longer(-c(phenotype, source), names_to = "quantity") |>
  dplyr::filter(value > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  ggplot(aes(value, colour = source)) +
  geom_area(aes(y = after_stat(sqrt(count)), fill = source),
            stat = "bin", bins = 40, alpha = 0.15,
            position = "identity", colour = NA) +
  geom_freqpoly(aes(y = after_stat(sqrt(count))), bins = 40, linewidth = 0.5) +
  facet_grid(phenotype ~ quantity, scales = "free") +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(x = NULL, y = expression(sqrt(count))) +
  theme_channel()
p_burst

save_fig(p_burst, "fig_burst", width = 8, height = 1.5 * n_ph + 1)
readr::write_csv(bursts_raw, file.path(CFG$out_dir, "bursts_per_burst.csv"))
```
bars
```{r burst_bars,  fig.height = 7}
d_burst <- bursts_raw |>
  dplyr::transmute(phenotype, source,
                   `Burst length (ms)` = length_ms,
                   `Openings per burst` = as.numeric(n_open)) |>
  tidyr::pivot_longer(-c(phenotype, source), names_to = "quantity") |>
  dplyr::filter(value > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")))

# Breaks per quantity, since burst length and opening count span quite
# different ranges, and log-spaced so every bar is one width on the log axis.
log_breaks <- function(v, n = 41) {
  r <- range(v)
  if (r[1] == r[2]) return(NULL)
  10^seq(log10(r[1]), log10(r[2]), length.out = n)
}
brk_burst <- split(d_burst$value, d_burst$quantity) |> lapply(log_breaks)

layer_hist <- function(q) {
  d <- dplyr::filter(d_burst, quantity == q)
  list(
    geom_histogram(data = d, aes(y = after_stat(sqrt(count)), fill = source),
                   breaks = brk_burst[[q]], alpha = 0.15,
                   position = "identity", colour = NA),
    geom_histogram(data = d, aes(y = after_stat(sqrt(count))), fill = NA,
                   breaks = brk_burst[[q]], linewidth = 0.4,
                   position = "identity")
  )
}


p_burst <- ggplot(d_burst, aes(value, colour = source)) +
  lapply(names(brk_burst), layer_hist) +
  facet_grid(
    phenotype ~ quantity,
    scales = "free",
    labeller = labeller(
      phenotype = function(x) gsub("_", "\n", x, fixed = TRUE)
    )
  ) +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(
    x = NULL,
    y = expression(sqrt(count))
  ) +
  theme_channel() +
  theme(
    strip.text.y = element_text(
      angle = 270,
      hjust = 0.5,
      vjust = 0.5
    )
  ) +
  guides(
    colour = guide_legend(
      override.aes = list(
        fill = scales::alpha(
          pal[c("Real", "Diffusion", "GAN")],
          0.15
        ),
        linewidth = 0.4
      )
    )
  )


p_burst


save_fig(p_burst, "fig_burst", width = 8, height = 1.5 * n_ph + 1)
readr::write_csv(bursts_raw, file.path(CFG$out_dir, "bursts_per_burst.csv"))
```



## Figure: slow envelope

```{r fig-env, fig.height = 6}
p_env <- pull_all("env") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  ggplot(aes(value, colour = source, fill = source)) +
  geom_density(alpha = 0.15, linewidth = 0.5) +
  facet_wrap(~ phenotype, scales = "free", ncol = 2) +
  scale_colour_manual(values = pal) + scale_fill_manual(values = pal) +
  labs(x = "Envelope current (pA)", y = "Density") +
  theme_channel()
p_env
save_fig(p_env, "fig_envelope", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
```

## Figure: manifold projections (optional)

```{r fig-manifold, eval = CFG$do_manifold, fig.width = 7.5, fig.height = 12}
if (!all(vapply(c("uwot", "Rtsne"), requireNamespace, logical(1), quietly = TRUE))) {
  stop("Set CFG$do_manifold = FALSE or install uwot and Rtsne.")
}

proj <- purrr::map_dfr(seq_len(nrow(manifest)), function(i) {
  row <- manifest[i, ]
  d <- list(Real = read_record(row$raw), Diffusion = read_record(row$ddpm))
  if (!is.na(row$gan)) d$GAN <- read_record(row$gan)
  Wp <- lapply(d, function(x) make_windows(x$current, CFG$proj_win, CFG$n_windows_proj))
  X <- do.call(rbind, Wp)
  lab <- rep(names(Wp), vapply(Wp, nrow, integer(1)))
  X <- scale(X); X[!is.finite(X)] <- 0
  um <- uwot::umap(X, n_neighbors = 15, min_dist = 0.1, n_components = 2)
  ts <- Rtsne::Rtsne(X, dims = 2, perplexity = 30, check_duplicates = FALSE,
                     pca = TRUE)$Y
  dplyr::bind_rows(
    tibble::tibble(x = um[, 1], y = um[, 2], source = lab, method = "UMAP"),
    tibble::tibble(x = ts[, 1], y = ts[, 2], source = lab, method = "t-SNE")
  ) |> dplyr::mutate(phenotype = row$phenotype)
})

proj <- proj |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  dplyr::group_by(phenotype, method) |>
  dplyr::mutate(x = scales::rescale(x), y = scales::rescale(y)) |>
  dplyr::ungroup()

# One arm per column, over a grey ghost of the full cloud so each panel still
# shows where the other two sat. Rescaling per phenotype lets the three columns
# share fixed scales, which free scales previously prevented.
plot_proj <- function(d) {
  ggplot(d, aes(x, y)) +
    geom_point(
      data = dplyr::select(d, -source),
      colour = "grey88",
      size = 0.7
    ) +
    geom_point(
      aes(colour = source),
      size = 1.1,
      alpha = 0.85
    ) +
    facet_grid(
      phenotype ~ source,
      labeller = labeller(
        phenotype = function(x) gsub("_", "\n", x)
      )
    ) +
    scale_colour_manual(values = pal) +
    guides(
      colour = guide_legend(
        override.aes = list(size = 4, alpha = 1)
      )
    ) +
    labs(x = NULL, y = NULL) +
    theme_channel(base_size = 11) +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      legend.text = element_text(size = 12),
      legend.key.size = unit(1.5, "lines"),
      panel.spacing = unit(0.7, "lines"),
      strip.text.y = element_text(
        angle = 270,
        hjust = 0.5,
        vjust = 0.5
      )
    )
}
p_proj_umap <- plot_proj(dplyr::filter(proj, method == "UMAP"))
p_proj_tsne <- plot_proj(dplyr::filter(proj, method == "t-SNE"))
p_proj_umap
p_proj_tsne

save_fig(p_proj_umap, "fig_manifold_umap", width = 7.5, height = 2.3 * n_ph + 1)
save_fig(p_proj_tsne, "fig_manifold_tsne", width = 7.5, height = 2.3 * n_ph + 1)
readr::write_csv(proj, file.path(CFG$out_dir, "manifold_coordinates.csv"))
```

# 7. Numbers for the manuscript

Rows keyed to `[[placeholder]]` names. Pooled values are medians across
phenotypes, with the range, because five records do not support a mean and an
interval being read as an estimate.

```{r placeholders}
fmt <- function(x, d = 4) formatC(x, digits = d, format = "g")

pooled <- scored |>
  dplyr::group_by(metric, method) |>
  dplyr::summarise(median = median(value), lo = min(value), hi = max(value),
                   .groups = "drop")

placeholders <- dplyr::bind_rows(
  tibble::tibble(placeholder = "n_phenotypes", value = as.character(n_ph),
                 note = "paired seed records"),
  tibble::tibble(placeholder = "n_samples_total",
                 value = format(sum(info$n_real), big.mark = ","),
                 note = "seed samples across all phenotypes"),
  # Rate before unique, or floating-point dust splits one rate into several.
  tibble::tibble(placeholder = "sample_khz",
                 value = paste(unique(round(sort(info$fs_real) / 1000, 3)),
                               collapse = ", "),
                 note = "acquisition rate(s), kHz"),
  pooled |>
    dplyr::transmute(
      placeholder = paste0(gsub("[^A-Za-z0-9]+", "_", tolower(metric)), "_",
                           tolower(method)),
      value = sprintf("%s (%s to %s)", fmt(median), fmt(lo), fmt(hi)),
      note = paste("median across phenotypes, range in brackets")),
  paired_stats |>
    dplyr::transmute(
      placeholder = paste0("unanimity_", gsub("[^A-Za-z0-9]+", "_", tolower(metric))),
      value = sprintf("%d/%d favour Diffusion", n_favour_diff, n),
      note = verdict)
)

knitr::kable(placeholders, caption = "Values keyed to the manuscript placeholders.")
```

## What this comparison does and does not settle

Both methods are scored identically above, so the numbers are comparable as
measures of how much the synthetic data resemble the real record. That is the
only question they answer.

They do not settle the question the paper is about. The GAN emits its raw trace
and its `Channels` column together, so its labels are a network output; the
diffusion pipeline is handed its labels before it starts and never touches them.
A GAN record could match the real data better on every metric here and still be
the less useful training set, because a downstream idealiser trained against it
is being scored on another network's opinion rather than on truth. If the GAN
wins on similarity, say so plainly and make the ground-truth argument on its own
terms.

# Session information

```{r session}
sessionInfo()
```
