---
title: 'DiffuChannel benchmark: diffusion vs DeepGANnel across all phenotypes'
author: "Barrett-Jolley lab"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    toc: true
    number_sections: false
    md_extensions: -tex_math_dollars
  html_document:
    toc: true
    df_print: paged
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 8, fig.height = 5, dpi = 150)
```

# Purpose

One pass over every phenotype folder. For each one it reads the real seed record,
the DeepGANnel record and the diffusion (DDPM) record, computes the same metric
set on the two synthetic arms, and then pools the five paired comparisons.

Every arm is scored identically. Whether that flatters the diffusion model OR NOT is not
this paper's business.

```{r packages}
required <- c("dplyr", "tidyr", "ggplot2", "readr", "purrr", "arrow",
              "scales", "patchwork", "tibble","Rtsne", "uwot", "kableExtra")
missing  <- required[!vapply(required, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing)) {
  stop("Install first: install.packages(c(\"",
       paste(missing, collapse = "\", \""), "\"))")
}

library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(purrr)
library(arrow)
library(patchwork)
library(uwot)
library(Rtsne)

set.seed(42)
```

# Configuration

```{r config}
CFG <- list(
  # ---- where things live --------------------------------------------------
  pheno_root = "./Phenotypes",   # holds Phenotype_A, Phenotype_B, ...
  batch_dir  = "./batch_out",    # holds Phenotype_A_DDPM.parquet, ...
  out_dir    = "./benchmark_out",

  raw_pattern = "raw",           # substring marking the seed csv (case-insensitive)
  gan_pattern = "gan",           # substring marking the DeepGANnel csv
  only        = NULL,            # e.g. c("Phenotype_A") to run a subset

  # ---- recording parameters ----------------------------------------------
  sample_hz  = 1e4,              # fallback only; a Time column always wins
  filter_hz  = 1e3,

  # ---- how to treat a generated file's own time column --------------------
  #   "seed"  every arm inherits the raw record's rate (treat as equivalent:
  #           the generated file's time column is cosmetic and wrong)
  #   "own"   an arm with a time column is taken at its word (treat as honest:
  #           it really was produced at that rate)
  rate_policy  = "seed",
  rate_compare = TRUE,           # also run the other policy where they differ

  # ---- must match training ------------------------------------------------
  seq_len     = 128,
  slow_factor = 64,

  # ---- analysis knobs -----------------------------------------------------
  n_windows_mmd = 400,           # O(n^2) in the kernel: keep modest
  n_perm        = 200,
  # Spectral segment length. "auto" is resolved per phenotype (see below); a
  # number is applied to every phenotype and must fit all of them.
  psd_seg       = "auto",
  psd_seg_min   = 256,           # below this, stop rather than report a coarse PSD
  psd_n_seg_min = 20,            # target segments per arm; fewer is flagged, not fatal
  equalise_noise = TRUE,         # hold every arm to the same segment and sample count
  acf_lag_max   = 200,
  # Which conductance level supplies the noise segments. "auto" picks, per
  # phenotype, whichever of closed or open every arm can support; a record that
  # barely closes has no baseline but plenty of open-level material.
  noise_level   = "auto",        # "auto", "closed" or "open"
  ks_subsample  = 2e5,
  trace_secs    = 0.3,

  # ---- generation block geometry ------------------------------------------
  # Only used by the block-join (seam) test. These must match what
  # diffusion_app.py actually generated with, or the boundaries are looked for
  # in the wrong places and the test passes for a reason that has nothing to do
  # with the inpainting.
  block_len      = 32768,       # generation block length, samples
  context_len    = 512,         # carried context at block joins, samples

  # ---- optional extras ----------------------------------------------------
  do_manifold    = TRUE,        # UMAP / t-SNE, needs uwot + Rtsne, slow
  n_windows_proj = 400,

  proj_win       = 512,

  # ClusterSignificance on the manifold projections, for [[cluster_p]]. The
  # package is on Bioconductor, not CRAN, so it is optional and the chunk
  # skips rather than fails when it is absent. The permutation is the slow
  # part: n_perm_cluster permutations per phenotype per projection method.
  do_cluster_test = TRUE,
  n_perm_cluster  = 100,

  cache = FALSE                   # reuse per-phenotype results if present
  #Careful with caching - it can cause real confusion!!
)

dir.create(CFG$out_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(file.path(CFG$out_dir, "cache"), showWarnings = FALSE, recursive = TRUE)

pal <- c("Real" = "#1b1b1b", "Diffusion" = "#c0392b", "GAN" = "#8e44ad",
         "Real floor" = "#7f8c8d", "White noise" = "#bdc3c7")

theme_channel <- function(base_size = 10) {
  theme_classic(base_size = base_size) +
    theme(strip.background = element_blank(),
          strip.text = element_text(face = "bold", hjust = 0),
          legend.position = "top",
          legend.title = element_blank(),
          legend.key.size = unit(0.8, "lines"),
          axis.text = element_text(colour = "black"))
}

# Figures are written as LZW-compressed TIFF, which is what the journals ask
# for and is lossless, so no generation of the figure is ever thrown away.
#
# The device is opened explicitly rather than left to ggsave's device
# dispatch: ggsave has to guess how to forward `compression` to whichever
# device it picks from the file extension, and when it guesses wrong it
# silently writes an uncompressed TIFF. Opening the device by hand means the
# compression argument goes exactly where it is meant to.
#
# ragg is preferred where it is installed because its text rendering on
# Windows is much better than the stock device's; grDevices::tiff is the
# fallback and takes the same compression argument.
save_fig <- function(plot, name, width = 8, height = 5, dpi = 300) {
  path <- file.path(CFG$out_dir, paste0(name, ".tiff"))

  if (requireNamespace("ragg", quietly = TRUE)) {
    ragg::agg_tiff(path, width = width, height = height, units = "in",
                   res = dpi, compression = "lzw", background = "white")
  } else {
    grDevices::tiff(path, width = width, height = height, units = "in",
                    res = dpi, compression = "lzw", bg = "white")
  }
  on.exit(grDevices::dev.off(), add = TRUE)

  print(plot)
  invisible(plot)
}

# kable(), scaled to fit the page in PDF so wide tables stop cramming their
# headers together. latex_options do nothing when knitting to HTML, so this
# is safe either way. hold_position stops the table drifting away from its
# caption once it's wrapped in a resize box.
nice_kable <- function(x, caption = NULL, ...) {
  if (!is.null(caption) && knitr::is_latex_output())
    caption <- gsub("([_$&%#])", "\\\\\\1", caption)
  knitr::kable(x, caption = caption, ...) |>
    kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
}
```

# Reading records

The seed and GAN csv files were not written by this pipeline, so their headers
cannot be assumed. Named columns are used when they are recognisable; otherwise
the columns are inferred the same way `batch_run.py` infers them, which is the
only way the R and Python sides can agree on which column was which. The
resolved mapping is printed in the manifest below and is worth reading once.

```{r fn-read}
normalise_names <- function(nms) gsub("[_-]", " ", trimws(tolower(nms)))

# 10%-of-range quantisation, as in batch_run.py stage_seed()
quantise <- function(x) {
  span <- diff(range(x, na.rm = TRUE))
  if (span > 0) round(x / (0.1 * span)) else rep(0, length(x))
}

# A channels column may hold no more than this many discrete levels (MAX_LEVELS
# in batch_run.py). Used both to infer the column and to check a named one.
MAX_LEVELS <- 10

is_time_col <- function(x) {
  d <- diff(x)
  length(d) > 1 && all(d > 0) &&
    isTRUE(all.equal(d, rep(d[1], length(d)), tolerance = 1e-3))
}

infer_cols <- function(df) {
  # The column count decides whether a time column exists at all: three columns
  # means one of them is time, two means there is none. Which one it is still
  # has to be found, because the order varies.
  time_col <- NA_integer_
  if (ncol(df) >= 3) {
    hit <- which(vapply(df, is_time_col, logical(1)))
    if (!length(hit)) {
      stop(ncol(df), " columns but none of them increases in constant steps, ",
           "so the time column cannot be identified")
    }
    time_col <- hit[1]
  }
  rest <- setdiff(seq_along(df), if (is.na(time_col)) integer(0) else time_col)
  lev  <- lapply(rest, function(j) sort(unique(quantise(df[[j]]))))
  n    <- vapply(lev, length, integer(1))
  k    <- which.min(n)
  if (n[k] > MAX_LEVELS) {
    stop("no column with <= ", MAX_LEVELS, " discrete levels; level counts ",
         paste(n, collapse = ", "))
  }
  others <- setdiff(seq_along(rest), k)
  list(time   = time_col,
       chan   = rest[k],
       cur    = rest[others[which.max(n[others])]],
       levels = lev[[k]])
}

read_record <- function(path) {
  df <- if (grepl("\\.parquet$", path, ignore.case = TRUE)) {
    as.data.frame(arrow::read_parquet(path))
  } else {
    d <- suppressWarnings(as.data.frame(
      readr::read_csv(path, show_col_types = FALSE, name_repair = "minimal")))
    # A file with no header gets its first data row promoted to names; spot that
    # by the names parsing as numbers, and re-read.
    if (all(!is.na(suppressWarnings(as.numeric(names(d)))))) {
      d <- as.data.frame(readr::read_csv(path, col_names = FALSE,
                                         show_col_types = FALSE))
    }
    d
  }

  df <- df[, vapply(df, function(x) is.numeric(x) || is.logical(x) ||
                      !all(is.na(suppressWarnings(as.numeric(as.character(x))))),
                    logical(1)), drop = FALSE]
  df[] <- lapply(df, function(x) suppressWarnings(as.numeric(as.character(x))))
  df <- df[stats::complete.cases(df), , drop = FALSE]
  if (ncol(df) < 2 || nrow(df) < 100) stop("too few usable rows/columns")

  nn  <- normalise_names(names(df))
  idx <- function(target) { h <- which(nn %in% target); if (length(h)) h[1] else NA_integer_ }
  i_chan <- idx(c("channels", "channel", "ideal", "idealisation", "idealization"))
  i_cur  <- idx(c("noisy current", "current", "raw", "raw current", "i"))
  i_time <- idx(c("time", "t"))

  # Headers are either absent or correct, so a recognised pair is taken at face
  # value. Anything else is inferred. Column order is never assumed either way.
  if (!is.na(i_chan) && !is.na(i_cur) && i_chan != i_cur) {
    how  <- "named columns"
    chan <- as.integer(round(df[[i_chan]]))
  } else {
    inf <- infer_cols(df)
    i_chan <- inf$chan; i_cur <- inf$cur; i_time <- inf$time
    how  <- "inferred"
    chan <- match(quantise(df[[i_chan]]), inf$levels) - 1L
  }

  out <- tibble::tibble(channels = as.numeric(chan),
                        current  = as.numeric(df[[i_cur]]))
  out$time <- if (!is.na(i_time)) as.numeric(df[[i_time]]) else
    (seq_len(nrow(out)) - 1) / CFG$sample_hz
  out <- dplyr::filter(out, is.finite(channels), is.finite(current))

  attr(out, "map") <- sprintf(
    "%s: chan=col%d cur=col%d time=%s", how, i_chan, i_cur,
    if (is.na(i_time)) sprintf("none, assuming %g Hz", CFG$sample_hz)
    else paste0("col", i_time))
  dplyr::select(out, time, channels, current)
}

fs_of <- function(d, fallback = CFG$sample_hz) {
  if (nrow(d) < 3) return(fallback)
  dt <- stats::median(diff(d$time), na.rm = TRUE)
  if (!is.finite(dt) || dt <= 0) fallback else 1 / dt
}

#' Crop a record to its first n rows, keeping the column map
#'
#' Subsetting a tibble drops attributes, so the "map" attribute that
#' read_record attached is stashed and restored. Anything reading it
#' downstream (inventory's has_time, the manifest print) would otherwise
#' see NULL.
#'
#' @param d  A record as returned by read_record: a tibble with columns
#'           time, channels, current, and a "map" attribute describing
#'           which source column supplied each.
#' @param n  Integer, the number of rows to keep, counted from the start
#'           of the record. Assumed to be no greater than nrow(d).
#' @return   The same tibble with rows 1:n and the "map" attribute intact.
crop_rows <- function(d, n) {
  m <- attr(d, "map")
  d <- d[seq_len(n), ]
  attr(d, "map") <- m
  d
}
```

# Metric functions

Carried over from `DiffuChannel_analysis.Rmd` without change.

```{r fn-metrics}
make_windows <- function(x, win, n_max = Inf, stride = NULL) {
  n <- length(x)
  if (n < win) return(matrix(numeric(0), nrow = 0, ncol = win))
  if (is.null(stride)) stride <- max(1L, floor((n - win) / max(1L, min(n_max, n))))
  starts <- seq(1L, n - win + 1L, by = stride)
  if (length(starts) > n_max) starts <- sample(starts, n_max)
  t(vapply(starts, function(s) x[s:(s + win - 1L)], numeric(win)))
}

sq_dists <- function(A, B) {
  pmax(outer(rowSums(A^2), rowSums(B^2), "+") - 2 * tcrossprod(A, B), 0)
}

# Median heuristic: gamma = 1 / median pairwise squared distance. The fixed 1/L
# used during training underflows to zero on current in pA and pins the
# statistic to its finite-sample floor.
median_gamma <- function(X, Y, n_sub = 200) {
  A <- X[sample(nrow(X), min(n_sub, nrow(X))), , drop = FALSE]
  B <- Y[sample(nrow(Y), min(n_sub, nrow(Y))), , drop = FALSE]
  m <- stats::median(sq_dists(A, B))
  if (!is.finite(m) || m <= 0) 1 / ncol(X) else 1 / m
}

mmd_rbf <- function(X, Y, gamma = NULL, unbiased = TRUE) {
  m <- nrow(X); n <- nrow(Y)
  if (m < 2 || n < 2) return(NA_real_)
  if (is.null(gamma)) gamma <- median_gamma(X, Y)
  Kxx <- exp(-gamma * sq_dists(X, X))
  Kyy <- exp(-gamma * sq_dists(Y, Y))
  Kxy <- exp(-gamma * sq_dists(X, Y))
  if (unbiased) {
    (sum(Kxx) - sum(diag(Kxx))) / (m * (m - 1)) +
      (sum(Kyy) - sum(diag(Kyy))) / (n * (n - 1)) - 2 * mean(Kxy)
  } else {
    mean(Kxx) + mean(Kyy) - 2 * mean(Kxy)
  }
}

mmd_permutation_p <- function(X, Y, n_perm = CFG$n_perm, gamma = NULL) {
  if (is.null(gamma)) gamma <- median_gamma(X, Y)
  obs <- mmd_rbf(X, Y, gamma)
  Z <- rbind(X, Y); m <- nrow(X)
  null <- replicate(n_perm, {
    i <- sample(nrow(Z))
    mmd_rbf(Z[i[seq_len(m)], , drop = FALSE], Z[i[-seq_len(m)], , drop = FALSE], gamma)
  })
  s <- stats::sd(null)
  list(mmd = obs,
       p   = (1 + sum(null >= obs)) / (n_perm + 1),
       z   = if (is.finite(s) && s > 0) (obs - mean(null)) / s else NA_real_)
}

# Noise segments are taken at a single conductance level, so that no gating step
# falls inside a segment. Runs are returned separately and never concatenated:
# gluing them together puts a step discontinuity at every join, and a step is
# broadband, lifting the high-frequency end of the spectrum and dragging the ACF
# towards zero in proportion to how often the record gates. That bias falls
# hardest on exactly the busy records the comparison is meant to include.
at_level <- function(d, level) {
  if (identical(level, "closed")) d$channels == 0 else d$channels >= max(d$channels)
}

level_runs <- function(d, level, min_len = 1L) {
  r <- rle(at_level(d, level))
  ends <- cumsum(r$lengths); starts <- ends - r$lengths + 1
  keep <- which(r$values & r$lengths >= min_len)
  lapply(keep, function(i) d$current[starts[i]:ends[i]])
}

longest_at <- function(d, level) {
  r <- rle(at_level(d, level))
  if (!any(r$values)) 0L else max(r$lengths[r$values])
}

# Lengths of the runs at one level, without materialising the runs themselves.
# Used to choose a segment length before any spectrum is computed.
run_lengths_at <- function(d, level) {
  r <- rle(at_level(d, level))
  r$lengths[r$values]
}

# How many Welch segments a set of run lengths yields at 50% overlap. Mirrors
# the loop in welch_psd, so the two cannot drift apart.
seg_count <- function(lens, nperseg) {
  lens <- lens[lens >= nperseg]
  if (!length(lens)) return(0L)
  as.integer(sum(1L + (lens - nperseg) %/% (nperseg %/% 2L)))
}

# Welch across a list of runs: every segment from every run long enough to hold
# one contributes to a single average periodogram.
#
# n_seg_max thins that set to a fixed count, taken evenly across the record. A
# Welch estimate's variance falls as 1/n_seg, so an arm that gates more often
# has shorter runs, fewer segments and a noisier spectrum, and the RMS distance
# to real inflates for reasons that have nothing to do with the model. Holding
# both synthetic arms to the same count makes that variance common to both and
# cancels it from their difference.
welch_psd <- function(runs, fs, nperseg, n_seg_max = Inf) {
  if (is.numeric(runs)) runs <- list(runs)
  runs <- runs[vapply(runs, length, integer(1)) >= nperseg]
  if (!length(runs)) stop("no run at the chosen level holds ", nperseg, " samples")

  step <- nperseg %/% 2
  idx <- do.call(rbind, lapply(seq_along(runs), function(i)
    cbind(run = i, start = seq(1L, length(runs[[i]]) - nperseg + 1L, by = step))))
  n_avail <- nrow(idx)

  if (is.finite(n_seg_max) && n_seg_max < n_avail) {
    k <- as.integer(n_seg_max)
    idx <- idx[1L + ((0:(k - 1L)) * n_avail) %/% k, , drop = FALSE]
  }

  w <- 0.5 - 0.5 * cos(2 * pi * (seq_len(nperseg) - 1) / (nperseg - 1))  # Hann
  U <- sum(w^2)
  acc <- numeric(nperseg %/% 2 + 1)
  for (r in seq_len(nrow(idx))) {
    s   <- idx[r, "start"]
    seg <- runs[[idx[r, "run"]]][s:(s + nperseg - 1L)]
    seg <- seg - mean(seg)
    X   <- stats::fft(seg * w)[seq_len(nperseg %/% 2 + 1)]
    acc <- acc + (Mod(X)^2) / (fs * U)
  }
  nseg <- nrow(idx)
  psd <- acc / nseg
  psd[-c(1, length(psd))] <- 2 * psd[-c(1, length(psd))]
  tibble::tibble(freq = seq(0, fs / 2, length.out = length(psd)), psd = psd,
                 n_seg = nseg, n_seg_avail = n_avail) |>
    dplyr::filter(freq > 0)
}

# Length-weighted mean of the per-run ACFs, for the same reason. n_max is the
# same argument as n_seg_max above: the ACF's sampling variance also falls with
# the material behind it, so arms are held to a common sample budget, spent on
# the runs in order and trimming the one that crosses it.
acf_of <- function(runs, lag_max, fs, n_max = Inf) {
  if (is.numeric(runs)) runs <- list(runs)
  runs <- runs[vapply(runs, length, integer(1)) > lag_max]
  if (!length(runs)) stop("no run at the chosen level is longer than ", lag_max,
                          " samples")

  if (is.finite(n_max)) {
    lens  <- vapply(runs, length, integer(1))
    spent <- c(0L, cumsum(lens))[seq_along(lens)]
    take  <- pmin(lens, pmax(0L, as.integer(n_max) - spent))
    keep  <- take > lag_max
    if (any(keep)) runs <- Map(function(x, k) x[seq_len(k)], runs[keep], take[keep])
  }

  w <- vapply(runs, length, integer(1))
  A <- vapply(runs, function(x)
    stats::acf(x - mean(x), lag.max = lag_max, plot = FALSE,
               demean = TRUE)$acf[, 1, 1], numeric(lag_max + 1))
  tibble::tibble(lag_n  = seq_len(lag_max + 1),
                 lag_ms = 1000 * (seq_len(lag_max + 1) - 1) / fs,
                 acf    = as.numeric(A %*% w / sum(w)),
                 n_used = sum(w))
}

dwell_times <- function(ideal, fs) {
  r <- rle(as.integer(round(ideal)))
  tibble::tibble(level = r$values, n_sample = r$lengths,
                 dwell_ms = 1000 * r$lengths / fs) |>
    dplyr::mutate(state = ifelse(level == 0, "Closed", "Open"))
}

# Half-amplitude idealiser. Two Gaussians of common width, fitted by EM.
# Hard assignment fails when one level holds 1% of the record: the far tail of
# the crowded level outnumbers the sparse one on its own side of any threshold
# and drags the estimate onto itself. Responsibilities do not have that
# problem. Polarity says which of the two fitted levels is shut.
idealise_half_amplitude <- function(current, unit_pA = NULL, baseline = NULL,
                                    n_levels = 1, polarity = 1) {
  x <- current[is.finite(current)]

  if (is.null(baseline) || is.null(unit_pA)) {
    # Fitting on a regular subsample keeps the knit time sane; a level holding
    # 1% of the record still contributes ~1000 points here.
    xf <- if (length(x) > 1e5) x[seq(1, length(x), length.out = 1e5)] else x

    # Seed outside both bulks, not at the 1st/99th percentile: in a record
    # that is open 99% of the time the 1st percentile is still open. The
    # starting weight is taken from the split at the midpoint rather than set
    # to a half, so the sparse component is not invited to bid for the
    # crowded one.
    mu  <- stats::quantile(xf, c(0.001, 0.999), names = FALSE)
    sg  <- max(stats::mad(xf), .Machine$double.eps^0.5)
    w   <- min(max(mean(xf < mean(mu)), 1e-3), 1 - 1e-3)
    tol <- 1e-8 * sg
    
    for (i in 1:300) {
      d1 <- stats::dnorm(xf, mu[1], sg) * w
      d2 <- stats::dnorm(xf, mu[2], sg) * (1 - w)
      r  <- d2 / (d1 + d2)
      r[!is.finite(r)] <- 0.5
      s2 <- sum(r); s1 <- sum(1 - r)
      if (s1 < 1 || s2 < 1) break
      new_mu <- c(sum((1 - r) * xf) / s1, sum(r * xf) / s2)
      new_sg <- max(sqrt((sum((1 - r) * (xf - new_mu[1])^2) +
                          sum(r * (xf - new_mu[2])^2)) / length(xf)),
                    .Machine$double.eps^0.5)
      w <- s1 / length(xf)
      done <- max(abs(c(new_mu - mu, new_sg - sg))) < tol
      mu <- new_mu; sg <- new_sg
      if (done) break
    }

    mu   <- sort(mu)
    shut <- if (polarity >= 0) mu[1] else mu[2]
    open <- if (polarity >= 0) mu[2] else mu[1]
    minority <- min(w, 1 - w)

    if (!is.finite(open - shut) || abs(open - shut) < 3 * sg ||
        minority < 5e-4) {
        out <- rep(NA_integer_, length(current))
        attr(out, "why") <- sprintf(
          "sep %.3g, width %.3g, rarer level %.3g",
          abs(open - shut), sg, minority)
        return(out)
    }

    if (is.null(baseline)) baseline <- shut
    if (is.null(unit_pA))  unit_pA  <- open - shut
  }

  lvl <- (current - baseline) / unit_pA
  pmin(pmax(as.integer(round(lvl)), 0L), as.integer(n_levels))
}

idealise_label_anchored_cv <- function(
    current,
    labels,
    n_levels = 1L,
    polarity = 1,
    block_len = 1000L,
    trim = 0.10,
    min_state_n = 50L
) {
  stopifnot(length(current) == length(labels))

  ok <- is.finite(current) & is.finite(labels)
  labels <- as.integer(labels > 0)

  n <- length(current)
  block <- ((seq_len(n) - 1L) %/% block_len) + 1L
  fold <- block %% 2L

  recovered <- rep(NA_integer_, n)
  fit_info <- vector("list", 2L)

  robust_centre <- function(x) {
    x <- x[is.finite(x)]

    if (!length(x)) {
      return(NA_real_)
    }

    mean(x, trim = trim)
  }

  for (test_fold in 0:1) {
    train <- ok & fold != test_fold
    test  <- ok & fold == test_fold

    x_closed <- current[train & labels == 0L]
    x_open   <- current[train & labels > 0L]

    if (length(x_closed) < min_state_n ||
        length(x_open) < min_state_n) {
      next
    }

    closed_centre <- robust_centre(x_closed)
    open_centre   <- robust_centre(x_open)

    if (!is.finite(closed_centre) ||
        !is.finite(open_centre) ||
        closed_centre == open_centre) {
      next
    }

    # Respect the polarity defined from the real record.
    if (polarity >= 0) {
      baseline <- closed_centre
      unit_pA  <- open_centre - closed_centre
    } else {
      baseline <- closed_centre
      unit_pA  <- open_centre - closed_centre
    }

    lvl <- (current[test] - baseline) / unit_pA

    recovered[test] <- pmin(
      pmax(as.integer(round(lvl)), 0L),
      as.integer(n_levels)
    )

    fit_info[[test_fold + 1L]] <- list(
      n_closed = length(x_closed),
      n_open = length(x_open),
      baseline = baseline,
      open_centre = open_centre,
      unit_pA = unit_pA
    )
  }

  if (all(is.na(recovered))) {
    attr(recovered, "why") <-
      "cross-fitted label anchoring lacked enough observations in one or both states"
  }

  attr(recovered, "fit") <- fit_info
  attr(recovered, "method") <- "cross-fitted label-anchored half-amplitude"

  recovered
}
# Cohen's kappa on two logical vectors. Chance agreement is high when a record
# is nearly always open or nearly always shut, which is exactly when raw
# agreement flatters an arm, so kappa rather than accuracy.
cohen_kappa <- function(a, b) {
  po <- mean(a == b)
  pe <- mean(a) * mean(b) + mean(!a) * mean(!b)
  if (!is.finite(pe) || pe >= 1) NA_real_ else (po - pe) / (1 - pe)
}

block_mean <- function(x, factor) {
  factor <- max(1L, as.integer(factor))
  nb <- length(x) %/% factor
  if (nb < 1) return(mean(x))
  colMeans(matrix(x[seq_len(nb * factor)], nrow = factor))
}

ks_safe <- function(a, b, n_max = CFG$ks_subsample) {
  a <- a[is.finite(a)]; b <- b[is.finite(b)]
  if (length(a) > n_max) a <- sample(a, n_max)
  if (length(b) > n_max) b <- sample(b, n_max)
  suppressWarnings(stats::ks.test(a, b))
}

# Window choice for the example traces, ported from _best_window in
# batch_run.py so the R figures and the Python previews pick the same way:
# prefer a stretch whose Popen sits in [lo, hi], else the one nearest the middle
# of that band. Taking the first n samples instead lands on whatever the record
# happens to open with, which for a sparse phenotype is often no events at all.
best_window <- function(chan, n, lo = 0.4, hi = 0.8) {
  if (length(chan) <= n) return(1L)
  cs <- c(0, cumsum(as.numeric(chan)))
  k  <- (cs[(n + 1):length(cs)] - cs[1:(length(cs) - n)]) / n
  inside <- which(k >= lo & k <= hi)
  if (length(inside)) return(inside[ceiling(length(inside) / 2)])
  which.min(abs(k - 0.5 * (lo + hi)))
}

rms_diff <- function(a, b) sqrt(mean((a - b)^2))

# Two-component exponential mixture on a set of dwell times, by EM, giving the
# time constants the manuscript asks for rather than a mean dwell. Returns one
# component rather than two when two cannot be supported, because a fit whose
# components have converged onto each other reports a spurious pair of taus
# that reads like a result. Same degeneracy test as fit_tcrit: a slow-to-fast
# ratio under min_ratio, or a component holding almost no area, is one
# component wearing two labels.
#
# Note this cannot rescue a record whose dwells are only a few samples long.
# At 10 kHz a mean closed dwell of 0.39 ms is under four samples, the dwell
# distribution is a handful of discrete values, and no exponential mixture is
# meaningful. Those come back as one component and should be reported as mean
# dwells instead.
fit_exp_mixture <- function(x, n_iter = 500, tol = 1e-10, min_n = 20L,
                            min_ratio = 3, min_area = 0.02) {
  x <- x[is.finite(x) & x > 0]

  one <- function() tibble::tibble(
    n_components = 1L,
    tau_fast_ms  = if (length(x)) mean(x) else NA_real_,
    tau_slow_ms  = NA_real_,
    area_fast    = NA_real_,
    mean_ms      = if (length(x)) mean(x) else NA_real_,
    n_events     = length(x))

  if (length(x) < min_n) return(one())

  tau <- as.numeric(stats::quantile(x, c(0.25, 0.9), names = FALSE))
  if (!(tau[2] > tau[1])) return(one())
  a <- c(0.5, 0.5)
  ll_old <- -Inf

  for (i in seq_len(n_iter)) {
    d <- cbind(a[1] * stats::dexp(x, 1 / tau[1]),
               a[2] * stats::dexp(x, 1 / tau[2]))
    s <- rowSums(d)
    s[s <= 0] <- .Machine$double.xmin
    w  <- d[, 1] / s
    sw <- sum(w)
    if (!is.finite(sw) || sw < 1 || (length(x) - sw) < 1) return(one())
    a   <- c(sw / length(x), 1 - sw / length(x))
    tau <- c(sum(w * x) / sw, sum((1 - w) * x) / (length(x) - sw))
    if (any(!is.finite(tau)) || any(tau <= 0)) return(one())
    ll <- sum(log(s))
    if (is.finite(ll) && is.finite(ll_old) &&
        abs(ll - ll_old) < tol * abs(ll)) break
    ll_old <- ll
  }

  o <- order(tau); tau <- tau[o]; a <- a[o]
  if (tau[2] / tau[1] < min_ratio || min(a) < min_area) return(one())

  tibble::tibble(n_components = 2L, tau_fast_ms = tau[1],
                 tau_slow_ms = tau[2], area_fast = a[1],
                 mean_ms = mean(x), n_events = length(x))
}
```

# 1. Find the files

```{r discover}
find_one <- function(folder, pattern) {
  f <- list.files(
    folder,
    pattern = "\\.(csv|parquet)$",
    full.names = TRUE,
    ignore.case = TRUE
  )

  hit <- f[
    grepl(pattern, basename(f), ignore.case = TRUE)
  ]

  if (!length(hit)) return(NA_character_)

  if (length(hit) > 1) {
    warning(
      folder, ": ", length(hit),
      " files match '", pattern,
      "', using ", basename(hit[1])
    )
  }

  hit[1]
}

folders <- list.dirs(CFG$pheno_root, recursive = FALSE)
if (!is.null(CFG$only)) folders <- folders[basename(folders) %in% CFG$only]
if (!length(folders)) stop("No phenotype folders in ", CFG$pheno_root)

parquets <- list.files(CFG$batch_dir, pattern = "\\.parquet$", full.names = TRUE)

manifest <- purrr::map_dfr(folders, function(fd) {
  tag <- gsub(" ", "_", basename(fd))
  ddpm <- parquets[grepl(paste0("^", tag), basename(parquets))]
  if (!length(ddpm)) ddpm <- parquets[grepl(tag, basename(parquets), fixed = TRUE)]
  tibble::tibble(
    phenotype = tag,
    raw  = find_one(fd, CFG$raw_pattern),
    gan  = find_one(fd, CFG$gan_pattern),
    ddpm = if (length(ddpm)) ddpm[1] else NA_character_
  )
})

nice_kable(
  manifest |> dplyr::mutate(dplyr::across(c(raw, gan, ddpm), basename)),
  caption = "Manifest. Check that each row's three files belong to the same phenotype."
)

bad <- manifest |> dplyr::filter(is.na(raw) | is.na(ddpm))
if (nrow(bad)) {
  warning("Dropping phenotypes with no seed or no DDPM file: ",
          paste(bad$phenotype, collapse = ", "))
  manifest <- dplyr::filter(manifest, !is.na(raw), !is.na(ddpm))
}
if (!nrow(manifest)) stop("Nothing left to analyse.")
```

# 2. Pre-flight

Every record is read and checked before a single metric is computed, so that a
record which cannot support the analysis stops the run here rather than becoming
an `NA` that propagates quietly into the pooled tables.

All records are held in memory for the rest of the notebook. That is the price
of reading each file exactly once; if it becomes a problem, run fewer phenotypes
at a time with `CFG$only`.

```{r preflight}
recs <- purrr::map(seq_len(nrow(manifest)), function(i) {
  row <- manifest[i, ]
  out <- list(Real = read_record(row$raw), Diffusion = read_record(row$ddpm))
  if (!is.na(row$gan)) out$GAN <- read_record(row$gan)
  out
})
names(recs) <- manifest$phenotype

n_before <- purrr::imap_dfr(recs, function(arms, ph)
  tibble::tibble(phenotype = ph, arm = names(arms),
                 n_read = vapply(arms, nrow, integer(1))))

# Table 1 of the manuscript, built here because this is the last point at
# which the seed records are at their full length. Every other row count in
# the paper should be traced to this file rather than re-typed: transcribing
# them by hand has already produced a column that is one short in every row.
seed_inventory <- purrr::imap_dfr(recs, function(arms, ph) {
  d  <- arms$Real
  fs <- fs_of(d)
  tibble::tibble(
    phenotype  = ph,
    seed_file  = basename(manifest$raw[manifest$phenotype == ph]),
    n_rows     = nrow(d),
    sample_khz = round(fs / 1000, 3),
    duration_s = nrow(d) / fs,
    max_levels = max(d$channels),
    popen      = mean(d$channels > 0),
    mean_pA    = mean(d$current),
    sd_pA      = stats::sd(d$current))
})

nice_kable(seed_inventory, digits = 4,
           caption = paste("Table 1. Seed records as read, before any",
                           "cropping. n_rows is the number of usable rows",
                           "read from each file and is the figure the",
                           "manuscript should quote."))

readr::write_csv(seed_inventory, file.path(CFG$out_dir, "table1_seed_data.csv"))

# The block-join test needs the diffusion arm at its full generated length.
# Cropping to the shortest arm shortens it by up to a quarter, which moves the
# last join past the end of the record and would leave the test quietly
# examining fewer joins than the record actually contains. Stashed here, before
# the crop, and used by nothing except the seam figures.
recs_full_diffusion <- purrr::map(recs, function(arms) arms$Diffusion)

# Every arm of a phenotype is cropped to the row count of its shortest arm.
# Rows, not seconds and not file size: the metrics are computed per sample, and
# several of them carry a length bias. MMD's z-score scales with the number of
# windows, the permutation p loses power when there are fewer, the KS statistic
# has a finite-sample floor, and both spectral distances are differences between
# estimates whose variance falls with the material behind them. An arm that is
# merely shorter would otherwise score differently for that reason alone.
for (ph in names(recs)) {
  arms <- recs[[ph]]
  min_len <- Inf
  for (a in names(arms)) { min_len <- min(min_len, nrow(arms[[a]])) }
  for (a in names(arms)) { arms[[a]] <- crop_rows(arms[[a]], min_len) }
  recs[[ph]] <- arms
}

crop_report <- n_before |>
  dplyr::left_join(
    purrr::imap_dfr(recs, function(arms, ph)
      tibble::tibble(phenotype = ph, arm = names(arms),
                     n_used = vapply(arms, nrow, integer(1)))),
    by = c("phenotype", "arm")) |>
  dplyr::mutate(kept = n_used / n_read)

nice_kable(crop_report, digits = 3,
             caption = "Rows read and rows retained after cropping each phenotype's arms to its shortest.")
readr::write_csv(crop_report, file.path(CFG$out_dir, "crop_report.csv"))

inventory <- purrr::imap_dfr(recs, function(arms, ph) {
  purrr::imap_dfr(arms, function(d, a) {
    closed <- d$current[d$channels == 0]
    open   <- d$current[d$channels > 0]
    tibble::tibble(
      phenotype = ph, arm = a, n = nrow(d),
      has_time = !grepl("time=none", attr(d, "map")),
      fs = fs_of(d),
      longest_closed = longest_at(d, "closed"),
      longest_open   = longest_at(d, "open"),
      popen = mean(d$channels > 0), max_level = max(d$channels),
      baseline = median(closed),
      unit_amp = median(open) - median(closed),
      cols = attr(d, "map"))
  })
})

# A record that has been min-max scaled, or left in the wrong units, has a
# unitary amplitude unlike its seed's. Every amplitude metric then compares
# units rather than distributions, and the arm loses on bookkeeping.
scale_check <- inventory |>
  dplyr::group_by(phenotype) |>
  dplyr::mutate(rel = unit_amp / unit_amp[arm == "Real"]) |>
  dplyr::ungroup() |>
  dplyr::filter(arm != "Real", rel < 1 / 3 | rel > 3)
if (nrow(scale_check)) {
  stop("Unitary amplitude differs from the seed by more than 3x in: ",
       paste(sprintf("%s/%s (%.3g vs %.3g)", scale_check$phenotype,
                     scale_check$arm, scale_check$unit_amp,
                     scale_check$unit_amp / scale_check$rel), collapse = "; "),
       ". That is a scaling or units problem, not a modelling result: the ",
       "amplitude metrics would compare units rather than distributions. ",
       "Check whether a min-max scaled copy has been used.")
}

nice_kable(inventory, digits = 4,
             caption = "Pre-flight inventory. Every record that will be analysed.")
readr::write_csv(inventory, file.path(CFG$out_dir, "preflight_inventory.csv"))
```

## Sample rate

Only the seed record's time column reflects a real acquisition.
`diffusion_app.py` writes its export time as `arange(n) * d * SAMPLE_DT` with
`SAMPLE_DT` fixed at 1e-4, so a DDPM record always claims 10 kHz whatever it was
trained on. A GAN record's time column is whatever DeepGANnel wrote, and when
that disagrees with the seed there are two readings, neither of which can be
settled from the file alone:

- **equivalent** (`rate_policy = "seed"`): the time column is cosmetic. The
  record is a sample-for-sample stand-in for the seed and inherits its rate.
- **honest** (`rate_policy = "own"`): the record really was produced at that
  rate, so it has half the bandwidth and its events last twice as long in
  wall-clock terms.

Relabelling a genuinely decimated record makes its dwell times and spectrum
wrong; taking a cosmetic label at face value does the same in the other
direction. So both are computed and compared below.

```{r rate}
seed_fs <- inventory |>
  dplyr::filter(arm == "Real") |>
  dplyr::mutate(fs = dplyr::if_else(has_time, fs, CFG$sample_hz))
seed_fs <- setNames(seed_fs$fs, seed_fs$phenotype)

rates_under <- function(policy) {
  inventory |>
    dplyr::transmute(
      phenotype, arm,
      fs = if (identical(policy, "own")) {
        dplyr::if_else(has_time, fs, unname(seed_fs[phenotype]))
      } else {
        unname(seed_fs[phenotype])
      })
}

rate_map <- function(policy) {
  r <- rates_under(policy)
  split(setNames(r$fs, r$arm), r$phenotype)
}

disputed <- rates_under("own") |>
  dplyr::rename(fs_own = fs) |>
  dplyr::left_join(dplyr::rename(rates_under("seed"), fs_seed = fs),
                   by = c("phenotype", "arm")) |>
  dplyr::filter(abs(fs_own - fs_seed) / fs_seed > 0.01)

nice_kable(rates_under("seed") |>
               tidyr::pivot_wider(names_from = arm, values_from = fs),
             digits = 1, caption = "Rate under the equivalent policy (Hz).")

if (nrow(disputed)) {
  nice_kable(disputed, digits = 1,
               caption = "Arms where the two policies disagree. These are the only ones whose rate-dependent metrics change between policies.")
} else {
  cat("No arm's own time column disagrees with its seed; the two policies coincide.\n")
}

FS_MAIN <- rate_map(CFG$rate_policy)
FS_ALT  <- rate_map(if (identical(CFG$rate_policy, "seed")) "own" else "seed")

# Which way openings go, per phenotype, read once off the real record's own
# labels and imposed on every arm. A recording convention rather than a tuned
# parameter, and the same bargain as taking one tcrit from the real record.
POL_MAIN <- inventory |>
  dplyr::filter(arm == "Real") |>
  dplyr::transmute(phenotype, polarity = sign(unit_amp))
POL_MAIN <- setNames(POL_MAIN$polarity, POL_MAIN$phenotype)

if (!all(POL_MAIN %in% c(-1, 1))) {
  stop("A real record has no measurable unitary amplitude, so the direction ",
       "of its openings is undefined: ",
       paste(names(POL_MAIN)[!(POL_MAIN %in% c(-1, 1))], collapse = ", "))
}
```


The lag-1 autocorrelation of the single-level noise, reported after the analysis,
is the physical evidence: decimation decorrelates adjacent samples, so a record
genuinely at half the rate sits well below its own seed while unaffected
phenotypes match.

## Noise level and segment length for the spectral metrics

The noise metrics need contiguous samples at one conductance level. Which level
that can be depends on the record: a quiet channel offers long closed periods, a
record at Popen 0.99 offers long open ones and closures only a sample or two
wide. The level is therefore chosen per phenotype as whichever *every* arm of
that phenotype can support, and reported so it can be stated in the legend.

Reporting an open-level spectrum is not a fudge: the requirement is that no
gating transition falls inside a segment, which the open level satisfies exactly
as well. It does measure something slightly different — open-channel noise sits
on top of the instrumentation noise — but it is the same quantity in both arms
of the comparison, which is all the paired metric needs.

The segment length is chosen per phenotype for the same reason the level is.
Only the three arms within a phenotype are ever compared, so nothing requires
A and E to share a segment length, and making them share one lets the single
worst record dictate the resolution for every other. What *is* required is that
the three arms of one phenotype share it, since the RMS distance is taken
between their spectra.

The `"auto"` rule also runs the other way from the obvious one. Taking the
largest segment a record tolerates maximises resolution and minimises the
number of segments, which is the wrong trade for a statistic whose variance
falls as 1/n_seg. So `"auto"` takes the largest power of two that still gives
every arm `psd_n_seg_min` segments, and only falls back towards the resolution
floor when no length manages it.

```{r level-and-seg}
cap_at <- inventory |>
  dplyr::group_by(phenotype) |>
  dplyr::summarise(closed = min(longest_closed), open = min(longest_open),
                   .groups = "drop")

levels_used <- cap_at |>
  dplyr::mutate(
    level = if (identical(CFG$noise_level, "auto")) {
      dplyr::if_else(open > closed, "open", "closed")
    } else {
      CFG$noise_level
    },
    cap = dplyr::if_else(level == "open", open, closed))

if (any(levels_used$cap == 0)) {
  z <- levels_used |> dplyr::filter(cap == 0)
  stop("No usable ", paste(unique(z$level), collapse = "/"), " segments at all in: ",
       paste(z$phenotype, collapse = ", "),
       ". Set CFG$noise_level to the other level, or exclude these phenotypes.")
}

nice_kable(levels_used, digits = 0,
             caption = "Longest run every arm can supply at each level, and the level used. 'cap' is that phenotype's ceiling on segment length.")

LEVEL_OF <- setNames(levels_used$level, levels_used$phenotype)
CAP_OF   <- setNames(levels_used$cap,   levels_used$phenotype)

# Run lengths at the chosen level, per arm, per phenotype. Everything below is
# decided from these, before a single spectrum is computed.
lens_of <- purrr::imap(recs, function(arms, ph)
  lapply(arms, run_lengths_at, level = LEVEL_OF[[ph]]))

seg_counts_at <- function(lens, nperseg)
  vapply(lens, seg_count, integer(1), nperseg = nperseg)

# The floor on segment length: coarser than psd_seg_min is not worth reporting,
# and the ACF needs segments longer than its own maximum lag.
SEG_FLOOR <- as.integer(2^ceiling(log2(max(CFG$psd_seg_min, CFG$acf_lag_max + 1))))

pick_nperseg <- function(lens, cap) {
  hi <- 2^floor(log2(cap))
  if (hi < SEG_FLOOR) return(NA_integer_)
  for (nps in 2^(floor(log2(hi)):floor(log2(SEG_FLOOR)))) {
    if (min(seg_counts_at(lens, nps)) >= CFG$psd_n_seg_min) return(as.integer(nps))
  }
  SEG_FLOOR
}

if (identical(CFG$psd_seg, "auto")) {
  NPERSEG_OF <- vapply(names(recs),
                       function(ph) pick_nperseg(lens_of[[ph]], CAP_OF[[ph]]),
                       integer(1))
  starved <- names(NPERSEG_OF)[is.na(NPERSEG_OF)]
  if (length(starved)) {
    stop(paste(starved, collapse = ", "), ": no arm supports ", SEG_FLOOR,
         " contiguous samples at the chosen level (caps ",
         paste(CAP_OF[starved], collapse = ", "), "). A spectrum that coarse ",
         "is not worth reporting. Exclude the phenotype with CFG$only, or ",
         "lower psd_seg_min and acf_lag_max deliberately and say so in the ",
         "methods.")
  }
} else {
  nps <- as.integer(CFG$psd_seg)
  if (nps < CFG$psd_seg_min) {
    stop("CFG$psd_seg = ", nps, " is below CFG$psd_seg_min = ", CFG$psd_seg_min, ".")
  }
  if (nps <= CFG$acf_lag_max) {
    stop("CFG$psd_seg = ", nps, " is not longer than CFG$acf_lag_max = ",
         CFG$acf_lag_max, ", so the ACF would be estimated over runs too short ",
         "to support it. Lower acf_lag_max.")
  }
  short <- names(CAP_OF)[CAP_OF < nps]
  if (length(short)) {
    stop("CFG$psd_seg = ", nps, " but ", paste(short, collapse = ", "),
         " supports only ", paste(CAP_OF[short], collapse = ", "),
         " samples at the chosen level. Set CFG$psd_seg = \"auto\".")
  }
  NPERSEG_OF <- setNames(rep(nps, length(recs)), names(recs))
}

seg_tbl <- purrr::imap_dfr(lens_of, function(lens, ph) {
  n <- seg_counts_at(lens, NPERSEG_OF[[ph]])
  tibble::tibble(
    phenotype = ph, level = LEVEL_OF[[ph]], cap = CAP_OF[[ph]],
    nperseg = NPERSEG_OF[[ph]],
    seg_ms  = 1000 * NPERSEG_OF[[ph]] / FS_MAIN[[ph]][["Real"]],
    res_hz  = FS_MAIN[[ph]][["Real"]] / NPERSEG_OF[[ph]],
    Real = unname(n["Real"]), Diffusion = unname(n["Diffusion"]),
    GAN = if ("GAN" %in% names(n)) unname(n["GAN"]) else NA_integer_,
    n_used = min(n))
})

nice_kable(seg_tbl, digits = 1,
             caption = paste("Segment length chosen per phenotype, and the",
                             "segments each arm can supply at it. n_used is",
                             "what every arm is held to when",
                             "equalise_noise is TRUE."))
readr::write_csv(seg_tbl, file.path(CFG$out_dir, "psd_segments.csv"))

thin <- seg_tbl |> dplyr::filter(n_used < CFG$psd_n_seg_min)
if (nrow(thin)) {
  warning("Fewer than CFG$psd_n_seg_min = ", CFG$psd_n_seg_min,
          " segments per arm in: ",
          paste(sprintf("%s (%d)", thin$phenotype, thin$n_used), collapse = ", "),
          ". Welch's variance goes as 1/n_seg, so those spectra are noisy and ",
          "their PSD RMS is not a reliable measure of spectral mismatch. The ",
          "equalisation makes the arms comparable with each other but cannot ",
          "manufacture precision. Flagged as psd_n_seg_ok in the provenance ",
          "table; say so in the methods or drop the phenotype from the ",
          "spectral metrics.")
}

if (dplyr::n_distinct(levels_used$level) > 1 ||
    dplyr::n_distinct(NPERSEG_OF) > 1) {
  cat("Level and/or segment length differ between phenotypes, so the noise\n",
      "metrics are paired within a phenotype but not comparable in absolute\n",
      "terms across them. State both per panel in the figure legend.\n", sep = "")
}
```
# 3. Per-phenotype analysis

Each phenotype is reduced to a long table of metrics, plus the curves needed for
the pooled figures. `direction` records which way is better, because pooling a
distance with a summary statistic that should *match* the real record would be
meaningless otherwise.

```{r analyse}
analyse_one <- function(row, arms_all, fs, level, nperseg) {
  ph <- row$phenotype

  real    <- arms_all$Real
  arms    <- arms_all[setdiff(names(arms_all), "Real")]
  has_gan <- "GAN" %in% names(arms)

  # ---- windows and MMD ----------------------------------------------------
  W <- list(Real = make_windows(real$current, CFG$seq_len, CFG$n_windows_mmd))
  for (a in names(arms)) W[[a]] <- make_windows(arms[[a]]$current, CFG$seq_len,
                                                CFG$n_windows_mmd)
  half <- floor(nrow(real) / 2)
  W_r1 <- make_windows(real$current[1:half], CFG$seq_len, CFG$n_windows_mmd)
  W_r2 <- make_windows(real$current[(half + 1):nrow(real)], CFG$seq_len, CFG$n_windows_mmd)
  W_wn <- make_windows(rnorm(nrow(real), mean(real$current), sd(real$current)),
                       CFG$seq_len, CFG$n_windows_mmd)

  # One bandwidth for every comparison in this phenotype, or the numbers are
  # not on a single scale.
  gam <- median_gamma(W$Real, W$Diffusion)

  # ---- single-level noise -------------------------------------------------
  # Runs are kept separate and every record has already been shown to hold at
  # least one segment, so there is no missing-data branch here.
  runs <- lapply(arms_all, level_runs, level = level, min_len = nperseg)

  # Equal row counts do not give equal usable material: an arm that gates more
  # often has shorter runs and so fewer segments. Both noise metrics are RMS
  # differences between estimates, so the noisier arm is penalised for its
  # gating twice, once in the dwell metrics and again here. Holding every arm
  # to the counts the thinnest can supply makes that variance common.
  lens       <- lapply(runs, function(x) vapply(x, length, integer(1)))
  seg_avail  <- vapply(lens, seg_count, integer(1), nperseg = nperseg)
  samp_avail <- vapply(lens, function(L) sum(L[L > CFG$acf_lag_max]), integer(1))
  seg_cap    <- if (isTRUE(CFG$equalise_noise)) min(seg_avail)  else Inf
  samp_cap   <- if (isTRUE(CFG$equalise_noise)) min(samp_avail) else Inf

  psd <- purrr::imap_dfr(runs, function(x, a)
    dplyr::mutate(welch_psd(x, fs[[a]], nperseg, seg_cap), source = a))
  acf_tbl <- purrr::imap_dfr(runs, function(x, a)
    dplyr::mutate(acf_of(x, CFG$acf_lag_max, fs[[a]], samp_cap), source = a))

  # Lag-1 autocorrelation of the single-level noise, per arm. A record genuinely
  # at a lower rate than the one assumed for it sits well below its own seed.
  lag1 <- acf_tbl |> dplyr::filter(lag_n == 2)
  lag1 <- setNames(lag1$acf, lag1$source)

  # Two arms at different rates have different frequency and lag grids, so both
  # comparisons interpolate onto a shared axis rather than pivoting on the grid
  # value. Under the equivalent policy the grids already coincide and the
  # interpolation is a no-op.
  curve <- function(tbl, a, xvar, yvar) {
    d <- tbl[tbl$source == a, ]
    list(x = d[[xvar]], y = d[[yvar]])
  }

  psd_rms <- function(a) {
    r <- curve(psd, "Real", "freq", "psd")
    g <- curve(psd, a, "freq", "psd")
    hi <- min(CFG$filter_hz, max(r$x), max(g$x))
    lo <- max(min(r$x), min(g$x))
    grid <- exp(seq(log(lo), log(hi), length.out = 512))
    rms_diff(approx(r$x, log10(r$y), grid)$y, approx(g$x, log10(g$y), grid)$y)
  }

  acf_rms <- function(a) {
    r <- curve(acf_tbl, "Real", "lag_ms", "acf")
    g <- curve(acf_tbl, a, "lag_ms", "acf")
    grid <- seq(0, min(max(r$x), max(g$x)), length.out = CFG$acf_lag_max + 1)
    rms_diff(approx(r$x, r$y, grid)$y, approx(g$x, g$y, grid)$y)
  }

  # ---- slow envelope ------------------------------------------------------
  env <- list(Real = block_mean(real$current, CFG$slow_factor))
  for (a in names(arms)) env[[a]] <- block_mean(arms[[a]]$current, CFG$slow_factor)
  We <- lapply(env, function(e) make_windows(e, min(64, length(e)), CFG$n_windows_mmd))

  # ---- dwell times --------------------------------------------------------
  dw <- purrr::imap_dfr(arms_all, function(d, a)
    dplyr::mutate(dwell_times(d$channels, fs[[a]]), source = a))
  mean_dwell <- function(a, s) {
    v <- dw |> dplyr::filter(source == a, state == s) |> dplyr::pull(dwell_ms)
    if (!length(v)) stop(ph, "/", a, ": no ", s, " dwells to average")
    mean(v)
  }

  level_pts <- function(d) d$current[at_level(d, level)]

  # ---- assemble -----------------------------------------------------------
  # One metric name whatever the level, so the pooled comparison keeps all
  # phenotypes; the level itself is recorded in the provenance table.
  sd_name <- "Noise SD (pA)"
  M <- function(metric, method, direction, value)
    tibble::tibble(phenotype = ph, metric = metric, method = method,
                   direction = direction, value = as.numeric(value))

  rows <- dplyr::bind_rows(
    M("MMD vs real",          "Real floor",  "reference", mmd_rbf(W_r1, W_r2, gamma = gam)),
    M("MMD vs real",          "White noise", "reference", mmd_rbf(W$Real, W_wn, gamma = gam)),
    M(sd_name,                "Real",        "reference", sd(level_pts(real))),
    M("Envelope SD (pA)",     "Real",        "reference", sd(env$Real)),
    M("Open probability",     "Real",        "reference", mean(real$channels > 0)),
    M("Mean open dwell (ms)", "Real",        "reference", mean_dwell("Real", "Open")),
    M("Mean closed dwell (ms)", "Real",      "reference", mean_dwell("Real", "Closed")),
    purrr::map_dfr(names(arms), function(a) {
      pm <- mmd_permutation_p(W$Real, W[[a]], gamma = gam)
      dplyr::bind_rows(
        M("MMD vs real",            a, "lower",  mmd_rbf(W$Real, W[[a]], gamma = gam)),
        M("MMD permutation p",      a, "higher", pm$p),
        M("MMD z-score",            a, "lower",  pm$z),
        M("KS amplitude",           a, "lower",
          unname(ks_safe(real$current, arms[[a]]$current)$statistic)),
        M("log10 PSD RMS",          a, "lower",  psd_rms(a)),
        M("ACF RMS",                a, "lower",  acf_rms(a)),
        M("MMD envelope",           a, "lower",  mmd_rbf(We$Real, We[[a]])),
        M(sd_name,                  a, "match_real", sd(level_pts(arms[[a]]))),
        M("Envelope SD (pA)",       a, "match_real", sd(env[[a]])),
        M("Open probability",       a, "match_real", mean(arms[[a]]$channels > 0)),
        M("Mean open dwell (ms)",   a, "match_real", mean_dwell(a, "Open")),
        M("Mean closed dwell (ms)", a, "match_real", mean_dwell(a, "Closed"))
      )
    })
  )

  # ---- curves and provenance ---------------------------------------------
  # time_ms is built from the sample index and the rate assigned to the arm, not
  # from the file's own time column, which for a DDPM export is a fixed 1e-4
  # step regardless of what it was trained on and makes panels differ in width.
  traces <- purrr::imap_dfr(arms_all, function(d, a) {
    n <- min(nrow(d), round(CFG$trace_secs * fs[[a]]))
    s <- best_window(d$channels, n)
    d[s:(s + n - 1), ] |>
      dplyr::mutate(source = a, phenotype = ph,
                    time_ms = 1000 * (dplyr::row_number() - 1) / fs[[a]],
                    win_start = s, win_popen = mean(channels > 0))
  })

  amp <- purrr::imap_dfr(arms_all, function(d, a)
    tibble::tibble(current = d$current, source = a, phenotype = ph))

  info <- tibble::tibble(
    phenotype = ph,
    seed_file = basename(row$raw),
    gan_file  = if (has_gan) basename(row$gan) else "none",
    ddpm_file = basename(row$ddpm),
    n_real = nrow(real),
    n_ddpm = nrow(arms$Diffusion),
    n_gan  = if (has_gan) nrow(arms$GAN) else NA_integer_,
    popen_real = mean(real$channels > 0),
    max_level_real = max(real$channels),
    fs_real = fs[["Real"]],
    fs_ddpm = fs[["Diffusion"]],
    fs_gan  = if (has_gan) fs[["GAN"]] else NA_real_,
    noise_level = level,
    psd_nperseg = nperseg,
    psd_seg_ms = 1000 * nperseg / fs[["Real"]],
    # Segments each arm could have supplied, then the count they were all held
    # to. Indexed by name rather than position: Real is first by construction
    # today, which is not a reason to rely on it.
    psd_n_seg_real = unname(seg_avail["Real"]),
    psd_n_seg_ddpm = unname(seg_avail["Diffusion"]),
    psd_n_seg_gan  = if (has_gan) unname(seg_avail["GAN"]) else NA_integer_,
    psd_n_seg_used = psd$n_seg[psd$source == "Real"][1],
    psd_n_seg_ok   = min(seg_avail) >= CFG$psd_n_seg_min,
    acf_n_used     = acf_tbl$n_used[acf_tbl$source == "Real"][1],
    lag1_real = unname(lag1["Real"]),
    lag1_ddpm = unname(lag1["Diffusion"]),
    lag1_gan  = if (has_gan) unname(lag1["GAN"]) else NA_real_,
    gamma = gam
  )

  list(metrics = rows,
       psd = dplyr::mutate(psd, phenotype = ph),
       acf = dplyr::mutate(acf_tbl, phenotype = ph),
       dwell = dplyr::mutate(dw, phenotype = ph),
       env = purrr::imap_dfr(env, function(e, a)
         tibble::tibble(idx = seq_along(e), value = e, source = a, phenotype = ph)),
       traces = traces,
       amp = amp,
       info = info)
}

# The cache is keyed on the segment length, whether the arms were equalised,
# the rates in force and the policy label: results computed under different
# assumptions are not comparable with these and must not be silently reused.
run_one <- function(row, fs, policy) {
  ph  <- row$phenotype
  nps <- NPERSEG_OF[[ph]]
  key <- sprintf("%s_seg%d_%s_%s_%s_%s", ph, nps, LEVEL_OF[[ph]], policy,
                 if (isTRUE(CFG$equalise_noise)) "eq" else "raw",
                 paste(sprintf("%s%g", substr(names(fs), 1, 1), fs), collapse = ""))
  cache_file <- file.path(CFG$out_dir, "cache", paste0(key, ".rds"))
  if (CFG$cache && file.exists(cache_file)) {
    message("Reusing cached ", key)
    return(readRDS(cache_file))
  }
  message("Analysing ", ph, " (", policy, " rates)")
  res <- analyse_one(row, recs[[ph]], fs, LEVEL_OF[[ph]], nps)
  if (CFG$cache) saveRDS(res, cache_file)
  res
}

results <- purrr::map(seq_len(nrow(manifest)), function(i) {
  row <- manifest[i, ]
  run_one(row, FS_MAIN[[row$phenotype]], CFG$rate_policy)
})
names(results) <- manifest$phenotype

pull_all <- function(what) purrr::map_dfr(results, what)

metrics <- pull_all("metrics")
info    <- pull_all("info")

# Nothing missing gets past this point. An NA here would become a phenotype
# silently dropped from a paired comparison, which is worse than a failed knit.
bad <- dplyr::filter(metrics, !is.finite(value))
if (nrow(bad)) {
  stop("Non-finite metrics, which must be understood rather than dropped:\n",
       paste(sprintf("  %s / %s / %s", bad$phenotype, bad$method, bad$metric),
             collapse = "\n"))
}

nice_kable(
  info |> dplyr::select(phenotype, seed_file, ddpm_file, gan_file,
                        n_real, n_ddpm, n_gan, popen_real, max_level_real),
  digits = 4, caption = "Records analysed."
)
nice_kable(info |> dplyr::select(phenotype, noise_level, fs_real, fs_ddpm,
                                   fs_gan, psd_nperseg, psd_seg_ms,
                                   psd_n_seg_real, psd_n_seg_ddpm,
                                   psd_n_seg_gan, psd_n_seg_used,
                                   psd_n_seg_ok, acf_n_used, gamma),
             digits = 4,
             caption = paste("Rate applied, spectral segment length, segments",
                             "each arm could supply, the count they were all",
                             "held to, and kernel bandwidth. psd_n_seg_ok is",
                             "FALSE where the spectra are too thinly averaged",
                             "to be worth much, however well equalised."))

nice_kable(info |> dplyr::select(phenotype, lag1_real, lag1_ddpm, lag1_gan),
             digits = 4,
             caption = "Lag-1 autocorrelation of the single-level noise. An arm well below its own real value is genuinely at a lower rate than the one assumed for it, not merely mislabelled.")
readr::write_csv(info, file.path(CFG$out_dir, "provenance.csv"))
readr::write_csv(metrics, file.path(CFG$out_dir, "metrics_long.csv"))
```

## Equivalent versus honest

Only rate-dependent metrics can move between the policies. The amplitude
metrics, MMD and open probability are computed per sample and are identical
either way, which is itself useful: it means the choice cannot be dodged by
appealing to the metrics that happen to be convenient.

```{r policy-compare, eval = CFG$rate_compare}
alt_phenos <- unique(disputed$phenotype)

if (!length(alt_phenos)) {
  cat("No arm's rate is disputed, so the two policies give identical results.\n")
} else {
  alt <- purrr::map(alt_phenos, function(ph) {
    row <- manifest[manifest$phenotype == ph, ]
    run_one(row, FS_ALT[[ph]], if (identical(CFG$rate_policy, "seed")) "own" else "seed")
  })
  names(alt) <- alt_phenos

  rate_sensitive <- c("log10 PSD RMS", "ACF RMS",
                      "Mean open dwell (ms)", "Mean closed dwell (ms)")

  cmp <- dplyr::bind_rows(
    metrics |> dplyr::filter(phenotype %in% alt_phenos) |>
      dplyr::mutate(policy = CFG$rate_policy),
    purrr::map_dfr(alt, "metrics") |>
      dplyr::mutate(policy = if (identical(CFG$rate_policy, "seed")) "own" else "seed")
  ) |>
    dplyr::filter(metric %in% rate_sensitive) |>
    dplyr::mutate(policy = dplyr::recode(policy, seed = "equivalent", own = "honest")) |>
    dplyr::select(phenotype, metric, method, policy, value) |>
    tidyr::pivot_wider(names_from = policy, values_from = value) |>
    dplyr::mutate(ratio = honest / equivalent)

  nice_kable(cmp, digits = 4,
               caption = "Rate-dependent metrics under both readings. Everything else is unchanged by construction.")
  readr::write_csv(cmp, file.path(CFG$out_dir, "rate_policy_comparison.csv"))

  lag_cmp <- dplyr::bind_rows(
    info |> dplyr::filter(phenotype %in% alt_phenos) |>
      dplyr::mutate(policy = CFG$rate_policy),
    purrr::map_dfr(alt, "info") |>
      dplyr::mutate(policy = if (identical(CFG$rate_policy, "seed")) "own" else "seed")
  ) |>
    dplyr::select(phenotype, policy, lag1_real, lag1_ddpm, lag1_gan)
  nice_kable(lag_cmp, digits = 4,
               caption = "Lag-1 autocorrelation, which does not depend on the rate label at all and is therefore the evidence.")
}
```

### How to read this

The lag-1 table is the part that decides it, because it is computed per sample
and is untouched by whichever label is applied. If the disputed arm's value sits
close to its own seed record, the samples are correlated like the seed's are and
the record is a sample-for-sample stand-in with a cosmetic time column:
**equivalent** is right, and the honest reading inflates its dwell times by the
rate ratio for no physical reason. If it sits well below, adjacent samples really
are less correlated, the record was produced at the lower rate, and **honest** is
right — but then the record has less bandwidth than the seed, which is a genuine
limitation of that arm and belongs in the results rather than being corrected
away.

The dwell ratios are the sanity check on whichever answer you accept: under the
honest reading a 5 kHz record's mean dwells double exactly, so if the honest
dwell times land implausibly far from the seed's while the equivalent ones sit
close, that is evidence for the equivalent reading and vice versa.

# 4. Scoring

Distances should be small, the permutation p should be large, and summary
statistics such as the noise SD should *match* the real record rather
than approach zero. The score column puts all three on a "smaller is better"
footing except the p, which is flagged separately.

```{r score}
refs <- metrics |>
  dplyr::filter(direction == "reference", method == "Real") |>
  dplyr::select(phenotype, metric, ref = value)

scored <- metrics |>
  dplyr::filter(direction %in% c("lower", "higher", "match_real")) |>
  dplyr::left_join(refs, by = c("phenotype", "metric")) |>
  dplyr::mutate(
    score = dplyr::case_when(
      direction == "match_real" ~ abs(value - ref),
      direction == "higher"     ~ 1 - value,
      TRUE                      ~ value
    ),
    better = "lower"
  ) |>
  dplyr::filter(is.finite(score))


nice_kable(
  scored |> dplyr::arrange(metric, method, phenotype) |>
    dplyr::select(metric, method, phenotype, value, ref, score),
  digits = 5, caption = "Raw values and the comparable score derived from them."
)
readr::write_csv(scored, file.path(CFG$out_dir, "scores_long.csv"))
```

## Head to head, per phenotype

```{r head-to-head}
head_to_head <- scored |>
  dplyr::select(metric, phenotype, method, value) |>
  tidyr::pivot_wider(names_from = method, values_from = value) |>
  dplyr::left_join(refs, by = c("phenotype", "metric")) |>
  dplyr::rename(`Reference (real)` = ref) |>
  dplyr::arrange(metric, phenotype)

nice_kable(head_to_head, digits = 5,
             caption = paste("Similarity metrics, computed identically",
                             "for both methods. Lower is better throughout except",
                             "the permutation p, where a large value means the",
                             "difference from real is no bigger than relabelling",
                             "produces."))
readr::write_csv(head_to_head, file.path(CFG$out_dir, "head_to_head_metrics.csv"))
```

# 5. Pooled comparison

The design is paired: both methods see the same seed record, so the quantity of
interest is the within-phenotype difference. Between-phenotype variation in
noise, open probability and duration is far larger than any method effect and
would swamp it if the records were treated as independent samples.

```{r paired}
pairable <- scored |>
  dplyr::filter(method %in% c("Diffusion", "GAN")) |>
  dplyr::group_by(metric, phenotype) |>
  dplyr::filter(dplyr::n_distinct(method) == 2) |>
  dplyr::ungroup()

n_ph <- dplyr::n_distinct(pairable$phenotype)

p_slope <- ggplot(pairable, aes(factor(method, levels = c("Diffusion", "GAN")),
                                score, group = phenotype)) +
  geom_line(colour = "grey60", linewidth = 0.4) +
  geom_point(aes(colour = method), size = 2.2) +
  facet_wrap(~ metric, scales = "free_y") +
  scale_colour_manual(values = pal, guide = "none") +
  labs(x = NULL, y = "Score (lower is better)")+
  theme_channel()
p_slope
save_fig(p_slope, "fig_paired_metrics", width = 9, height = 6.5)
```

Each line is one phenotype. Lines that all slope the same way are the reportable
result; lines that cross mean the ranking depends on which recording you happen
to analyse.

```{r pooled-summary}
wide <- pairable |>
  dplyr::select(metric, phenotype, better, method, score) |>
  tidyr::pivot_wider(names_from = method, values_from = score) |>
  dplyr::mutate(diff = Diffusion - GAN)

paired_stats <- wide |>
  dplyr::group_by(metric, better) |>
  dplyr::summarise(
    n              = dplyr::n(),
    n_ties         = sum(diff == 0),
    median_diff    = median(diff),
    mean_diff      = mean(diff),
    sd_diff        = sd(diff),
    n_favour_diff  = sum(if (dplyr::first(better) == "lower") diff < 0 else diff > 0),
    p_paired_t     = if (dplyr::n() > 2 && sd(diff) > 0)
                       stats::t.test(diff)$p.value else NA_real_,
    ci_lo          = if (dplyr::n() > 2 && sd(diff) > 0)
                       stats::t.test(diff)$conf.int[1] else NA_real_,
    ci_hi          = if (dplyr::n() > 2 && sd(diff) > 0)
                       stats::t.test(diff)$conf.int[2] else NA_real_,
    # The sign test must be taken in the same tail as `better`, and exact ties
    # are absence of evidence rather than evidence against: counting them as
    # unfavourable reported 0.0625 for a metric that was three-fifths tied.
    p_sign         = {
      fav <- if (dplyr::first(better) == "lower") sum(diff < 0) else sum(diff > 0)
      nz  <- sum(diff != 0)
      if (nz > 0) stats::binom.test(fav, nz)$p.value else NA_real_
    },
    .groups = "drop"
  ) |>
  dplyr::mutate(
    n_eff   = n - n_ties,
    # Decided from the counts, not from the verdict string. The string gets
    # " (n tied)" appended below, which defeated the != comparison that used
    # to count unanimous metrics and reported a tied metric as unanimous.
    unanimous = n_eff > 0 & (n_favour_diff == n_eff | n_favour_diff == 0),
    verdict = dplyr::case_when(
      n_eff == 0             ~ "no separation: all differences exactly tied",
      n_favour_diff == n_eff ~ "consistent: Diffusion",
      n_favour_diff == 0     ~ "consistent: GAN",
      TRUE                   ~ "inconsistent across phenotypes"),
    verdict = dplyr::if_else(n_ties > 0 & n_eff > 0,
                             paste0(verdict, " (", n_ties, " tied)"), verdict)) |>
  dplyr::select(metric, better, n, n_ties, median_diff, mean_diff, sd_diff,
                n_favour_diff, p_paired_t, ci_lo, ci_hi, p_sign, unanimous,
                verdict)

nice_kable(paired_stats, digits = 5,
             caption = paste("Pooled paired comparison, Diffusion minus",
                             "GAN within each phenotype. The t-test is on the",
                             "within-phenotype differences; the sign test excludes",
                             "exact ties, so with five records it cannot fall below",
                             "0.0625 and the unanimity count is the reportable",
                             "result."))
readr::write_csv(paired_stats, file.path(CFG$out_dir, "paired_summary.csv"))
```

## What this many phenotypes will and will not support

```{r power}
min_p_sign <- min(1, 2 * 0.5^n_ph)
cat(sprintf("Phenotypes pooled: %d\n", n_ph))
cat(sprintf("Smallest two-sided p a sign test can return here: %.4f\n", min_p_sign))

if (min_p_sign > 0.05) {
  cat("\nThat is above 0.05, so no sign test on these data can reach significance,\n",
      "however large or however unanimous the effect. The same floor applies to the\n",
      "Wilcoxon signed-rank test at this n. The p_sign column above is therefore\n",
      "reported for completeness and should not be leaned on.\n",
      "\nWhat these data do support: the per-phenotype values, the direction, whether\n",
      "it is unanimous, and the paired t interval on the differences. Those are the\n",
      "defensible summaries at this sample size.\n", sep = "")
} else {
  cat("\nA sign test is at least capable of reaching p < 0.05 here.\n")
}

cat(sprintf("\nUnanimous metrics: %d of %d.\n",
            sum(paired_stats$unanimous), nrow(paired_stats)))

if (any(paired_stats$n_ties > 0)) {
  cat("Metrics with exact ties, which are excluded from the unanimity count ",
      "rather than counted against it: ",
      paste(paired_stats$metric[paired_stats$n_ties > 0], collapse = ", "),
      ".\n", sep = "")
}
```

The paired t-test above assumes the five differences are roughly normal, which
five points cannot demonstrate. It is quoted as an effect estimate with an
interval rather than as a hypothesis test, and the interval is wide by
construction. A metric that comes out inconsistent is a finding in itself: it
means the two methods are close enough that recording-specific detail decides
the ranking.

# 6. Publication figures

## Figure: example traces

```{r fig-traces, fig.height = 7}
traces <- pull_all("traces") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")))

# Popen of the window actually shown, so a panel cannot quietly misrepresent the
# record it came from.
win_lab <- traces |>
  dplyr::group_by(phenotype, source) |>
  dplyr::summarise(win_popen = dplyr::first(win_popen),
                   win_start = dplyr::first(win_start),
                   x = min(time_ms), y = max(current), .groups = "drop") |>
  dplyr::mutate(label = sprintf("Popen %.2f (from %d)", win_popen, win_start))

p_traces <- ggplot(traces, aes(time_ms, current, colour = source)) +
  geom_line(linewidth = 0.2) +
  geom_text(data = win_lab, aes(x, y, label = label), hjust = 0, vjust = 1,
            size = 2.4, colour = "grey30", inherit.aes = FALSE) +
  facet_grid(phenotype ~ source, scales = "free_y") +
  scale_colour_manual(values = pal, guide = "none") +
  labs(x = "Time (ms)", y = "Current (pA)") +
  theme_channel()
p_traces
save_fig(p_traces, "fig_traces", width = 8, height = 1.4 * n_ph + 1)
```

The window is chosen by the same rule the Python previews use, so a panel shows
gating rather than whichever stretch the file happens to start with. Where no
window falls inside the target band — Phenotype_B is too quiet, Phenotype_D too
busy — the nearest available is used, and its Popen is printed so the gap is
visible rather than implied.

## Figure: idealisations

```{r fig-ideal, fig.height = 7}
p_ideal <- ggplot(traces, aes(time_ms, channels, colour = source)) +
  geom_step(linewidth = 0.3) +
  facet_grid(phenotype ~ source) +
  scale_colour_manual(values = pal, guide = "none") +
  labs(x = "Time (ms)", y = "Open channels") +
  theme_channel()
p_ideal
save_fig(p_ideal, "fig_idealisations", width = 8, height = 1.2 * n_ph + 1)
```

## Figure: all-points amplitude histograms

```{r fig-amp, fig.height = 6}
p_amp <- pull_all("amp") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  ggplot(aes(current, colour = source, fill = source)) +
  geom_density(alpha = 0.15, linewidth = 0.5, adjust = 0.6) +
  facet_wrap(~ phenotype, scales = "free", ncol = 2) +
  scale_colour_manual(values = pal) + scale_fill_manual(values = pal) +
  labs(x = "Current (pA)", y = "Density") +
  theme_channel()
p_amp
save_fig(p_amp, "fig_amplitude", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
```

## Figure: single-level noise

```{r fig-noise, fig.height = 6}
# Panels are labelled with the level they were computed at, so the legend can
# state it and a phenotype that offers no baseline is not silently different.
panel_label <- setNames(sprintf("%s (%s level, %d-sample segments)",
                                names(LEVEL_OF), LEVEL_OF,
                                NPERSEG_OF[names(LEVEL_OF)]),
                        names(LEVEL_OF))

p_psd <- pull_all("psd") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")),
                phenotype = panel_label[phenotype]) |>
  ggplot(aes(freq, psd, colour = source)) +
  geom_line(linewidth = 0.4) +
  geom_vline(xintercept = CFG$filter_hz, linetype = "dashed",
             colour = "grey60", linewidth = 0.3) +
  facet_wrap(~ phenotype, ncol = 2) +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_colour_manual(values = pal) +
  labs(x = "Frequency (Hz)", y = expression(PSD~(pA^2/Hz))) +
  theme_channel()

p_acf <- pull_all("acf") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")),
                phenotype = panel_label[phenotype]) |>
  ggplot(aes(lag_ms, acf, colour = source)) +
  geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.3) +
  geom_line(linewidth = 0.4) +
  facet_wrap(~ phenotype, ncol = 2) +
  scale_colour_manual(values = pal) +
  labs(x = "Lag (ms)", y = "Autocorrelation") +
  theme_channel()

p_psd
p_acf
save_fig(p_psd, "fig_psd", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
save_fig(p_acf, "fig_acf", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
```

## Figure: dwell times

Sigworth-Sine display: log abscissa, square-root ordinate. The diffusion arm's
`Channels` column is the conditioning idealisation it was handed, so it is a
known ground truth. The GAN emitted its own labels, so those are a network
output. They are plotted together for comparison and must not be described as
equivalent in the manuscript.

```{r fig-dwell, fig.height = 7}
p_dwell <- pull_all("dwell") |>
  dplyr::filter(dwell_ms > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  ggplot(aes(dwell_ms, colour = source)) +
  geom_area(aes(y = after_stat(sqrt(count)), fill = source),
            stat = "bin", bins = 40, alpha = 0.15,
            position = "identity", colour = NA) +
  geom_freqpoly(aes(y = after_stat(sqrt(count))), bins = 40, linewidth = 0.45) +
  facet_grid(phenotype ~ state, scales = "free_y") +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(x = "Dwell time (ms)", y = expression(sqrt(count))) +
  theme_channel()
p_dwell
save_fig(p_dwell, "fig_dwell", width = 8, height = 1.5 * n_ph + 1)
```

and bars
```{r dwellbars,  fig.height = 7}
d_dwell <- pull_all("dwell") |>
  dplyr::filter(dwell_ms > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")))

# Log-spaced breaks, so every bar is the same width on the log axis. Equal
# linear bins would be wide at short dwells and thin at long ones.
brk <- 10^seq(log10(min(d_dwell$dwell_ms)), log10(max(d_dwell$dwell_ms)),
              length.out = 41)

p_dwell <- ggplot(d_dwell, aes(dwell_ms, colour = source)) +
  geom_histogram(aes(y = after_stat(sqrt(count)), fill = source),
                 breaks = brk, alpha = 0.15, position = "identity", colour = NA) +
  geom_histogram(aes(y = after_stat(sqrt(count))), fill = NA,
                 breaks = brk, linewidth = 0.4, position = "identity") +
  facet_grid(phenotype ~ state, 
             scales = "free_y",
             labeller = labeller(
             phenotype = function(x) gsub("_", "\n", x, fixed = TRUE)
    )) +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(x = "Dwell time (ms)", y = expression(sqrt(count))) +
  theme_channel()+
  theme(
    strip.text.y = element_text(
      angle = 270,
      hjust = 0.5,
      vjust = 0.5
    )
  ) +
    guides(colour = guide_legend(override.aes = list(
    fill = scales::alpha(pal[c("Real", "Diffusion", "GAN")], 0.15),
    linewidth = 0.4)))
p_dwell

save_fig(p_dwell, "fig_dwell", width = 8, height = 1.5 * n_ph + 1)
```



## Bursts

```{r bursts, fig.height = 6}
# Two-component exponential mixture on the closed times by EM, then the
# Colquhoun-Sakmann equal-areas tcrit: the gap length at which as many brief
# closures are misread as between-burst as long ones are misread as within.
fit_tcrit <- function(closed_ms, n_iter = 200) {
  x <- closed_ms[is.finite(closed_ms) & closed_ms > 0]
  if (length(x) < 50) return(NA_real_)
  tau <- stats::quantile(x, c(0.25, 0.9), names = FALSE)
  if (!(tau[2] > tau[1])) return(NA_real_)
  a <- c(0.5, 0.5)
  for (i in seq_len(n_iter)) {
    d1 <- a[1] * stats::dexp(x, 1 / tau[1])
    d2 <- a[2] * stats::dexp(x, 1 / tau[2])
    w  <- d1 / (d1 + d2); w[!is.finite(w)] <- 0.5
    a   <- c(mean(w), 1 - mean(w))
    tau <- c(sum(w * x) / sum(w), sum((1 - w) * x) / sum(1 - w))
    if (any(!is.finite(tau)) || any(tau <= 0)) return(NA_real_)
  }
  o <- order(tau); tau <- tau[o]; a <- a[o]
  # A record with one closed component has no bursts to find, and forcing a
  # tcrit onto it invents them.
  if (tau[2] / tau[1] < 3 || min(a) < 0.02) return(NA_real_)
  f <- function(t) a[1] * exp(-t / tau[1]) - a[2] * (1 - exp(-t / tau[2]))
  lo <- tau[1] * 1e-3; hi <- tau[2]
  if (!is.finite(f(lo)) || !is.finite(f(hi)) || f(lo) * f(hi) > 0) return(NA_real_)
  stats::uniroot(f, c(lo, hi), tol = 1e-9)$root
}

burst_of <- function(dw, tcrit) {
  if (!is.finite(tcrit)) return(NULL)
  sep <- dw$state == "Closed" & dw$dwell_ms >= tcrit
  b   <- cumsum(sep)
  d <- dw[!sep, , drop = FALSE]; b <- b[!sep]
  if (!nrow(d)) return(NULL)
  # The first and last bursts are cut off by the ends of the record, so both
  # would drag the mean burst length down.
  inner <- b > min(b) & b < max(b)
  d <- d[inner, , drop = FALSE]; b <- b[inner]
  if (!nrow(d)) return(NULL)
  tibble::tibble(burst = b, state = d$state, ms = d$dwell_ms) |>
    dplyr::group_by(burst) |>
    dplyr::summarise(length_ms = sum(ms),
                     n_open    = sum(state == "Open"),
                     open_ms   = sum(ms[state == "Open"]),
                     n_gap     = sum(state == "Closed"),
                     gap_ms    = sum(ms[state == "Closed"]),
                     .groups = "drop") |>
    dplyr::filter(n_open > 0)
}

res_burst <- purrr::map(manifest$phenotype, function(ph) {
  arms <- recs[[ph]]; fs <- FS_MAIN[[ph]]
  dwl <- lapply(names(arms), function(a) dwell_times(arms[[a]]$channels, fs[[a]]))
  names(dwl) <- names(arms)
  # One tcrit per phenotype, taken from the real record and imposed on every
  # arm. Per-arm tcrits would make the burst counts incomparable.
  tcrit <- fit_tcrit(dwl$Real$dwell_ms[dwl$Real$state == "Closed"])
  per_burst <- purrr::map_dfr(names(arms), function(a) {
    b <- burst_of(dwl[[a]], tcrit)
    if (is.null(b) || !nrow(b)) return(NULL)
    dplyr::mutate(b, phenotype = ph, source = a)
  })
  if (!nrow(per_burst)) return(list(summary = NULL, per_burst = NULL))
  list(
    per_burst = per_burst,
    summary = per_burst |>
      dplyr::group_by(phenotype, source) |>
      dplyr::summarise(
        tcrit_ms = tcrit, n_bursts = dplyr::n(),
        `Burst length (ms)`       = mean(length_ms),
        `Openings per burst`      = mean(n_open),
        `Open time in burst (ms)` = sum(open_ms) / sum(n_open),
        `Gap in burst (ms)`       = if (sum(n_gap)) sum(gap_ms) / sum(n_gap) else NA_real_,
        `Burst Popen`             = sum(open_ms) / sum(length_ms),
        .groups = "drop") |>
      dplyr::mutate(`Bursts per s` = purrr::map_dbl(source, function(a)
        sum(per_burst$source == a) / (nrow(arms[[a]]) / fs[[a]]))))
})

burst      <- purrr::map_dfr(res_burst, "summary")
bursts_raw <- purrr::map_dfr(res_burst, "per_burst")

nice_kable(dplyr::select(burst, phenotype, source, tcrit_ms, n_bursts,
                           dplyr::everything()), digits = 4,
             caption = paste("Burst kinetics. tcrit is fitted on the",
                             "real record of each phenotype and applied to all",
                             "three arms; a phenotype whose closed times are",
                             "monoexponential yields no tcrit and drops out."))

burst_cmp <- burst |>
  dplyr::select(-tcrit_ms, -n_bursts) |>
  tidyr::pivot_longer(-c(phenotype, source), names_to = "metric") |>
  tidyr::pivot_wider(names_from = source, values_from = value) |>
  dplyr::filter(is.finite(Real), is.finite(Diffusion), is.finite(GAN)) |>
  dplyr::mutate(err_diffusion = abs(Diffusion - Real),
                err_gan       = abs(GAN - Real),
                favours = dplyr::case_when(
  err_diffusion < err_gan ~ "Diffusion",
  err_gan < err_diffusion ~ "GAN",
  TRUE                    ~ "Tie"
  ))

nice_kable(burst_cmp, digits = 4,
             caption = "Burst metrics, absolute error against the real record.")

nice_kable(
  burst_cmp |>
    dplyr::group_by(metric) |>
    dplyr::summarise(n = dplyr::n(),
                     n_favour_diffusion = sum(favours == "Diffusion"),
                     .groups = "drop"),
  caption = "Unanimity of the burst metrics across phenotypes.")

readr::write_csv(burst,     file.path(CFG$out_dir, "burst_kinetics.csv"))
readr::write_csv(burst_cmp, file.path(CFG$out_dir, "burst_comparison.csv"))
```

## Burst Figures
```{r burst_fig, fig.height = 7 }
p_burst <- bursts_raw |>
  dplyr::transmute(phenotype, source,
                   `Burst length (ms)` = length_ms,
                   `Openings per burst` = as.numeric(n_open)) |>
  tidyr::pivot_longer(-c(phenotype, source), names_to = "quantity") |>
  dplyr::filter(value > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  ggplot(aes(value, colour = source)) +
  geom_area(aes(y = after_stat(sqrt(count)), fill = source),
            stat = "bin", bins = 40, alpha = 0.15,
            position = "identity", colour = NA) +
  geom_freqpoly(aes(y = after_stat(sqrt(count))), bins = 40, linewidth = 0.5) +
  facet_grid(phenotype ~ quantity, scales = "free") +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(x = NULL, y = expression(sqrt(count))) +
  theme_channel()
p_burst

save_fig(p_burst, "fig_burst", width = 8, height = 1.5 * n_ph + 1)
readr::write_csv(bursts_raw, file.path(CFG$out_dir, "bursts_per_burst.csv"))
```
bars
```{r burst_bars,  fig.height = 7}
d_burst <- bursts_raw |>
  dplyr::transmute(phenotype, source,
                   `Burst length (ms)` = length_ms,
                   `Openings per burst` = as.numeric(n_open)) |>
  tidyr::pivot_longer(-c(phenotype, source), names_to = "quantity") |>
  dplyr::filter(value > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")))

# Breaks per quantity, since burst length and opening count span quite
# different ranges, and log-spaced so every bar is one width on the log axis.
log_breaks <- function(v, n = 41) {
  r <- range(v)
  if (r[1] == r[2]) return(NULL)
  10^seq(log10(r[1]), log10(r[2]), length.out = n)
}
brk_burst <- split(d_burst$value, d_burst$quantity) |> lapply(log_breaks)

layer_hist <- function(q) {
  d <- dplyr::filter(d_burst, quantity == q)
  list(
    geom_histogram(data = d, aes(y = after_stat(sqrt(count)), fill = source),
                   breaks = brk_burst[[q]], alpha = 0.15,
                   position = "identity", colour = NA),
    geom_histogram(data = d, aes(y = after_stat(sqrt(count))), fill = NA,
                   breaks = brk_burst[[q]], linewidth = 0.4,
                   position = "identity")
  )
}


p_burst <- ggplot(d_burst, aes(value, colour = source)) +
  lapply(names(brk_burst), layer_hist) +
  facet_grid(
    phenotype ~ quantity,
    scales = "free",
    labeller = labeller(
      phenotype = function(x) gsub("_", "\n", x, fixed = TRUE)
    )
  ) +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(
    x = NULL,
    y = expression(sqrt(count))
  ) +
  theme_channel() +
  theme(
    strip.text.y = element_text(
      angle = 270,
      hjust = 0.5,
      vjust = 0.5
    )
  ) +
  guides(
    colour = guide_legend(
      override.aes = list(
        fill = scales::alpha(
          pal[c("Real", "Diffusion", "GAN")],
          0.15
        ),
        linewidth = 0.4
      )
    )
  )


p_burst


save_fig(p_burst, "fig_burst", width = 8, height = 1.5 * n_ph + 1)
readr::write_csv(bursts_raw, file.path(CFG$out_dir, "bursts_per_burst.csv"))
```



## Label self-consistency

Carried over from `DiffuChannel_analysis.Rmd`, where it ran on one record at a
time. Each arm's current trace is idealised on its own, with a conventional
half-amplitude criterion and no reference to any other arm, and the result is
compared against the `Channels` column that arm shipped with.

This asks a different question from everything above. The GAN emitted its labels
and its trace together, so there is no external truth to score them against --
but there is an internal one: the labels should agree with what the arm's own
current is doing. The diffusion arm was handed its labels before it started, so
its agreement measures whether the generated current followed the conditioning
it was given. The real record supplies the floor, because expert idealisation
and a half-amplitude criterion do not agree perfectly either, and an arm cannot
reasonably be asked to do better than the seed.

```{r self-consistency}
res_selfcon <- purrr::map(manifest$phenotype, function(ph) {

  arms <- recs[[ph]]
  fs   <- FS_MAIN[[ph]]

  # First pass: try unsupervised EM on every arm.
  em_rec <- purrr::map(names(arms), function(a) {

    d <- arms[[a]]

    n_lev <- max(
      1,
      max(d$channels, na.rm = TRUE)
    )

    idealise_half_amplitude(
      d$current,
      n_levels = n_lev,
      polarity = POL_MAIN[[ph]]
    )
  })

  names(em_rec) <- names(arms)

  # If EM fails for any arm, use label-anchored CV for every
  # arm in this phenotype, keeping the comparison consistent.
  em_failed <- vapply(
    em_rec,
    function(x) all(is.na(x)),
    logical(1)
  )

  phenotype_fallback <- any(em_failed)

  if (phenotype_fallback) {
    failed_arms <- names(arms)[em_failed]

    method_note <- paste0(
      "Phenotype-wide fallback because unsupervised EM failed for: ",
      paste(failed_arms, collapse = ", ")
    )
  } else {
    method_note <- ""
  }

  # Second pass: calculate self-consistency using the method
  # selected for this entire phenotype.
  out <- purrr::map(names(arms), function(a) {

    d <- arms[[a]]

    n_lev <- max(
      1,
      max(d$channels, na.rm = TRUE)
    )

    lab <- d$channels > 0

    if (phenotype_fallback) {

      rec <- idealise_label_anchored_cv(
        current = d$current,
        labels = d$channels,
        n_levels = n_lev,
        polarity = POL_MAIN[[ph]],
        block_len = 1000L,
        trim = 0.10,
        min_state_n = 50L
      )

      ideal_method <- "Phenotype-wide label-anchored CV fallback"

    } else {

      rec <- em_rec[[a]]
      ideal_method <- "Unsupervised EM"
    }

    # This occurs only if the selected phenotype-wide
    # method still fails for this arm.
    if (all(is.na(rec))) {

      failure_reason <- attr(rec, "why")

      if (is.null(failure_reason)) {
        failure_reason <- "no reason was recorded"
      }

      return(list(
        summary = tibble::tibble(
          phenotype = ph,
          source = a,
          kappa = NA_real_,
          f1_open = NA_real_,
          popen_labels = mean(lab),
          popen_recovered = NA_real_,
          open_ms_labels = NA_real_,
          open_ms_recovered = NA_real_,
          flicker = NA,
          ideal_method = ideal_method,
          note = paste(
            method_note,
            "Selected method failed:",
            failure_reason
          )
        ),
        dwell = NULL
      ))
    }

    op <- rec > 0

    tp <- sum(lab & op)
    fp <- sum(!lab & op)
    fn <- sum(lab & !op)

    dl <- dwell_times(
      as.integer(lab),
      fs[[a]]
    )

    dr <- dwell_times(
      as.integer(op),
      fs[[a]]
    )

    mo <- function(x) {

      v <- x$dwell_ms[x$state == "Open"]

      if (length(v)) {
        mean(v)
      } else {
        NA_real_
      }
    }

    list(
      summary = tibble::tibble(
        phenotype = ph,
        source = a,
        kappa = cohen_kappa(lab, op),
        f1_open = if (2 * tp + fp + fn > 0) {
          2 * tp / (2 * tp + fp + fn)
        } else {
          NA_real_
        },
        popen_labels = mean(lab),
        popen_recovered = mean(op),
        open_ms_labels = mo(dl),
        open_ms_recovered = mo(dr),
        flicker = length(rle(op)$lengths) > 0.5 * length(op),
        ideal_method = ideal_method,
        note = method_note
      ),

      dwell = dplyr::bind_rows(
        dplyr::mutate(
          dl,
          origin = "Own labels"
        ),
        dplyr::mutate(
          dr,
          origin = "Recovered from current"
        )
      ) |>
        dplyr::mutate(
          phenotype = ph,
          source = a
        )
    )
  })

  list(
    summary = purrr::map_dfr(out, "summary"),
    dwell   = purrr::map_dfr(out, "dwell")
  )
})

self_consist <- purrr::map_dfr(res_selfcon, "summary")
sc_dwell     <- purrr::map_dfr(res_selfcon, "dwell")

if (any(self_consist$flicker %in% TRUE)) {
  warning("The recovered idealisation switches on more than half the samples ",
          "in: ", paste(sprintf("%s/%s", self_consist$phenotype[self_consist$flicker %in% TRUE],
                                self_consist$source[self_consist$flicker %in% TRUE]),
                        collapse = ", "),
          ". That is a wrong level estimate rather than a fast record, and the ",
          "kappa for those rows means nothing. Check the amplitude histogram.")
}

# A recovered Popen that is the complement of the labelled one is an inverted
# idealisation, not a disagreement, and its kappa is the negative of the real
# number. It should not happen now the polarity is imposed, so say so loudly.
inverted <- self_consist |>
  dplyr::filter(is.finite(kappa),
                abs(popen_recovered - (1 - popen_labels)) < 0.05,
                kappa < 0)
if (nrow(inverted)) {
  warning("The idealisation looks inverted in: ",
          paste(sprintf("%s/%s", inverted$phenotype, inverted$source),
                collapse = ", "), ". Check POL_MAIN for those phenotypes.")
}

nice_kable(self_consist, digits = 4,
             caption = paste("Label self-consistency. Each arm's",
                             "Channels column against a half-amplitude",
                             "idealisation of that same arm's current. The real",
                             "record is the floor, not a target of 1."))

sc_cmp <- self_consist |>
  dplyr::select(phenotype, source, kappa) |>
  tidyr::pivot_wider(names_from = source, values_from = kappa) |>
  dplyr::filter(
    is.finite(Real),
    is.finite(Diffusion),
    is.finite(GAN)
  ) |>
  dplyr::mutate(
  deficit_diffusion = Real - Diffusion,
  deficit_gan       = Real - GAN,

  error_diffusion = abs(Diffusion - Real),
  error_gan       = abs(GAN - Real),

  pct_error_diffusion = 100 * error_diffusion / abs(Real),
  pct_error_gan       = 100 * error_gan / abs(Real),

  favours = dplyr::case_when(
    error_diffusion < error_gan ~ "Diffusion",
    error_gan < error_diffusion ~ "GAN",
    TRUE                        ~ "Tie"
  )
)

nice_kable(sc_cmp, digits = 4,
             caption = paste("Shortfall in kappa against the real",
                             "record's own floor. A negative deficit means the",
                             "arm's labels agree with its current better than",
                             "the expert idealisation agrees with the seed,",
                             "which is a sign the arm is too clean rather than",
                             "a sign it is good."))

cat(sprintf("\nPhenotypes where the diffusion arm is the more self-consistent: %d of %d.\n",
            sum(sc_cmp$favours == "Diffusion"), nrow(sc_cmp)))

readr::write_csv(self_consist, file.path(CFG$out_dir, "self_consistency.csv"))
readr::write_csv(sc_cmp, file.path(CFG$out_dir, "self_consistency_comparison.csv"))
```

The gap between the two dwell distributions below is the missed-event loss: the
brief openings and closures that a half-amplitude criterion cannot recover from
the current, however good the trace is. Being able to measure it at all is what
having the conditioning idealisation buys.

```{r fig-selfcon, fig.width = 10, fig.height = 7}
pal_origin <- c("Own labels" = "#1b1b1b",
                "Recovered from current" = "#c0392b")

d_sc <- sc_dwell |>
  dplyr::filter(dwell_ms > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")),
                origin = factor(origin, levels = names(pal_origin)))

brk_sc <- 10^seq(log10(min(d_sc$dwell_ms)), log10(max(d_sc$dwell_ms)),
                 length.out = 41)

p_selfcon <- ggplot(d_sc, aes(dwell_ms, colour = origin)) +
  geom_histogram(aes(y = after_stat(sqrt(count)), fill = origin),
                 breaks = brk_sc, alpha = 0.15, position = "identity",
                 colour = NA) +
  geom_histogram(aes(y = after_stat(sqrt(count))), fill = NA,
                 breaks = brk_sc, linewidth = 0.4, position = "identity") +
  facet_grid(
    phenotype ~ source + state,
    scales = "free_y",
    labeller = labeller(
      phenotype = function(x) gsub("_", "\n", x, fixed = TRUE)
    )
  ) +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal_origin) +
  scale_fill_manual(values = pal_origin, guide = "none") +
  labs(x = "Dwell time (ms)", y = expression(sqrt(count))) +
  theme_channel() +
  theme(
    strip.text.y = element_text(
      angle = 270,
      hjust = 0.5,
      vjust = 0.5
    )
  ) +
  guides(
    colour = guide_legend(
      override.aes = list(
        fill = scales::alpha(pal_origin, 0.15),
        linewidth = 0.4
      )
    )
  )

p_selfcon

save_fig(p_selfcon, "fig_self_consistency", width = 10, height = 1.5 * n_ph + 1)
readr::write_csv(sc_dwell, file.path(CFG$out_dir, "self_consistency_dwells.csv"))
```

## Dwell time constants

Section 3.6 of the manuscript asks for exponential time constants, not mean
dwells. Both are fitted here from the same two dwell sets the self-consistency
step already produces: the arm's own labels, which for the diffusion arm are
the exact conditioning ground truth, and the idealisation recovered from that
arm's current by a half-amplitude criterion. The difference between them is the
missed-event loss.

Where the column reports one component, two could not be supported and the mean
dwell is the only defensible summary for that row. That is a property of the
record, not a failure of the fit: a phenotype whose closed dwells average three
or four samples has no exponential structure left to resolve.

```{r dwell-tau}
tau_fits <- sc_dwell |>
  dplyr::group_by(phenotype, source, state, origin) |>
  dplyr::group_modify(~ fit_exp_mixture(.x$dwell_ms)) |>
  dplyr::ungroup() |>
  dplyr::mutate(origin = dplyr::recode(origin,
                  `Own labels` = "true",
                  `Recovered from current` = "recovered"))

nice_kable(tau_fits, digits = 4,
           caption = paste("Dwell time constants. 'true' is the arm's own",
                           "label column, which for the diffusion arm is the",
                           "conditioning ground truth; 'recovered' is the",
                           "idealisation recovered from that arm's current.",
                           "n_components is 1 where a two-component fit could",
                           "not be supported."))

readr::write_csv(tau_fits, file.path(CFG$out_dir, "dwell_tau_fits.csv"))

cat(sprintf(
  "\nTwo-component fits supportable in %d of %d phenotype/state/origin cells.\n",
  sum(tau_fits$n_components == 2), nrow(tau_fits)))
```

## Figure: slow envelope

```{r fig-env, fig.height = 6}
p_env <- pull_all("env") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  ggplot(aes(value, colour = source, fill = source)) +
  geom_density(alpha = 0.15, linewidth = 0.5) +
  facet_wrap(~ phenotype, scales = "free", ncol = 2) +
  scale_colour_manual(values = pal) + scale_fill_manual(values = pal) +
  labs(x = "Envelope current (pA)", y = "Density") +
  theme_channel()
p_env
save_fig(p_env, "fig_envelope", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
```

The density above says whether the envelope has the right spread; it says
nothing about whether it wanders on the right timescale, because a distribution
has no time axis. The panel below is the envelope as a time course, which is
what the manuscript's Figure 4A shows. It was previously produced only by
`DiffuChannel_analysis.Rmd`, one record at a time.

```{r fig-env-time, fig.height = 7}
# idx counts blocks of slow_factor samples, so seconds need the rate of the arm
# that block came from. Joined from a lookup rather than computed per row: the
# envelope tables run to tens of thousands of rows per phenotype.
fs_tbl <- purrr::imap_dfr(FS_MAIN, function(fs, ph)
  tibble::tibble(phenotype = ph, source = names(fs), fs = unname(fs)))

env_time <- pull_all("env") |>
  dplyr::left_join(fs_tbl, by = c("phenotype", "source")) |>
  dplyr::mutate(time_s = idx * CFG$slow_factor / fs,
                source = factor(source, levels = c("Real", "Diffusion", "GAN")))

if (any(!is.finite(env_time$time_s)))
  stop("An envelope row has no rate, so its time axis is undefined. Check that ",
       "FS_MAIN names every arm present in the env tables.")

p_env_time <- ggplot(env_time, aes(time_s, value, colour = source)) +
  geom_line(linewidth = 0.3) +
  facet_grid(phenotype ~ source, scales = "free_y",
             labeller = labeller(
               phenotype = function(x) gsub("_", "\n", x, fixed = TRUE))) +
  scale_colour_manual(values = pal, guide = "none") +
  labs(x = "Time (s)", y = "Envelope current (pA)") +
  theme_channel() +
  theme(strip.text.y = element_text(angle = 270, hjust = 0.5, vjust = 0.5))

p_env_time
save_fig(p_env_time, "fig_envelope_time", width = 8, height = 1.4 * n_ph + 1)
```

## Figure: manifold projections (optional)

```{r fig-manifold, eval = CFG$do_manifold, fig.width = 7.5, fig.height = 12}
if (!all(vapply(c("uwot", "Rtsne"), requireNamespace, logical(1), quietly = TRUE))) {
  stop("Set CFG$do_manifold = FALSE or install uwot and Rtsne.")
}

proj <- purrr::map_dfr(seq_len(nrow(manifest)), function(i) {
  row <- manifest[i, ]
  d <- list(Real = read_record(row$raw), Diffusion = read_record(row$ddpm))
  if (!is.na(row$gan)) d$GAN <- read_record(row$gan)
  Wp <- lapply(d, function(x) make_windows(x$current, CFG$proj_win, CFG$n_windows_proj))
  X <- do.call(rbind, Wp)
  lab <- rep(names(Wp), vapply(Wp, nrow, integer(1)))
  X <- scale(X); X[!is.finite(X)] <- 0
  um <- uwot::umap(X, n_neighbors = 15, min_dist = 0.1, n_components = 2)
  ts <- Rtsne::Rtsne(X, dims = 2, perplexity = 30, check_duplicates = FALSE,
                     pca = TRUE)$Y
  dplyr::bind_rows(
    tibble::tibble(x = um[, 1], y = um[, 2], source = lab, method = "UMAP"),
    tibble::tibble(x = ts[, 1], y = ts[, 2], source = lab, method = "t-SNE")
  ) |> dplyr::mutate(phenotype = row$phenotype)
})

proj <- proj |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  dplyr::group_by(phenotype, method) |>
  dplyr::mutate(x = scales::rescale(x), y = scales::rescale(y)) |>
  dplyr::ungroup()

# One arm per column, over a grey ghost of the full cloud so each panel still
# shows where the other two sat. Rescaling per phenotype lets the three columns
# share fixed scales, which free scales previously prevented.
plot_proj <- function(d) {
  ggplot(d, aes(x, y)) +
    geom_point(
      data = dplyr::select(d, -source),
      colour = "grey88",
      size = 0.7
    ) +
    geom_point(
      aes(colour = source),
      size = 1.1,
      alpha = 0.85
    ) +
    facet_grid(
      phenotype ~ source,
      labeller = labeller(
        phenotype = function(x) gsub("_", "\n", x)
      )
    ) +
    scale_colour_manual(values = pal) +
    guides(
      colour = guide_legend(
        override.aes = list(size = 4, alpha = 1)
      )
    ) +
    labs(x = NULL, y = NULL) +
    theme_channel(base_size = 11) +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      legend.text = element_text(size = 12),
      legend.key.size = unit(1.5, "lines"),
      panel.spacing = unit(0.7, "lines"),
      strip.text.y = element_text(
        angle = 270,
        hjust = 0.5,
        vjust = 0.5
      )
    )
}
p_proj_umap <- plot_proj(dplyr::filter(proj, method == "UMAP"))
p_proj_tsne <- plot_proj(dplyr::filter(proj, method == "t-SNE"))
p_proj_umap
p_proj_tsne

save_fig(p_proj_umap, "fig_manifold_umap", width = 7.5, height = 2.3 * n_ph + 1)
save_fig(p_proj_tsne, "fig_manifold_tsne", width = 7.5, height = 2.3 * n_ph + 1)
readr::write_csv(proj, file.path(CFG$out_dir, "manifold_coordinates.csv"))
```

## Cluster separation of the projections

The projections above are judged by eye, which is not a measurement. This tests
whether the real and synthetic point clouds are separable in the embedding,
using the permutation approach of ClusterSignificance as in the DeepGANnel
paper. A *large* p is the favourable outcome here: it means the two clouds
cannot be told apart, which is the opposite of the usual reading and is worth
stating explicitly in the legend.

The test is run within a phenotype and within a projection method, since the
coordinates of one embedding mean nothing in another.

```{r cluster-test, eval = CFG$do_manifold && isTRUE(CFG$do_cluster_test)}
if (!requireNamespace("ClusterSignificance", quietly = TRUE)) {
  message("ClusterSignificance is not installed, so [[cluster_p]] cannot be ",
          "filled. It is a Bioconductor package:\n",
          '  BiocManager::install("ClusterSignificance")\n',
          "Either install it, or delete the ClusterSignificance sentence and ",
          "reference [49] from Methods 2.9 and the claim from Section 3.8.")
  cluster_tbl <- NULL
} else {
  # pcp() takes a two-column matrix and a grouping factor and returns a
  # p-value for every pairwise group comparison, so one call per phenotype
  # per method covers both Real-vs-Diffusion and Real-vs-GAN.
  cluster_one <- function(d) {
    mat <- as.matrix(d[, c("x", "y")])
    grp <- factor(as.character(d$source))
    if (dplyr::n_distinct(grp) < 2 || nrow(mat) < 10) return(NULL)
    pe <- ClusterSignificance::permute(
      mat = mat, groups = grp, iter = CFG$n_perm_cluster, projmethod = "pcp")
    p <- ClusterSignificance::pvalue(pe)
    tibble::tibble(comparison = names(p), p = as.numeric(p))
  }

  cluster_tbl <- proj |>
    dplyr::group_by(phenotype, method) |>
    dplyr::group_modify(~ {
      r <- tryCatch(cluster_one(.x),
                    error = function(e)
                      tibble::tibble(comparison = "failed",
                                     p = NA_real_,
                                     note = conditionMessage(e)))
      if (is.null(r)) tibble::tibble(comparison = NA_character_, p = NA_real_) else r
    }) |>
    dplyr::ungroup()

  nice_kable(cluster_tbl, digits = 4,
             caption = paste("ClusterSignificance permutation p-values for",
                             "separation between the point clouds in each",
                             "embedding. A large p means the clouds are not",
                             "separable, which is the favourable result."))

  readr::write_csv(cluster_tbl,
                   file.path(CFG$out_dir, "cluster_significance.csv"))

  if (any(is.na(cluster_tbl$p)))
    warning("Some cluster tests returned no p-value. Read the note column ",
            "before quoting anything from this table.")
}
```

## Figure: generation block joins

Long records are assembled from overlapping blocks bridged by context-carry
inpainting, so the one artefact this pipeline could plausibly introduce is a
seam. If the inpainting works, the distribution of sample-to-sample differences
in the neighbourhood of a join is indistinguishable from that within a block.
If it does not, it shows up here rather than in review.

This ran per record in `DiffuChannel_analysis.Rmd` and is done here for every
phenotype, on the **uncropped** diffusion arm, since cropping would move the
joins. Only the diffusion arm has joins at all: the GAN emits a record in one
pass and the seed record was never assembled from anything.

```{r fig-seams, fig.height = 7}
step <- CFG$block_len - CFG$context_len
half <- CFG$context_len %/% 2

# Joins of one record, dropping any that fall within half a context length of
# either end, where there is not enough material on both sides to compare.
joins_of <- function(n) {
  # seq() errors rather than returning nothing when the record is shorter
  # than one block, so that case is caught before it is reached.
  if (n <= step) return(numeric(0))
  b <- seq(step, n - 1L, by = step)
  b[b > CFG$context_len & b < (n - CFG$context_len)]
}

seam_one <- function(d, ph) {
  n <- nrow(d)
  b <- joins_of(n)
  diffs <- diff(d$current)

  if (!length(b)) {
    return(list(
      summary = tibble::tibble(phenotype = ph, n_samples = n, n_joins = 0L,
                               ks_seam = NA_real_, p_seam = NA_real_,
                               note = "single block, no joins to test"),
      dens = NULL, trace = NULL, joins = numeric(0)))
  }

  near <- unique(unlist(lapply(b, function(i) seq(i - half, i + half))))
  near <- near[near >= 1 & near <= length(diffs)]
  far  <- setdiff(seq_along(diffs), near)

  ks <- ks_safe(diffs[near], diffs[far])

  # The two sets are compared at equal size, so the density panel cannot be
  # read as a difference in smoothness that is really a difference in n.
  far_s <- sample(far, min(length(far), length(near)))

  # Decimated for drawing only. The statistic above uses every sample.
  dec <- max(1L, n %/% 20000L)
  keep <- seq(1L, n, by = dec)

  list(
    summary = tibble::tibble(
      phenotype = ph, n_samples = n, n_joins = length(b),
      ks_seam = unname(ks$statistic), p_seam = unname(ks$p.value),
      note = sprintf("%d samples near a join, %d within blocks",
                     length(near), length(far))),
    dens = dplyr::bind_rows(
      tibble::tibble(d = diffs[near],  where = "Near block join"),
      tibble::tibble(d = diffs[far_s], where = "Within block")) |>
      dplyr::mutate(phenotype = ph),
    trace = tibble::tibble(phenotype = ph, time_s = d$time[keep],
                           current = d$current[keep]),
    joins = d$time[b])
}

res_seam <- purrr::imap(recs_full_diffusion, function(d, ph) seam_one(d, ph))

seam_tbl   <- purrr::map_dfr(res_seam, "summary")
seam_dens  <- purrr::map_dfr(res_seam, "dens")
seam_trace <- purrr::map_dfr(res_seam, "trace")
seam_lines <- purrr::imap_dfr(res_seam, function(r, ph)
  if (length(r$joins)) tibble::tibble(phenotype = ph, time_s = r$joins) else NULL)

nice_kable(seam_tbl, digits = 4,
           caption = paste("Block joins in the diffusion arm. ks_seam compares",
                           "sample-to-sample differences within half a carried",
                           "context of a join against those within blocks. A",
                           "small statistic means the join is not detectable in",
                           "the output."))

readr::write_csv(seam_tbl, file.path(CFG$out_dir, "seam_test.csv"))
```

```{r fig-seam-panels, fig.height = 7}
if (nrow(seam_trace)) {
  p_long <- ggplot(seam_trace, aes(time_s, current)) +
    geom_line(linewidth = 0.15, colour = pal[["Diffusion"]]) +
    { if (nrow(seam_lines))
        geom_vline(data = seam_lines, aes(xintercept = time_s),
                   linetype = "dashed", colour = "grey55", linewidth = 0.3) } +
    facet_wrap(~ phenotype, scales = "free", ncol = 1) +
    labs(x = "Time (s)", y = "Current (pA)") +
    theme_channel()
  print(p_long)
  save_fig(p_long, "fig_long_record", width = 8, height = 1.4 * n_ph + 1)
}

if (nrow(seam_dens)) {
  p_seam <- ggplot(seam_dens, aes(d, colour = where)) +
    geom_density(linewidth = 0.6, adjust = 0.8) +
    facet_wrap(~ phenotype, scales = "free", ncol = 2) +
    scale_colour_manual(values = c("Near block join" = "#c0392b",
                                   "Within block"    = "#1b1b1b")) +
    labs(x = "Sample-to-sample difference (pA)", y = "Density") +
    theme_channel()
  print(p_seam)
  save_fig(p_seam, "fig_seam_test", width = 8,
           height = 2.2 * ceiling(n_ph / 2) + 1)
} else {
  message("No phenotype's diffusion record spans more than one block, so there ",
          "are no joins to test. Check CFG$block_len against the record length.")
}
```

# 7. Numbers for the manuscript

Rows keyed to `[[placeholder]]` names. Pooled values are medians across
phenotypes, with the range, because five records do not support a mean and an
interval being read as an estimate.

```{r placeholders}
fmt <- function(x, d = 4) formatC(x, digits = d, format = "g")

pooled <- scored |>
  dplyr::group_by(metric, method) |>
  dplyr::summarise(median = median(value), lo = min(value), hi = max(value),
                   .groups = "drop")

placeholders <- dplyr::bind_rows(
  tibble::tibble(placeholder = "n_phenotypes", value = as.character(n_ph),
                 note = "paired seed records"),
  # Two different totals, previously conflated under one name: the seed
  # material that exists, and the subset actually scored after every arm of a
  # phenotype was cropped to its shortest. The manuscript wants the first.
  tibble::tibble(placeholder = "n_samples_total",
                 value = format(sum(seed_inventory$n_rows), big.mark = ","),
                 note = "seed samples across all phenotypes, before cropping"),
  tibble::tibble(placeholder = "n_samples_analysed",
                 value = format(sum(info$n_real), big.mark = ","),
                 note = "seed samples actually scored, after cropping each phenotype to its shortest arm"),
  # Rate before unique, or floating-point dust splits one rate into several.
  tibble::tibble(placeholder = "sample_khz",
                 value = paste(unique(round(sort(info$fs_real) / 1000, 3)),
                               collapse = ", "),
                 note = "acquisition rate(s), kHz"),
  pooled |>
    dplyr::transmute(
      placeholder = paste0(gsub("[^A-Za-z0-9]+", "_", tolower(metric)), "_",
                           tolower(method)),
      value = sprintf("%s (%s to %s)", fmt(median), fmt(lo), fmt(hi)),
      note = paste("median across phenotypes, range in brackets")),
  paired_stats |>
    dplyr::transmute(
      placeholder = paste0("unanimity_", gsub("[^A-Za-z0-9]+", "_", tolower(metric))),
      value = sprintf("%d/%d favour Diffusion", n_favour_diff, n),
      note = verdict),
  # Dwell time constants for the diffusion arm, recovered against true, over
  # the phenotypes where a two-component fit held. Reported per phenotype
  # rather than pooled: a time constant is not comparable across records the
  # way a distance is.
  {
    tf <- tau_fits |>
      dplyr::filter(source == "Diffusion", n_components == 2L)
    if (nrow(tf)) {
      tf |>
        dplyr::mutate(ph = gsub("^Phenotype_", "", phenotype)) |>
        dplyr::group_by(state, origin) |>
        dplyr::summarise(
          value = paste(sprintf("%s %.3g/%.3g", ph, tau_fast_ms, tau_slow_ms),
                        collapse = ", "),
          n_ph = dplyr::n(), .groups = "drop") |>
        dplyr::transmute(
          placeholder = sprintf("tau_%s_%s", tolower(state), origin),
          value = value,
          note = sprintf("fast/slow tau in ms, per phenotype, %d of %d phenotypes supported two components",
                         n_ph, dplyr::n_distinct(tau_fits$phenotype)))
    } else {
      tibble::tibble(placeholder = "tau_fits", value = "NOT SUPPORTED",
                     note = "no phenotype supported a two-component dwell fit")
    }
  },
  # Cluster separation, if the optional test ran.
  {
    if (exists("cluster_tbl") && !is.null(cluster_tbl) &&
        any(is.finite(cluster_tbl$p))) {
      cp <- dplyr::filter(cluster_tbl, is.finite(p))
      tibble::tibble(
        placeholder = "cluster_p",
        value = sprintf("%s (%s to %s)", fmt(median(cp$p)),
                        fmt(min(cp$p)), fmt(max(cp$p))),
        note = sprintf("ClusterSignificance permutation p, median over %d tests; large p means not separable",
                       nrow(cp)))
    } else {
      tibble::tibble(placeholder = "cluster_p", value = "NOT RUN",
                     note = "ClusterSignificance absent or the test failed")
    }
  },
  # Seam test. Reported as median and range like everything else pooled, and
  # only over the phenotypes that actually had a join to test.
  {
    s <- dplyr::filter(seam_tbl, is.finite(ks_seam))
    if (nrow(s))
      tibble::tibble(
        placeholder = "ks_seam",
        value = sprintf("%s (%s to %s)", fmt(median(s$ks_seam)),
                        fmt(min(s$ks_seam)), fmt(max(s$ks_seam))),
        note = sprintf("median across %d phenotypes, %d joins tested in total",
                       nrow(s), sum(s$n_joins)))
    else
      tibble::tibble(placeholder = "ks_seam", value = "NOT RUN",
                     note = "no phenotype spanned more than one block")
  }
)

nice_kable(placeholders, caption = "Values keyed to the manuscript placeholders.")
```

## What this comparison does and does not settle

Both methods are scored identically above, so the numbers are comparable as
measures of how much the synthetic data resemble the real record. That is the
only question they answer.

They do not settle the question the paper is about. The GAN emits its raw trace
and its `Channels` column together, so its labels are a network output; the
diffusion pipeline is handed its labels before it starts and never touches them.
A GAN record could match the real data better on every metric here and still be
the less useful training set, because a downstream idealiser trained against it
is being scored on another network's opinion rather than on truth. If the GAN
wins on similarity, say so plainly and make the ground-truth argument on its own
terms.

# Session information

```{r session}
sessionInfo()
```
