---
title: "DiffuChannel: pooled analysis across seed records"
author: "Barrett-Jolley lab"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    theme: flatly
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 7, fig.height = 4.5, dpi = 150)
```

# Purpose

`DiffuChannel_analysis.Rmd` writes one tidy file per seed record into `runs_dir`.
This notebook pools them, so that a claim rests on several recordings rather than
on whichever one was analysed first.

The design is **paired**: every method sees the same seed record, so the useful
quantity is the within-seed difference between methods, not the spread across
seeds. Seed records differ enormously in noise, open probability and duration,
and that between-seed variation would swamp any method difference if the runs
were treated as independent samples.

```{r packages}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(purrr)

set.seed(42)
```

# Configuration

```{r config}
META <- list(
  runs_dir = "./runs",        # folder written by the per-record notebook
  out_dir  = "./figures_meta",
  # Methods to place side by side. The first is treated as the reference arm
  # in the paired comparisons.
  compare  = c("Diffusion", "GAN")
)

dir.create(META$out_dir, showWarnings = FALSE, recursive = TRUE)

theme_meta <- function(base = 11) {
  theme_classic(base_size = base) +
    theme(strip.background = element_blank(),
          strip.text = element_text(face = "bold", hjust = 0),
          legend.position = "top", legend.title = element_blank(),
          axis.text = element_text(colour = "black"))
}

pal <- c("Diffusion" = "#c0392b", "GAN" = "#8e44ad",
         "Single-stage" = "#2980b9", "Real" = "#1b1b1b",
         "Real floor" = "#7f8c8d", "White noise" = "#bdc3c7")
```

# 1. Collect the runs

```{r collect}
files <- list.files(META$runs_dir, pattern = "__metrics\\.csv$", full.names = TRUE)

if (!length(files)) {
  stop("No run files in ", META$runs_dir, ". Run DiffuChannel_analysis.Rmd ",
       "once per seed record first; each run writes one file here.")
}

all_rows <- purrr::map_dfr(files, readr::read_csv, show_col_types = FALSE)

runs <- sort(unique(all_rows$run_id))
cat(sprintf("%d run(s) found: %s\n", length(runs), paste(runs, collapse = ", ")))

if (anyDuplicated(paste(all_rows$run_id, all_rows$metric, all_rows$method,
                        all_rows$kind))) {
  warning("Duplicate run_id/metric/method rows. Two runs probably shared a ",
          "run_id and one has overwritten part of the other. Check ",
          META$runs_dir, " and set CFG$run_id explicitly.")
}
```

## Provenance

```{r provenance}
info <- all_rows |>
  dplyr::filter(kind == "info") |>
  dplyr::select(run_id, metric, text) |>
  tidyr::pivot_wider(names_from = metric, values_from = text)

knitr::kable(info, caption = "What each run was computed from.")
readr::write_csv(info, file.path(META$out_dir, "meta_provenance.csv"))
```

A run that used a different `seq_len`, `slow_factor` or sampling rate is not
strictly comparable with the others. The table above is there to be read, not
skipped: pooling runs that were computed differently is the commonest way a
meta-analysis goes quietly wrong.

```{r consistency}
for (v in c("seq_len", "slow_factor")) {
  if (v %in% names(info) && dplyr::n_distinct(info[[v]]) > 1) {
    warning("Runs differ in ", v, " (", paste(unique(info[[v]]), collapse = ", "),
            "). They are being pooled anyway, but the comparison is no longer ",
            "like for like.")
  }
}
```

# 2. Put the metrics on a comparable footing

Metrics are not all oriented the same way. Distances should be small; the
permutation p should be large; and summary statistics such as the closed-state
standard deviation should match the real record rather than approach zero. For
that last group the comparable quantity is the absolute difference from the
real reference, which the per-record notebook exports alongside them.

```{r score}
metrics <- dplyr::filter(all_rows, kind == "metric")

refs <- metrics |>
  dplyr::filter(direction == "reference", method == "Real") |>
  dplyr::select(run_id, metric, ref = value)

scored <- metrics |>
  dplyr::filter(direction %in% c("lower", "higher", "match_real")) |>
  dplyr::left_join(refs, by = c("run_id", "metric")) |>
  dplyr::mutate(
    score = dplyr::case_when(
      direction == "match_real" ~ abs(value - ref),
      TRUE                      ~ value
    ),
    # after scoring, every metric is either "smaller is better" or the single
    # "larger is better" case
    better = ifelse(direction == "higher", "higher", "lower")
  ) |>
  dplyr::select(run_id, metric, method, direction, better, value, ref, score)

knitr::kable(
  scored |> dplyr::arrange(metric, method, run_id) |>
    dplyr::select(metric, method, run_id, value, ref, score),
  digits = 6, caption = "Raw values and the comparable score derived from them."
)
```

# 3. Paired comparison, one line per seed record

```{r slopegraph, fig.height = 6}
pairable <- scored |>
  dplyr::filter(method %in% META$compare) |>
  dplyr::group_by(metric, run_id) |>
  dplyr::filter(dplyr::n_distinct(method) == length(META$compare)) |>
  dplyr::ungroup()

if (!nrow(pairable)) {
  message("No metric has all of ", paste(META$compare, collapse = " and "),
          " present in the same run, so there is nothing to pair.")
} else {
  p_slope <- ggplot(pairable,
                    aes(factor(method, levels = META$compare), score,
                        group = run_id)) +
    geom_line(colour = "grey55", linewidth = 0.4) +
    geom_point(aes(colour = method), size = 2.2) +
    facet_wrap(~ metric, scales = "free_y") +
    scale_colour_manual(values = pal, guide = "none") +
    labs(x = NULL, y = "Score (see note above)") +
    theme_meta()
  print(p_slope)
  ggsave(file.path(META$out_dir, "meta_paired.png"), p_slope,
         width = 9, height = 6, dpi = 300)
}
```

Each line is one seed record. A consistent slope across every line is the
result worth reporting; lines that cross mean the ranking depends on which
recording you happened to analyse.

# 4. Pooled summary

```{r summary}
wide <- pairable |>
  dplyr::select(metric, run_id, better, method, score) |>
  tidyr::pivot_wider(names_from = method, values_from = score)

if (nrow(wide)) {
  a <- META$compare[1]; b <- META$compare[2]

  meta_summary <- wide |>
    dplyr::mutate(diff = .data[[a]] - .data[[b]]) |>
    dplyr::group_by(metric, better) |>
    dplyr::summarise(
      n_runs        = dplyr::n(),
      median_a      = median(.data[[a]]),
      median_b      = median(.data[[b]]),
      median_diff   = median(diff),
      min_diff      = min(diff),
      max_diff      = max(diff),
      # how many seeds favour the first method, given the metric's polarity
      n_favour_a    = sum(if (first(better) == "lower") diff < 0 else diff > 0),
      unanimous     = n_favour_a %in% c(0, dplyr::n()),
      .groups = "drop"
    ) |>
    dplyr::rename(!!paste0("median_", a) := median_a,
                  !!paste0("median_", b) := median_b,
                  !!paste0("n_favour_", a) := n_favour_a)

  knitr::kable(meta_summary, digits = 6,
               caption = paste("Pooled comparison.", a, "minus", b,
                               "within each seed record."))
  readr::write_csv(meta_summary, file.path(META$out_dir, "meta_summary.csv"))
  readr::write_csv(scored, file.path(META$out_dir, "meta_scores_long.csv"))
} else {
  meta_summary <- NULL
  message("Nothing to summarise.")
}
```

# 5. What can and cannot be concluded from this many runs

```{r power}
n_runs <- length(runs)
cat(sprintf("Runs pooled: %d\n\n", n_runs))

# Smallest two-sided p a sign test can return for n paired observations, i.e.
# the p you get when every single run favours the same method.
min_p_sign <- function(n) min(1, 2 * 0.5^n)

cat(sprintf(paste("With %d paired runs, the smallest two-sided p a sign test",
                  "can produce is %.3f,\neven if every run favours the same",
                  "method.\n"), n_runs, min_p_sign(n_runs)))

if (min_p_sign(n_runs) > 0.05) {
  cat("\nThat is above 0.05, so no significance test on these data can reach\n",
      "significance regardless of how large or how consistent the effect is.\n",
      "Reporting a p value here would be misleading rather than conservative.\n",
      "Report instead: the per-run values, the direction, and whether it is\n",
      "unanimous. Those are the honest summaries at this sample size.\n", sep = "")
} else {
  cat("\nA sign test is at least capable of reaching p < 0.05 here.\n")
  if (!is.null(meta_summary)) {
    sign_tests <- wide |>
      dplyr::mutate(diff = .data[[META$compare[1]]] - .data[[META$compare[2]]]) |>
      dplyr::group_by(metric) |>
      dplyr::summarise(
        n = dplyr::n(),
        p_sign = binom.test(sum(diff < 0), dplyr::n())$p.value,
        .groups = "drop"
      )
    knitr::kable(sign_tests, digits = 4, caption = "Sign test per metric.")
  }
}
```

```{r direction}
if (!is.null(meta_summary)) {
  a <- META$compare[1]
  dir_tbl <- meta_summary |>
    dplyr::transmute(
      Metric = metric,
      `Seeds pooled` = n_runs,
      `Seeds favouring` = sprintf("%d/%d %s",
                                  .data[[paste0("n_favour_", a)]], n_runs, a),
      Verdict = dplyr::case_when(
        .data[[paste0("n_favour_", a)]] == n_runs ~ paste("consistent:", a),
        .data[[paste0("n_favour_", a)]] == 0      ~ paste("consistent:", META$compare[2]),
        TRUE                                      ~ "inconsistent across seeds"
      )
    )
  knitr::kable(dir_tbl, caption = "Direction of effect, which is what this
               sample size supports reporting.")
  readr::write_csv(dir_tbl, file.path(META$out_dir, "meta_direction.csv"))
}
```

Three seed records is enough to show that a result is not an accident of one
recording. It is not enough to estimate an effect size, and a confidence
interval computed from three paired differences will be too wide to constrain
anything. If a metric comes out inconsistent, that is a finding in itself and
worth saying: it means the two methods are close enough that recording-specific
detail decides the ranking.

# Session information

```{r session}
sessionInfo()
```
